---
title: "Preparing clean spatial data"
format: html
---

# Introduction

The spatial data will be loaded and saved as topoJSON files. These files include topology.

The data will need to be cleaned before saving. This involves identifying gaps between polygons and sliver polygons, and then rectifying the polygons to ensure no gaps and no sliver polygons.

The spatial data to be cleaned include the Nigerian state boundaries (which are an important reference layer for Nigeria) and the livelihood zones.

```{r results='hide', warning=FALSE, echo=FALSE, message=FALSE}

library(tidyverse)
library(terra)
library(tidyterra)
library(readODS)
library(ggplot2)
library(NMdata)
library(rmapshaper)
library(geojsonio)
```



# Loading the data

## Load the Livelihood Zones

```{r}
#| label: lz_zones
#| message: false
#| warning: false

vect_lz_geo <- vect("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/FEWS NET LZ/NG_LHZ_2018/NG_LHZ_2018.shp")
vect_lz_geo <- makeValid(vect_lz_geo)
vect_lz_geo$LZCODE <- as.factor(vect_lz_geo$LZCODE)
levels(vect_lz_geo$LZCODE) # Check levels
plot(vect_lz_geo, main = "FEWS NET 2018 Livelihood Zones", levels=TRUE)

```

## Load the States

```{r}
#| label: states
#| message: false
#| warning: false

vect_state_geo <- vect("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NGA Admin 2019-2025 OSGOF-Ehealth-UNCS/nga_admin1.geojson")
vect_state_geo <- makeValid(vect_state_geo)
vect_state_geo$adm1_pcode <- as.factor(vect_state_geo$adm1_pcode)
levels(vect_state_geo$adm1_pcode) # Check levels
plot(vect_state_geo, main = "Nigeria administrative level 1 boundaries (COD-AB)", levels=TRUE)

```

## Create Geo-political zone

```{r}

adm1_pcode <- c("NG001", "NG002", "NG003", "NG004", "NG005", "NG006", "NG007", "NG008", "NG009", "NG010", "NG011", "NG012", "NG013", "NG014", "NG015", "NG016", "NG017", "NG018", "NG019", "NG020", "NG021", "NG022", "NG023", "NG024", "NG025", "NG026", "NG027", "NG028", "NG029", "NG030", "NG031", "NG032", "NG033", "NG034", "NG035", "NG036", "NG037")

geopolitical_zone<- c("South East", "North East", "South South", "South East", "North East", "South South", "North Central", "North East", "South South", "South South", "South East", "South South", "South West", "South East", "North Central", "North East", "South East", "North West", "North West", "North West", "North West", "North West", "North Central", "North Central", "South West", "North Central", "North Central", "South West", "South West", "South West", "South West", "North Central", "South South", "North West", "North East", "North East", "North West")

df_gpz <- tibble(adm1_pcode, geopolitical_zone)

vect_gpz_geo <- dplyr::left_join(vect_state_geo, df_gpz, by = c("adm1_pcode")) |> terra::aggregate(by = "geopolitical_zone")

vect_gpz_geo <- makeValid(vect_gpz_geo)
vect_gpz_geo$geopolitical_zone <- as.factor(vect_gpz_geo$geopolitical_zone)
levels(vect_gpz_geo$geopolitical_zone) # Check levels
plot(vect_gpz_geo, main = "Nigeria administrative geopolitical zone boundaries (COD-AB)", levels=TRUE)

```


## Load the Agro-ecological zones


```{r}
#| label: aez
#| message: false
#| warning: false

vect_aez_geo <- vect("spatial_data/IITA_AEZ_LGP_2020_SRTM_RCL_NGA.geojson")
vect_aez_geo <- makeValid(vect_aez_geo)
vect_aez_geo$AEZ <- as.factor(vect_aez_geo$AEZ)
levels(vect_aez_geo$AEZ) # Check levels
plot(vect_aez_geo, main = "Nigeria Agro-ecological zones based on 2020 LGP and elevation", levels=TRUE)

```


# Defining a CRS for Nigeria

For a single CRS covering all of Nigeria, in metres, commonly used for modern GIS analysis, the best choice is:
WGS 84 / UTM Zone 32N (EPSG:32632)
This is the most widely used option because:

Nigeria falls mostly within UTM Zone 32N (roughly 6°E–12°E longitude)
It uses metres as the unit
It is based on WGS 84, making it compatible with GPS and modern datasets
False easting ensures that all coordinates are positive.

```{r}
#| label: wkt_nig_utm
#| message: false
#| warning: false


wkt_nig_utm <- paste0("PROJCS[\"WGS 84 / UTM zone 32N\",
    GEOGCS[\"WGS 84\",
        DATUM[\"WGS_1984\",
            SPHEROID[\"WGS 84\",6378137,298.257223563,
                AUTHORITY[\"EPSG\",\"7030\"]],
            AUTHORITY[\"EPSG\",\"6326\"]],
        PRIMEM[\"Greenwich\",0,
            AUTHORITY[\"EPSG\",\"8901\"]],
        UNIT[\"degree\",0.0174532925199433,
            AUTHORITY[\"EPSG\",\"9122\"]],
        AUTHORITY[\"EPSG\",\"4326\"]],
    PROJECTION[\"Transverse_Mercator\"],
    PARAMETER[\"latitude_of_origin\",0],
    PARAMETER[\"central_meridian\",9],
    PARAMETER[\"scale_factor\",0.9996],
    PARAMETER[\"false_easting\",1000000],
    PARAMETER[\"false_northing\",0],
    UNIT[\"metre\",1,
        AUTHORITY[\"EPSG\",\"9001\"]],
    AXIS[\"Easting\",EAST],
    AXIS[\"Northing\",NORTH]]")


```


# Changing CRS for cleaning

```{r}
#| label: project
#| message: false
#| warning: false
#| 
# project the livelihood zones to wkt_nig_utm

vect_lz_utm <-  project(vect_lz_geo, wkt_nig_utm)
plot(vect_lz_utm, main = "FEWS NET 2018 Livelihood Zones UTM", levels=TRUE)

# project the states to wkt_nig_utm

vect_state_utm <-  project(vect_state_geo, wkt_nig_utm)
plot(vect_state_utm, main = "Nigeria administrative level 1 boundaries (COD-AB) UTM", levels=TRUE)

# project the gpz to wkt_nig_utm

vect_gpz_utm <-  project(vect_gpz_geo, wkt_nig_utm)
plot(vect_gpz_utm, main = "Nigeria administrative geopolitical zones boundaries (COD-AB) UTM", levels=TRUE)

# project the aez to wkt_nig_utm

vect_aez_utm <-  project(vect_aez_geo, wkt_nig_utm)
plot(vect_aez_utm, main = "Nigeria Agro-ecological zones based on 2020 LGP and elevation UTM", levels=TRUE)


```


# Check for slivers (high perimeter-to-area ratio)

```{r}
#| label: lz_slivers
#| message: false
#| warning: false


areas_lz  <- expanse(vect_lz_utm) # area in m² 

perims_lz <- perim(vect_lz_utm)                 # perimeter in m

ratios_lz <- perims_lz^2 / areas_lz

slivers_lz <- vect_lz_utm[ratios_lz > 1000, ]     # tune threshold to taste
crs(slivers_lz, describe = TRUE)

print(paste("Number of sliver polygons (lz):", nrow(slivers_lz)))

print(slivers_lz)

if (nrow(slivers_lz) > 0) {
ggplot(slivers_lz) +
  geom_spatvector(aes(fill= LZCODE), colour = "grey") +
  geom_spatvector_text(aes(label = LZCODE), size = 3) +
  coord_sf(crs = sf::st_crs(wkt_nig_utm), datum = sf::st_crs(wkt_nig_utm))
}    

```

```{r}
#| label: state_slivers
#| message: false
#| warning: false


areas_state  <- expanse(vect_state_utm) # area in m² 

perims_state <- perim(vect_state_utm)                 # perimeter in m

ratios_state <- perims_state^2 / areas_state

slivers_state <- vect_state_utm[ratios_state > 1000, ]     # tune threshold to taste
crs(slivers_state, describe = TRUE)

print(paste("Number of sliver polygons (state):", nrow(slivers_state)))

print(slivers_state)

if (nrow(slivers_state) > 0) {
  ggplot(slivers_state) +
    geom_spatvector(aes(fill = adm1_pcode), colour = "grey") +
    geom_spatvector_text(aes(label = adm1_pcode), size = 3) +
    coord_sf(crs = sf::st_crs(wkt_nig_utm),
             datum = sf::st_crs(wkt_nig_utm))
}  

```

```{r}
#| label: aez_slivers
#| message: false
#| warning: false


areas_aez  <- expanse(vect_aez_utm) # area in m² 

perims_aez <- perim(vect_aez_utm)                 # perimeter in m

ratios_aez <- perims_aez^2 / areas_aez

slivers_aez <- vect_aez_utm[ratios_aez > 1000, ]     # tune threshold to taste
crs(slivers_aez, describe = TRUE)

print(paste("Number of sliver polygons (aez):", nrow(slivers_aez)))

print(slivers_aez)

if (nrow(slivers_aez) > 0) {
  ggplot(slivers_aez) +
    geom_spatvector(aes(fill = AEZ), colour = "grey") +
    geom_spatvector_text(aes(label = AEZ), size = 3) +
    coord_sf(crs = sf::st_crs(wkt_nig_utm),
             datum = sf::st_crs(wkt_nig_utm))
}  

```

# Check for gaps by comparing union to convex hull

```{r}
#| label: lz_gaps
#| message: false
#| warning: false


unioned_lz <- aggregate(vect_lz_utm)     # dissolves all to single polygon
# Fill all holes to get a "solid" version
filled_lz <- fillHoles(unioned_lz)
# Gaps are what's in the filled version but not in the union
gaps_lz <- erase(filled_lz, unioned_lz)
gaps_lz <- gaps_lz[expanse(gaps_lz) > 10, ]  # filter gaps smaller than 10m²
print(gaps_lz)

ext_gaps_lz <- ext(gaps_lz)
print(ext_gaps_lz)
ext_vect_lz_utm <- ext(vect_lz_utm)
print(ext_vect_lz_utm)

# Number of rows (features)
nrow(gaps_lz)

# Number of actual geometric parts
nrow(geom(gaps_lz))

# Check which rows have valid geometries
is_valid_lz <- is.valid(gaps_lz)
is_empty_lz <- is.empty(gaps_lz)

print(is_valid_lz)
print(is_empty_lz)

if (!is_empty_lz) {
  # Check number of parts per row
  parts_per_row_lz <- sapply(seq_len(nrow(gaps_lz)), function(i)
    nrow(geom(gaps_lz[i, ])))
  print(parts_per_row_lz)
  
  g_lz <- geom(gaps_lz)
  head(g_lz, 40)
  
  
  n_parts_lz <- max(g_lz[, "part"])
  
  # Reconstruct each part as its own SpatVector and combine
  parts_list_lz <- lapply(seq_len(n_parts_lz), function(i) {
    part_coords_lz <- g_lz[g_lz[, "part"] == i, ]
    vect(part_coords_lz, type = "polygons", crs = crs(gaps_lz))
  })
  
  gaps_single_lz <- do.call(rbind, parts_list_lz)
  nrow(gaps_single_lz)
  gaps_single_lz <- gaps_single_lz[expanse(gaps_single_lz) > 10, ]  # filter gaps smaller than 10m²
  
  
  
  for (i in seq_len(nrow(gaps_single_lz))) {
    # Define your zoom extent explicitly
    xmin_lz <- ext(gaps_single_lz[i, ])[1]
    xmax_lz <- ext(gaps_single_lz[i, ])[2]
    ymin_lz <- ext(gaps_single_lz[i, ])[3]
    ymax_lz <- ext(gaps_single_lz[i, ])[4]
    
    # Crop the background polygons to the visible extent
    # This ensures only polygons (or parts of polygons) in view are labeled
    zoom_ext_lz <- ext(xmin_lz, xmax_lz, ymin_lz, ymax_lz)
    background_cropped_lz <- crop(vect_lz_utm, zoom_ext_lz)
    
    # Get vertices as a dataframe
    bg_vertices_lz <- as.data.frame(geom(background_cropped_lz))
    fg_vertices_lz <- as.data.frame(geom(gaps_single_lz[i, ]))
    
    # geom() returns columns: geom, part, x, y, hole
    print(paste("Gap poly:", i))
    print(fg_vertices_lz)
    print(paste("Background polys:", i))
    print(bg_vertices_lz)
    
    
    p_lz <- ggplot() +
      geom_spatvector(data = background_cropped_lz,
                      fill = "lightgrey",
                      colour = "grey") +
      geom_spatvector_text(
        data = background_cropped_lz,
        aes(label = LZCODE),
        size = 3,
        color = "black"
      ) +
      geom_spatvector(data = gaps_single_lz[i, ],
                      fill = "red",
                      colour = "darkred") +
      xlim(xmin_lz, xmax_lz) +
      ylim(ymin_lz, ymax_lz) +
      coord_sf(crs = sf::st_crs(wkt_nig_utm),
               datum = sf::st_crs(wkt_nig_utm)) +
      labs(title = paste("Polygon", i, "Gap = ", expanse(gaps_single_lz[i, ]), "m2"))
    
    print(p_lz)
  }
} else {
  print(paste("No gaps in lz dataset"))
}

```


```{r}
#| label: state_gaps
#| message: false
#| warning: false


unioned_state <- aggregate(vect_state_utm)     # dissolves all to single polygon
# Fill all holes to get a "solid" version
filled_state <- fillHoles(unioned_state)
# Gaps are what's in the filled version but not in the union
gaps_state <- erase(filled_state, unioned_state)
gaps_state <- gaps_state[expanse(gaps_state) > 10, ]  # filter gaps smaller than 10m²
print(gaps_state)

ext_gaps_state <- ext(gaps_state)
print(ext_gaps_state)
ext_vect_state_utm <- ext(vect_state_utm)
print(ext_vect_state_utm)

# Number of rows (features)
nrow(gaps_state)

# Number of actual geometric parts
nrow(geom(gaps_state))

# Check which rows have valid geometries
is_valid_state <- is.valid(gaps_state)
is_empty_state <- is.empty(gaps_state)

print(is_valid_state)
print(is_empty_state)


if (!is_empty_state) {
  # Check number of parts per row
  parts_per_row_state <- sapply(seq_len(nrow(gaps_state)), function(i)
    nrow(geom(gaps_state[i, ])))
  print(parts_per_row_state)
  
  g_state <- geom(gaps_state)
  head(g_state, 40)
  
  
  n_parts_state <- max(g_state[, "part"])
  
  # Reconstruct each part as its own SpatVector and combine
  parts_list_state <- lapply(seq_len(n_parts_state), function(i) {
    part_coords_state <- g_state[g_state[, "part"] == i, ]
    vect(part_coords_state,
         type = "polygons",
         crs = crs(gaps_state))
  })
  
  gaps_single_state <- do.call(rbind, parts_list_state)
  nrow(gaps_single_state)
  gaps_single_state <- gaps_single_state[expanse(gaps_single_state) > 10, ]  # filter gaps smaller than 10m²
  
  
  
  for (i in seq_len(nrow(gaps_single_state))) {
    # Define your zoom extent explicitly
    xmin_state <- ext(gaps_single_state[i, ])[1]
    xmax_state <- ext(gaps_single_state[i, ])[2]
    ymin_state <- ext(gaps_single_state[i, ])[3]
    ymax_state <- ext(gaps_single_state[i, ])[4]
    
    # Crop the background polygons to the visible extent
    # This ensures only polygons (or parts of polygons) in view are labeled
    zoom_ext_state <- ext(xmin_state, xmax_state, ymin_state, ymax_state)
    background_cropped_state <- crop(vect_state_utm, zoom_ext_state)
    
    # Get vertices as a dataframe
    bg_vertices_state <- as.data.frame(geom(background_cropped_state))
    fg_vertices_state <- as.data.frame(geom(gaps_single_state[i, ]))
    
    # geom() returns columns: geom, part, x, y, hole
    print(paste("Gap poly:", i))
    print(fg_vertices_state)
    print(paste("Background polys:", i))
    print(bg_vertices_state)
    
    
    p_state <- ggplot() +
      geom_spatvector(data = background_cropped_state,
                      fill = "lightgrey",
                      colour = "grey") +
      geom_spatvector_text(
        data = background_cropped_state,
        aes(label = adm1_pcode),
        size = 3,
        color = "black"
      ) +
      geom_spatvector(data = gaps_single_state[i, ],
                      fill = "red",
                      colour = "darkred") +
      xlim(xmin_state, xmax_state) +
      ylim(ymin_state, ymax_state) +
      coord_sf(crs = sf::st_crs(wkt_nig_utm),
               datum = sf::st_crs(wkt_nig_utm)) +
      labs(title = paste("Polygon", i, "Gap = ", expanse(gaps_single_state[i, ]), "m2"))
    
    print(p_state)
  }
} else {print(paste("No gaps in state dataset")) }


```

```{r}
#| label: aez_gaps
#| message: false
#| warning: false


unioned_aez <- aggregate(vect_aez_utm)     # dissolves all to single polygon
# Fill all holes to get a "solid" version
filled_aez <- fillHoles(unioned_aez)
# Gaps are what's in the filled version but not in the union
gaps_aez <- erase(filled_aez, unioned_aez)
gaps_aez <- gaps_aez[expanse(gaps_aez) > 10, ]  # filter gaps smaller than 10m²
print(gaps_aez)

ext_gaps_aez <- ext(gaps_aez)
print(ext_gaps_aez)
ext_vect_aez_utm <- ext(vect_aez_utm)
print(ext_vect_aez_utm)

# Number of rows (features)
nrow(gaps_aez)

# Number of actual geometric parts
nrow(geom(gaps_aez))

# Check which rows have valid geometries
is_valid_aez <- is.valid(gaps_aez)
is_empty_aez <- is.empty(gaps_aez)

print(is_valid_aez)
print(is_empty_aez)


if (!is_empty_aez) {
  # Check number of parts per row
  parts_per_row_aez <- sapply(seq_len(nrow(gaps_aez)), function(i)
    nrow(geom(gaps_aez[i, ])))
  print(parts_per_row_aez)
  
  g_aez <- geom(gaps_aez)
  head(g_aez, 40)
  
  
  n_parts_aez <- max(g_aez[, "part"])
  
  # Reconstruct each part as its own SpatVector and combine
  parts_list_aez <- lapply(seq_len(n_parts_aez), function(i) {
    part_coords_aez <- g_aez[g_aez[, "part"] == i, ]
    vect(part_coords_aez,
         type = "polygons",
         crs = crs(gaps_aez))
  })
  
  gaps_single_aez <- do.call(rbind, parts_list_aez)
  nrow(gaps_single_aez)
  gaps_single_aez <- gaps_single_aez[expanse(gaps_single_aez) > 10, ]  # filter gaps smaller than 10m²
  
  
  
  for (i in seq_len(nrow(gaps_single_aez))) {
    # Define your zoom extent explicitly
    xmin_aez <- ext(gaps_single_aez[i, ])[1]
    xmax_aez <- ext(gaps_single_aez[i, ])[2]
    ymin_aez <- ext(gaps_single_aez[i, ])[3]
    ymax_aez <- ext(gaps_single_aez[i, ])[4]
    
    # Crop the background polygons to the visible extent
    # This ensures only polygons (or parts of polygons) in view are labeled
    zoom_ext_aez <- ext(xmin_aez, xmax_aez, ymin_aez, ymax_aez)
    background_cropped_aez <- crop(vect_aez_utm, zoom_ext_aez)
    
    # Get vertices as a dataframe
    bg_vertices_aez <- as.data.frame(geom(background_cropped_aez))
    fg_vertices_aez <- as.data.frame(geom(gaps_single_aez[i, ]))
    
    # geom() returns columns: geom, part, x, y, hole
    print(paste("Gap poly:", i))
    print(fg_vertices_aez)
    print(paste("Background polys:", i))
    print(bg_vertices_aez)
    
    
    p_aez <- ggplot() +
      geom_spatvector(data = background_cropped_aez,
                      fill = "lightgrey",
                      colour = "grey") +
      geom_spatvector_text(
        data = background_cropped_aez,
        aes(label = AEZ),
        size = 3,
        color = "black"
      ) +
      geom_spatvector(data = gaps_single_aez[i, ],
                      fill = "red",
                      colour = "darkred") +
      xlim(xmin_aez, xmax_aez) +
      ylim(ymin_aez, ymax_aez) +
      coord_sf(crs = sf::st_crs(wkt_nig_utm),
               datum = sf::st_crs(wkt_nig_utm)) +
      labs(title = paste("Polygon", i, "Gap = ", expanse(gaps_single_aez[i, ]), "m2"))
    
    print(p_aez)
  }
} else {print(paste("No gaps in aez dataset")) }


```

# Check for overlaps

```{r}
#| label: lz_overlaps
#| message: false
#| warning: false


# Check for overlaps between polygon pairs
overlaps_lz <- relate(vect_lz_utm, relation = "overlaps")  # returns a matrix
any(overlaps_lz)

if (any(overlaps_lz)) {
# Get the actual overlap geometries
overlap_pairs_lz <- which(overlaps_lz, arr.ind = TRUE)
overlap_pairs_lz <- overlap_pairs_lz[overlap_pairs_lz[,1] < overlap_pairs_lz[,2], ]  # remove duplicates

overlap_geoms_lz <- do.call(rbind, lapply(1:nrow(overlap_pairs_lz), function(i) {
  intersect(vect_lz_utm[overlap_pairs_lz[i,1]], vect_lz_utm[overlap_pairs_lz[i,2]])
}))


p_lz <- ggplot() +
    geom_spatvector(data = vect_lz_utm, fill = "lightgrey", colour = "grey") +
    geom_spatvector_text(data = vect_lz_utm, aes(label = LZCODE), size = 3,
    color = "black") +
    geom_spatvector(data = overlap_geoms_lz, fill = "red", colour = "darkred") +
    coord_sf(crs = sf::st_crs(wkt_nig_utm), datum = sf::st_crs(wkt_nig_utm))
  
  print(p_lz)
} else {print(paste("No overlaps in lz dataset")) }
```


```{r}
#| label: state_overlaps
#| message: false
#| warning: false


# Check for overlaps between polygon pairs
overlaps_state <- relate(vect_state_utm, relation = "overlaps")  # returns a matrix

if (any(overlaps_state)) {
  # Get the actual overlap geometries
  overlap_pairs_state <- which(overlaps_state, arr.ind = TRUE)
  overlap_pairs_state <- overlap_pairs_state[overlap_pairs_state[, 1] < overlap_pairs_state[, 2], ]  # remove duplicates
  
  overlap_geoms_state <- do.call(rbind, lapply(1:nrow(overlap_pairs_state), function(i) {
    intersect(vect_state_utm[overlap_pairs_state[i, 1]], vect_state_utm[overlap_pairs_state[i, 2]])
  }))
  
  
  p_state <- ggplot() +
    geom_spatvector(data = vect_state_utm,
                    fill = "lightgrey",
                    colour = "grey") +
    geom_spatvector_text(
      data = vect_state_utm,
      aes(label = adm1_pcode),
      size = 3,
      color = "black"
    ) +
    geom_spatvector(data = overlap_geoms_state,
                    fill = "red",
                    colour = "darkred") +
    coord_sf(crs = sf::st_crs(wkt_nig_utm),
             datum = sf::st_crs(wkt_nig_utm))
  
  print(p_state)
  
} else {print(paste("No overlaps in state dataset")) }

```


```{r}
#| label: aez_overlaps
#| message: false
#| warning: false


# Check for overlaps between polygon pairs
overlaps_aez <- relate(vect_aez_utm, relation = "overlaps")  # returns a matrix

if (any(overlaps_aez)) {
  # Get the actual overlap geometries
  overlap_pairs_aez <- which(overlaps_aez, arr.ind = TRUE)
  overlap_pairs_aez <- overlap_pairs_aez[overlap_pairs_aez[, 1] < overlap_pairs_aez[, 2], ]  # remove duplicates
  
  overlap_geoms_aez <- do.call(rbind, lapply(1:nrow(overlap_pairs_aez), function(i) {
    intersect(vect_aez_utm[overlap_pairs_aez[i, 1]], vect_aez_utm[overlap_pairs_aez[i, 2]])
  }))
  
  
  p_aez <- ggplot() +
    geom_spatvector(data = vect_aez_utm,
                    fill = "lightgrey",
                    colour = "grey") +
    geom_spatvector_text(
      data = vect_aez_utm,
      aes(label = AEZ),
      size = 3,
      color = "black"
    ) +
    geom_spatvector(data = overlap_geoms_aez,
                    fill = "red",
                    colour = "darkred") +
    coord_sf(crs = sf::st_crs(wkt_nig_utm),
             datum = sf::st_crs(wkt_nig_utm))
  
  print(p_aez)
  
} else {print(paste("No overlaps in aez dataset")) }

```

# Clean and Convert to topojson

```{r}
#| label: lz_convert
#| message: false
#| warning: false


# 1. Fix validity
vect_lz_utm <- makeValid(vect_lz_utm)

# 2. Convert to sf for snap, rmapshaper, clean, convert back

sf_lz_utm    <- sf::st_as_sf(vect_lz_utm)
sf_lz_utm_snap <- sf::st_snap(sf_lz_utm, sf_lz_utm, tolerance = 0.1)  # tune tolerance in metres

sf_lz_utm_snap_clean <- sf::st_make_valid(sf_lz_utm_snap)  # re-validate after snapping

# 3. (Optional) Simplify topology-aware
sf_lz_utm_snap_clean_simple <- ms_simplify(sf_lz_utm_snap_clean, keep = 0.5, keep_shapes = TRUE)

# 4. Convert back to lat/long
vect_lz_utm_snap_clean_simple <- vect(sf_lz_utm_snap_clean_simple)
vect_lz_geo_snap_clean_simple <-  project(vect_lz_utm_snap_clean_simple, "epsg:4326")

sf_lz_geo_snap_clean_simple    <- sf::st_as_sf(vect_lz_geo_snap_clean_simple)

# 5. Write to TopoJSON
topojson_write(geojson_list(sf_lz_geo_snap_clean_simple), file =
                 "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/FEWS NET LZ/NG_LHZ_2018/NG_LHZ_2018.topojson", overwrite = TRUE, geometry = "polygon", object_name = "NG_LHZ_2018")

```

```{r}
#| label: state_convert
#| message: false
#| warning: false


# 1. Fix validity
vect_state_utm <- makeValid(vect_state_utm)

# 2. Convert to sf for snap, rmapshaper, clean, convert back

sf_state_utm    <- sf::st_as_sf(vect_state_utm)
sf_state_utm_snap <- sf::st_snap(sf_state_utm, sf_state_utm, tolerance = 0.1)  # tune tolerance in metres

sf_state_utm_snap_clean <- sf::st_make_valid(sf_state_utm_snap)  # re-validate after snapping

# 3. (Optional) Simplify topology-aware
sf_state_utm_snap_clean_simple <- ms_simplify(sf_state_utm_snap_clean, keep = 1, keep_shapes = TRUE)

# 4. Convert back to lat/long
vect_state_utm_snap_clean_simple <- vect(sf_state_utm_snap_clean_simple)
vect_state_geo_snap_clean_simple <-  project(vect_state_utm_snap_clean_simple, "epsg:4326")

sf_state_geo_snap_clean_simple    <- sf::st_as_sf(vect_state_geo_snap_clean_simple)

# 5. Write to TopoJSON
topojson_write(geojson_list(sf_state_geo_snap_clean_simple), file =                  "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NGA Admin 2019-2025 OSGOF-Ehealth-UNCS/nga_admin1_topo.topojson", overwrite = TRUE, geometry = "polygon", object_name = "NG_ADMIN1_2019_25")

```


```{r}
#| label: gpz_convert
#| message: false
#| warning: false

# 1. Fix validity
vect_gpz_utm <- makeValid(vect_gpz_utm)

# 2. Convert to sf for snap, rmapshaper, clean, convert back

sf_gpz_utm    <- sf::st_as_sf(vect_gpz_utm)
sf_gpz_utm_snap <- sf::st_snap(sf_gpz_utm, sf_gpz_utm, tolerance = 0.1)  # tune tolerance in metres

sf_gpz_utm_snap_clean <- sf::st_make_valid(sf_gpz_utm_snap)  # re-validate after snapping

# 3. (Optional) Simplify topology-aware
sf_gpz_utm_snap_clean_simple <- ms_simplify(sf_gpz_utm_snap_clean, keep = 1, keep_shapes = TRUE)

# 4. Convert back to lat/long
vect_gpz_utm_snap_clean_simple <- vect(sf_gpz_utm_snap_clean_simple)
vect_gpz_geo_snap_clean_simple <-  project(vect_gpz_utm_snap_clean_simple, "epsg:4326")

sf_gpz_geo_snap_clean_simple    <- sf::st_as_sf(vect_gpz_geo_snap_clean_simple)

# 5. Write to TopoJSON
topojson_write(geojson_list(sf_gpz_geo_snap_clean_simple), file =                  "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NGA Admin 2019-2025 OSGOF-Ehealth-UNCS/nga_gpz_topo.topojson", overwrite = TRUE, geometry = "polygon", object_name = "NG_GPZ_2019_25")


```

```{r}
#| label: aez_convert
#| message: false
#| warning: false


# 1. Fix validity
vect_aez_utm <- makeValid(vect_aez_utm)

# 2. Convert to sf for snap, rmapshaper, clean, convert back

sf_aez_utm    <- sf::st_as_sf(vect_aez_utm)
sf_aez_utm_snap <- sf::st_snap(sf_aez_utm, sf_aez_utm, tolerance = 0.1)  # tune tolerance in metres

sf_aez_utm_snap_clean <- sf::st_make_valid(sf_aez_utm_snap)  # re-validate after snapping

# 3. (Optional) Simplify topology-aware
sf_aez_utm_snap_clean_simple <- ms_simplify(sf_aez_utm_snap_clean, keep = 1, keep_shapes = TRUE)

# 4. Convert back to lat/long
vect_aez_utm_snap_clean_simple <- vect(sf_aez_utm_snap_clean_simple)
vect_aez_geo_snap_clean_simple <-  project(vect_aez_utm_snap_clean_simple, "epsg:4326")

sf_aez_geo_snap_clean_simple    <- sf::st_as_sf(vect_aez_geo_snap_clean_simple)

# 5. Write to TopoJSON
topojson_write(geojson_list(sf_aez_geo_snap_clean_simple), file =                  "spatial_data/IITA_AEZ_LGP_2020_SRTM_RCL_NGA_topo.topojson", overwrite = TRUE, geometry = "polygon", object_name = "AEZ")

```