---
title: "Nigeria Crop Consumption"
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    page-layout: full
    embed-resources: true
    self-contained-math: true
    toc-expand: true
    theme: united
    code-fold: true
    code-summary: "Show the code"
    css: css/style.css
bibliography: biofort_quarto_better.bib
---

```{r results='hide', warning=FALSE, echo=FALSE, message=FALSE}

library(tidyverse)
library(terra)
library(tidyterra)
library(readODS)
library(ggplot2)
library(ggtext)
library(tidyverse)
library(readODS)
library(kableExtra)
library(DT)
library(geojsonsf)

```
# Household level GHS data

The raw data files are available for use if necessary, the summaries per geopolitical zone are sufficient for some aspects of the atlas, but other indicators - such as the consumption of individual crop food products - require access to the raw household level survey data.

::: callout-note
## Relevant Datasets in the GHS

There are five GHS datasets that are relevant for the theme of crop food product consumption:

*sect6b_plantingw5.csv*\
Data collected through Post-Planting Household Questionnaire Section 6B (Food Consumption). This section records household-food item level information on the quantity and value of food consumed within the household during the past 7 days, as well as the quantity and value of food purchased during the past 30 days.

*sect6c_plantingw5.csv*\
Data collected through Post-Planting Household Questionnaire Section 6C (Aggregate Food Consumption). This section records household-food item level information on the number of days the household consumed each food group item during the last 7 days.

*sect5b_harvestw5.csv*\
Data collected through Post- Harvest Household Questionnaire Section 5B (Food Consumption). This section records household-food item level information on the quantity and value of food consumed within the household during the past 7 days, as well as the quantity and value of food purchased during the past 30 days.

*sect5c_harvestw5.csv*\
Data collected through Post-Harvest Household Questionnaire Section 5C (Aggregate Food Consumption). This section records household-food item level information on the number of days the household consumed each food group item during the last 7 days.

*sect3d_harvestw5.csv*\
Data collected through Post-Harvest Household Questionnaire Section 3D (Women Dietary Diversity). This section records women-level information on women’s nutrition diet for the previous day.
:::

## Description of datasets

#### Number of households


```{r}
#| label: GHS_household_data
#| message: false
#| warning: false
#| cache: false
#| results: asis

# 1. Post planting survey

df_GHS_secta_plantingw5 <- read.csv(
  "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Planting Wave 5/Household/secta_plantingw5.csv")

zone_sample_pp <- df_GHS_secta_plantingw5 |>
  group_by(
    zone
  ) |>
  summarise(
    n_households_pp = n(),
    .groups = "drop"
  ) |>
  arrange(zone)

state_sample_pp <- df_GHS_secta_plantingw5 |>
  group_by(
    zone,
    state
  ) |>
  summarise(
    n_households_pp = n(),
    .groups = "drop"
  ) |> 
  arrange(zone, state)


# 2. Post harvest survey

df_GHS_secta_harvestw5 <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Harvest Wave 5/Household/secta_harvestw5.csv")

zone_sample_ph <- df_GHS_secta_harvestw5 |>
  group_by(
    zone
  ) |>
  summarise(
    n_households_ph = n(),
    .groups = "drop"
  ) |>
  arrange(zone)

state_sample_ph <- df_GHS_secta_harvestw5 |>
  group_by(
    zone,
    state
  ) |>
  summarise(
    n_households_ph = n(),
    .groups = "drop"
  ) |>
  arrange(zone, state)

```

There were 5,067 households in the post planting survey, and 4771 households in the post harvest survey.

These were distributed more or less equally among the zones (@tbl-zone), but there were large differences between states. The state with the highest number of households was Abia, with 258 households in the post planting survey and 244 in the post harvest survey (@tbl-state). The state with the fewest number of households surveyed was Nasarawa, with only 24 households in both surveys.

```{r}
#| label: tbl-zone
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "hh_zone"

zone_sample <- zone_sample_pp |> left_join(zone_sample_ph, by = "zone")

zone_sample |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "# of households (post planting survey)", "# of households (post harvest survey)"),
    caption = "Households per zone",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

```{r}
#| label: tbl-state
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "hh_state"

state_sample <- state_sample_pp |> left_join(select(state_sample_ph, -zone), by = "state")

state_sample |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "State", "# of households (post planting survey)", "# of households (post harvest survey)"),
    caption = "Households per state",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(0, 'asc'), list(2, 'asc')),
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

### Household composition

We are interested in calculating the food consumption of different crops per person per year. The food consumption in the survey is at the household level. We therefore need to know the number of people per household so that the food quantities can be shared among the household and thereafter extrapolated to the population.

The composition of the household is contained within a 'roster' dataset. We remove household members in the roster who are no longer residing in the household (the GHS is a panel survey), and add any new members.

```{r}
#| label: GHS_roster_data
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ----1. post planting household size----

df_GHS_sect1_plantingw5 <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Planting Wave 5/Household/sect1_plantingw5.csv") |>
  mutate(
    household_member = s1q4 == "1. YES" |
      NEWMEMBER == 1
  )


# get size of each household

df_hh_size_pp <- df_GHS_sect1_plantingw5 |>
  filter(household_member) |>
  group_by(hhid) |>
  mutate(hh_size = n()) |>
  select(zone, state, hhid, hh_size) |>
  distinct()


# Get zone samples of hh_size

zone_sample_hh_size_pp <- df_hh_size_pp |>
  group_by(
    zone
  ) |>
  summarise(
    median_hh_size_pp = median(hh_size),
    max_hh_size_pp = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone)


# Get state samples of hh_size

state_sample_hh_size_pp <- df_hh_size_pp |>
  group_by(
    zone, state
  ) |>
  summarise(
    median_hh_size_pp = median(hh_size),
    max_hh_size_pp = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone, state)


# ----2. post harvest household size----

df_GHS_sect1_harvestw5 <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Harvest Wave 5/Household/sect1_harvestw5.csv") |>
  mutate(
    household_member = s1q4 == "1. YES" |
      NEWMEMBER == 1
  )

# get size of each household

df_hh_size_ph <- df_GHS_sect1_harvestw5 |>
  filter(household_member) |>
  group_by(hhid) |>
  mutate(hh_size = n()) |>
  select(zone, state, hhid, hh_size) |>
  distinct()


# Get zone samples of hh_size

zone_sample_hh_size_ph <- df_hh_size_ph |>
  group_by(
    zone
  ) |>
  summarise(
    median_hh_size_ph = median(hh_size),
    max_hh_size_ph = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone)


# Get state samples of hh_size

state_sample_hh_size_ph <- df_hh_size_ph |>
  group_by(
    zone, state
  ) |>
  summarise(
    median_hh_size_ph = median(hh_size),
    max_hh_size_ph = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone, state)

```

The average (median) household size per zone varies between 4 and 8 in the post-planting survey, with the three northern zones displaying higher averages than the three southern zones (@tbl-hh_size_zone). The zone with the largest individual household size was North West zone.

This pattern is mirrored in the post-harvest survey.

```{r}
#| label: tbl-hh_size_zone
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "Household size per zone"

zone_sample_hh_size <- zone_sample_hh_size_pp |> left_join(zone_sample_hh_size_ph, by = "zone")

zone_sample_hh_size |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "Median household size (post planting survey)", "Max household size (post planting survey)", "Median household size (post harvest survey)", "Max household size (post harvest survey)"),
    caption = "Household size per zone",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 1),
      # lengthMenu = c(15, 40),
      pageLength = 6
    ))

```

At the state level the (median) household size per zone varies between 3 in Cross River state and 9 in Borno state in the post-planting survey. The state that had the highest individual household was Kaduna (@tbl-hh_size_state).

In the post-harvest survey there were five southern states with a median of 3 households, and four northern states with a median of eight members per household.

```{r}
#| label: tbl-hh_size_state
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "Household size per state"

state_sample_hh_size <- state_sample_hh_size_pp |> left_join(select(state_sample_hh_size_ph, -zone), by = "state")

state_sample_hh_size |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "State", "Median household size (post planting survey)", "Max household size (post planting survey)", "Median household size (post harvest survey)", "Max household size (post harvest survey)"),
    caption = "Household size per state",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(0, 'asc'), list(2, 'desc')),
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 1),
      # lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

### Valid households for our analysis

Not all of the households in the surveys are valid for our analysis. Some households in the panel were not able to be contacted or the interview was incomplete.
Those households are removed to leave a consistent set of 4,715 households from the two surveys.

```{r}
#| label: GHS_analysis_households
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ----1. post planting survey ----

df_hh_analysis_pp <- df_GHS_secta_plantingw5 %>%
  filter(
    interview_result == "1. COMPLETE",
    !is.na(wt_cross_wave5),
    wt_cross_wave5 > 0
  ) %>%
  select(
    hhid,
    zone,
    state,
    lga,
    sector,
    ea,
    cluster,
    strata,
    wt_cross_wave5
  ) %>%
  left_join(
    select(df_hh_size_pp, -c(zone, state)),
    by = "hhid"
  ) %>%
  filter(
    !is.na(hh_size),
    hh_size > 0
  )

# ----2. post harvest survey ----

df_hh_analysis_ph <- df_GHS_secta_harvestw5 %>%
  filter(
    interview_result == "1. COMPLETE",
    !is.na(wt_cross_wave5),
    wt_cross_wave5 > 0
  ) %>%
  select(
    hhid,
    zone,
    state,
    lga,
    sector,
    ea,
    cluster,
    strata,
    wt_cross_wave5
  ) %>%
  left_join(
    select(df_hh_size_ph, -c(zone, state)),
    by = "hhid"
  ) %>%
  filter(
    !is.na(hh_size),
    hh_size > 0
  )

```


```{r}
#| label: GHS_sample_analysis_households
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ----1. post planting survey ----

zone_sample_analysis_pp <- df_hh_analysis_pp |>
  group_by(
    zone
  ) |>
  summarise(
    n_households_pp = n(),
    .groups = "drop"
  ) |>
  arrange(zone)

state_sample_analysis_pp <- df_hh_analysis_pp |>
  group_by(
    zone,
    state
  ) |>
  summarise(
    n_households_pp = n(),
    .groups = "drop"
  ) |> 
  arrange(zone, state)

# ----2. post harvest survey ----

zone_sample_analysis_ph <- df_hh_analysis_ph |>
  group_by(
    zone
  ) |>
  summarise(
    n_households_ph = n(),
    .groups = "drop"
  ) |>
  arrange(zone)

state_sample_analysis_ph <- df_hh_analysis_ph |>
  group_by(
    zone,
    state
  ) |>
  summarise(
    n_households_ph = n(),
    .groups = "drop"
  ) |>
  arrange(zone, state)
```

More households were dropped from the post-planting survey than from the post-harvest survey.

The pattern of the distribution of households among the zones is similar to the full set of households. North East zone has most households in both surveys, while South West has the fewest (@tbl-analysis_zone).

```{r}
#| label: tbl-analysis_zone
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "hh_analysis_zone"

zone_sample_analysis <- zone_sample_analysis_pp |> left_join(zone_sample_analysis_ph, by = "zone")

zone_sample_analysis |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "# of households (post planting survey)", "# of households (post harvest survey)"),
    caption = "Households per zone",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

At the state level the changes are greater, especially for the post planting survey, with states like Zamfara experiencing a reduced number of households for our analysis. However, the effect on the post-harvest number of households is less severe (@tbl-analysis_state), and the number of households per state is now almost the same when comparing the post planting and post harvest surveys.


```{r}
#| label: tbl-analysis_state
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "hh_analysis_state"

state_sample_analysis <- state_sample_analysis_pp |> left_join(select(state_sample_analysis_ph, -zone), by = "state")

state_sample_analysis |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "State", "# of households (post planting survey)", "# of households (post harvest survey)"),
    caption = "Households per state",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(0, 'asc'), list(2, 'asc')),
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))
```

```{r}
#| label: GHS_roster_data_analysis
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ----1. post planting household size----


# filter the roster using the valid hh_analysis hhid

df_hh_analysis_size_pp <- df_hh_size_pp |> right_join(select(df_hh_analysis_pp, hhid), by = "hhid")

# Get zone samples of hh_size

zone_sample_hh_analysis_size_pp <- df_hh_analysis_size_pp |>
  group_by(
    zone
  ) |>
  summarise(
    median_hh_size_pp = median(hh_size),
    max_hh_size_pp = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone)


# Get state samples of hh_size

state_sample_hh_analysis_size_pp <- df_hh_analysis_size_pp |>
  group_by(
    zone, state
  ) |>
  summarise(
    median_hh_size_pp = median(hh_size),
    max_hh_size_pp = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone, state)

# ----2. post harvest household size----
# filter the roster using the valid hh_analysis hhid

df_hh_analysis_size_ph <- df_hh_size_ph |> right_join(select(df_hh_analysis_ph, hhid), by = "hhid")


# Get zone samples of hh_size

zone_sample_hh_analysis_size_ph <- df_hh_analysis_size_ph |>
  group_by(
    zone
  ) |>
  summarise(
    median_hh_size_ph = median(hh_size),
    max_hh_size_ph = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone)


# Get state samples of hh_size

state_sample_hh_analysis_size_ph <- df_hh_analysis_size_ph |>
  group_by(
    zone, state
  ) |>
  summarise(
    median_hh_size_ph = median(hh_size),
    max_hh_size_ph = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone, state)

```

The average sizes of households in the analysis do not change, either for the zonal averages (@tbl-hh_analysis_size_zone) or the state averages (@tbl-hh_analysis_size_state).

```{r}
#| label: tbl-hh_analysis_size_zone
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "Households in analysis size per zone"

zone_sample_hh_analysis_size <- zone_sample_hh_analysis_size_pp |> left_join(zone_sample_hh_analysis_size_ph, by = "zone")

zone_sample_hh_analysis_size |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "Median household size (post planting survey)", "Max household size (post planting survey)", "Median household size (post harvest survey)", "Max household size (post harvest survey)"),
    caption = "Household size per zone",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 1),
      # lengthMenu = c(15, 40),
      pageLength = 6
    ))
```

```{r}
#| label: tbl-hh_analysis_size_state
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "Households in analysis size per state"

state_sample_hh_analysis_size <- state_sample_hh_analysis_size_pp |> left_join(select(state_sample_hh_analysis_size_ph, -zone), by = "state")

state_sample_hh_analysis_size |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "State", "Median household size (post planting survey)", "Max household size (post planting survey)", "Median household size (post harvest survey)", "Max household size (post harvest survey)"),
    caption = "Household size per state",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(0, 'asc'), list(2, 'desc')),
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 1),
      # lengthMenu = c(15, 40),
      pageLength = 15
    ))
```

### Extrapolating from the sample to the population

Since our objective is a representative estimate of food consumption in 2023/24 (rather than a longitudinal comparison) we should use the cross-sectional weight `wt_cross_wave5`. NBS explicitly says the cross-sectional weight is calibrated to the 2023 population and is intended for representative current-population estimates.

```{r}
#| label: GHS_analysis_weights
#| message: false
#| warning: false
#| cache: false
#| results: asis

# Check

summary(df_hh_analysis_pp$wt_cross_wave5)
summary(df_hh_analysis_ph$wt_cross_wave5)

# Get zone samples of hh_size

zone_sample_wt_cross_pp <- df_hh_analysis_pp |>
  group_by(
    zone
  ) |>
  summarise(
    mean_wt_cross_pp = mean(wt_cross_wave5),
    .groups = "drop"
  ) |>
  arrange(zone)

```

We will use the cross sectional weight in further calculations.

### Household Food consumption

This section records household-food item level information on the quantity and value of food consumed within the household during the past 7 days.

The data unit for these data are food products per household. The response for each food product is either "1. Yes" or "2. No".

There are over 100 different food products that can be recorded, and as a consequence of our samples of over 4,000 households, there are over half a million records for both the post-planting and post-harvest datasets.

To calculate household food consumption the relevant variables are:

Variable: Post Planting Survey | Variable: Post Harvest Survey | Description
------------------------------|------------------------------|--------------
`s6bq1` | `s5bq1`  | consumed this item in previous 7 days?
`s6bq2a` | `s5bq2a` | quantity consumed
`s6bq2b` | `s5bq2b` | unit
`s6bq2b_os` | `s5bq2b_os` | other unit
`s6bq2c` | `s5bq2c` | size
`s6bq2_cvn` | `s5bq2_cvn` | conversion factor

NBS describes `s6bq2_cvn` and `s5bq2_cvn` specifically as:

conversion factor item-unit-size-zone

and it ranges from 0.001 to 50 in the released data.

So the basic calculation is:

quantity_kg_7days =    `s6bq2a`  * `s6bq2_cvn`



```{r}
#| label: GHS_food_consumed_hh_analysis
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ----1. post planting household food consumption----

df_GHS_sect6b_plantingw5 <- read.csv(
  "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Planting Wave 5/Household/sect6b_plantingw5.csv")

df_food_pp <- df_GHS_sect6b_plantingw5 |>
  mutate(
    s6bq2a = as.numeric(s6bq2a),
    s6bq2_cvn = as.numeric(s6bq2_cvn),
    item_cd_int = as.numeric(stringi::stri_extract_first_regex(
      item_cd,
      "[0-9]+"
    ))
  ) 

df_food_pp_consumed <- df_food_pp |>
  filter(
    s6bq1 == "1. Yes",
    !is.na(s6bq2a),
    s6bq2a >= 0
  ) %>%
  mutate(
    kg_week = s6bq2a * s6bq2_cvn) %>%
  filter(!is.na(kg_week))

df_food_item_pp_consumed <- df_food_pp_consumed |>
  group_by(
    item_cd_int
  ) |>
  summarise(
    mean_kg = mean(kg_week),
    .groups = "drop"
  ) |>
  arrange(item_cd_int)


# ----2. post harvest household food consumption----

df_GHS_sect5b_plantingw5 <- read.csv(
  "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Harvest Wave 5/Household/sect5b_harvestw5.csv")

df_food_ph <- df_GHS_sect5b_plantingw5 |>
  mutate(
    s5bq2a = as.numeric(s5bq2a),
    s5bq2_cvn = as.numeric(s5bq2_cvn),
    item_cd_int = as.numeric(stringi::stri_extract_first_regex(
      item_cd,
      "[0-9]+"
    ))
  ) 

df_food_ph_consumed <- df_food_ph |>
  filter(
    s5bq1 == "1. Yes",
    !is.na(s5bq2a),
    s5bq2a >= 0
  ) %>%
  mutate(
    kg_week = s5bq2a * s5bq2_cvn,
  ) %>%
  filter(!is.na(kg_week))

df_food_item_ph_consumed <- df_food_ph_consumed |>
  group_by(
    item_cd_int
  ) |>
  summarise(
    mean_kg = mean(kg_week),
    .groups = "drop"
  ) |>
  arrange(item_cd_int)
```

```{r}
#| label: GHS_food_product_lookup
#| message: false
#| warning: false
#| cache: false
#| results: asis


food_lookup <- tribble(
  ~item_cd_int , ~product                     ,

        10 , "Guinea corn / sorghum"      ,
        11 , "Millet"                     ,
        13 , "Rice - local"               ,
        14 , "Rice - imported"            ,
        16 , "Maize flour"                ,
        17 , "Yam flour"                  ,
        18 , "Cassava flour"              ,
        19 , "Wheat flour"                ,
        20 , "Maize - unshelled/on cob"   ,
        22 , "Maize - shelled/off cob"    ,
        23 , "Other grains/flour"         ,

        25 , "Bread"                      ,
        26 , "Cake"                       ,
        27 , "Buns/Pofpof/Donuts"         ,
        28 , "Biscuits"                   ,
        29 , "Meat pie/Sausage roll"      ,

        30 , "Cassava roots"              ,
        31 , "Yam roots"                  ,
        32 , "Gari - white"               ,
        33 , "Gari - yellow"              ,
        34 , "Cocoyam"                    ,
        35 , "Plantains"                  ,
        36 , "Sweet potatoes"             ,
        37 , "Potatoes"                   ,
        38 , "Other roots/tubers"         ,

        40 , "Soya beans"                 ,
        41 , "Brown beans"                ,
        42 , "White beans"                ,
        43 , "Groundnuts - unshelled"     ,
        44 , "Groundnuts - shelled"       ,
        45 , "Other nuts/seeds/pulses"    ,
        46 , "Coconut"                    ,
        47 , "Kola nut"                   ,
        48 , "Cashew nut"                 ,

        50 , "Palm oil"                   ,
        51 , "Butter/Margarine"           ,
        52 , "Groundnut oil"              ,
        53 , "Other oil/fat"              ,
        54 , "Shea butter"                ,
        56 , "Animal fat"                 ,

        60 , "Bananas"                    ,
        61 , "Orange/Tangerine"           ,
        62 , "Mangoes"                    ,
        63 , "Avocado pear"               ,
        64 , "Pineapples"                 ,
        66 , "Other fruits"               ,
        67 , "Pawpaw"                     ,
        68 , "Watermelon"                 ,
        69 , "Apples"                     ,

        70 , "Tomatoes"                   ,
        71 , "Tomato puree - canned"      ,
        72 , "Onions"                     ,
        73 , "Garden eggs/Eggplant"       ,
        74 , "Okra - fresh"               ,
        75 , "Okra - dried"               ,
        76 , "Fresh pepper"               ,
        77 , "Dry pepper"                 ,
        78 , "Leaves"                     ,
        79 , "Other vegetables"           ,

        80 , "Chicken"                    ,
        81 , "Duck"                       ,
        82 , "Other domestic poultry"     ,
        83 , "Agricultural eggs"          ,
        84 , "Local eggs"                 ,
        85 , "Other eggs"                 ,

        90 , "Beef"                       ,
        91 , "Mutton"                     ,
        92 , "Pork"                       ,
        93 , "Goat"                       ,
        94 , "Wild game/Bush meat"        ,
        95 , "Canned beef"                ,
        96 , "Other meat"                 ,

       100 , "Fish - fresh"               ,
       101 , "Fish - frozen"              ,
       102 , "Fish - smoked"              ,
       103 , "Fish - dried"               ,
       104 , "Snails"                     ,
       105 , "Seafood"                    ,
       106 , "Canned fish/seafood"        ,
       107 , "Other fish/seafood"         ,

       110 , "Fresh milk"                 ,
       111 , "Milk powder"                ,
       112 , "Baby milk powder"           ,
       113 , "Milk - tinned"              ,
       114 , "Cheese/wara"                ,
       115 , "Other milk products"        ,

       120 , "Coffee"                     ,
       121 , "Chocolate drinks"           ,
       122 , "Tea"                        ,

       130 , "Sugar"                      ,
       132 , "Honey"                      ,
       133 , "Other sweets"               ,

       141 , "Salt"                       ,
       142 , "Ogbono - unground"          ,
       143 , "Ogbono - ground"            ,
       144 , "Ground pepper"              ,
       145 , "Melon - shelled"            ,
       146 , "Melon - unshelled"          ,
       147 , "Melon - ground"             ,
       148 , "Other spices"               ,

       150 , "Bottled water"              ,
       151 , "Sachet water"               ,
       152 , "Malt drinks"                ,
       153 , "Soft drinks"                ,
       154 , "Fruit juice"                ,
       155 , "Other non-alcoholic drinks" ,

       160 , "Beer"                       ,
       161 , "Palm wine"                  ,
       162 , "Pito"                       ,
       163 , "Gin"                        ,
       164 , "Other alcoholic beverages"  ,

       601 , "Guava"
)


```


```{r}
#| label: tbl-food_kg_week
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "food_kg_week"


df_food_item_consumed <- df_food_item_pp_consumed |>
  merge(df_food_item_ph_consumed, by = "item_cd_int", all = T) |>
left_join(food_lookup, by = "item_cd_int")  |>
  mutate(
    kg_week_pp = prettyunits::pretty_round(mean_kg.x, digits = 2),
    kg_week_ph = prettyunits::pretty_round(mean_kg.y, digits = 2)
  )|>
  select(item_cd_int, product, kg_week_pp, kg_week_ph)

df_food_item_consumed |> 
  DT::datatable(
    rownames = F,
    colnames=c("ID", "Food Product", "Mean kg per household per week (post planting survey)", "Mean kg per household per week (post harvest survey)"),
    caption = "Weekly Food Product Consumption",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(2, 'desc'), list(3, 'asc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```



```{r}
# missing_conversion_pp <- food_pp_consumed %>%
#   filter(
#     is.na(s6bq2_cvn)
#   ) %>%
#   count(
#     item_cd,
#     item_cd_alt,
#     item_os,
#     sort = TRUE
#   )
# 
# missing_conversion_pp


```




## Per capita consumption of specific food products and crop equivalents

The global Biofortification Priority Index (BPI) [@asare-marfoPrioritizingCountriesBiofortifcation2013] uses an indicator of
consumption per capita per year for each specific crop, standardised to fit with the maximum and minimum values encountered.

These consumption values are not readily available in the reports of the 2021 NFCMS or the 2023/24 General Household Survey–Panel. An advantage of the 2023/24 General Household Survey–Panel (GHS-Panel Wave 5) is that it has a household food-consumption module publicly available from the Nigeria National Bureau of Statistics.

The 2023/24 GHS contains a Post-Harvest Section 5B – Food Consumption with, for each household and food item:

 - quantity consumed during the past 7 days
 - unit of measurement
 - conversion factor to standard quantity
 - quantity obtained from:
 + purchases
 + own production
 + gifts/other sources
 - quantity purchased during the past 30 days
 - expenditure on the food

The dataset has 509,220 food-item/household observations and 26 variables.

The microdata have individual `item_cd` food items, so we can investigate the actual item list and identify things such as maize, maize flour, maize meal, gari, cassava flour, fufu, etc. The conversion factor is particularly valuable because households report quantities in local/non-standard units and the survey provides an item-unit-size-zone conversion factor.

We can also calculate the crop-equivalent consumption:

::: {.callout-note title="Calculating the crop-equivalent consumption"}
For example, suppose the household records:

Gari → 10 kg consumed

We could have two possible measures:

Food-product consumption

Gari = 10 kg/person/year

or, with the appropriate processing/yield conversion:

Crop-equivalent consumption

Cassava equivalent = X kg/person/year

That would allow you to aggregate all cassava-derived products:

gari + fufu + cassava flour + other cassava foods
→ total cassava-equivalent consumption kg/person/year

Likewise:

maize grain + maize flour + maize meal + maize products
→ total maize-equivalent consumption
:::


1. The dataset we want

For the post-harvest 2023/24 visit, the key file is:

`sect5b_harvestw5.dta` — Section 5B: Food Consumption

It contains 509,220 household-food observations and 26 variables. Each household has a record for the food items in the food roster, whether or not that particular item was consumed.

The corresponding post-planting file is:

`sect6b_plantingw5.dta`

with 529,470 observations, and essentially the same food-consumption structure.

The survey is representative of the six geopolitical zones, so zonal estimates are an intended use of the data.

Importantly, the GHS report says that when information was collected at both visits, the published results normally use the most recent information, i.e. the post-harvest visit.

So for our purpose we use `sect5b_harvestw5.dta` as the *primary* dataset.

2. What food items are actually available?

The standard `item_cd` roster contains the following crop-related items, among others:

Code |	Food item
-----|-----------
10	| Guinea corn/sorghum
11	| 	Millet
13	| 	Rice – local
14	| 	Rice – imported
16	| 	Maize flour
17	| 	Yam flour
18	| 	Cassava flour
20	| 	Maize – unshelled/on the cob
22	| 	Maize – shelled/off the cob
23	| 	Other grains and flour
30	| 	Cassava – roots
31	| 	Yam – roots
32	| 	Gari – white
33	| 	Gari – yellow
34	| 	Cocoyam
35	| 	Plantains
36	| 	Sweet potatoes
37	| 	Potatoes
38	| 	Other roots and tubers
40	| 	Soya beans
41	| 	Brown beans
42	| 	White beans
43	| 	Groundnuts – unshelled
44	| 	Groundnuts – shelled
45	| 	Other nuts/seeds/pulses
46	| 	Coconut
47	| 	Kola nut
48	| 	Cashew nut
50	| 	Palm oil
52	| 	Groundnut oil
60	| 	Bananas
61	| 	Orange/tangerine
62	| 	Mangoes
63	| 	Avocado pear
64	| 	Pineapples
67	| 	Pawpaw
68	| 	Watermelon
70	| 	Tomatoes
72	| 	Onions
73	| 	Garden eggs/eggplant
74	| 	Okra – fresh
75	| 	Okra – dried
76	| 	Fresh pepper
77	| 	Dry pepper
78	| 	Leaves
79	| 	Other vegetables

There are also numerous non-crop foods. The complete catalogue runs from item code 10 through 601, with gaps between food groups.

But there is an important second level

The variable `item_os_cd` captures additional foods when the respondent selected "other item".

This is particularly interesting for our purposes because it contains:

Code |	Food item
-----|-----------
231	|  Semovita/semolina
232	|  Akanu/Pap/Obi/Fufu
234	|  Plantain flour
235	|  Macaroni/pasta/noodles
237	|  Acha/fonio
381	|  Yam – second variety
453	|  Bambara nut
454	|  Palm nut
455	|  Palm kernel
456	|  Locust beans
661	|  Agbalumo/Udara/star apple
662	|  Cashew fruit
790	|  Jute leaves
793	|  Kuka/Baobab
798	|  Pumpkin leaves
1071	|  Crayfish
1152	|  Yoghurt
1331	|  Sugar cane
1551	|  Zobo
1552	|  Kunu
etc.

There were 503 valid `item_os_cd` observations in the post-harvest file.

So fufu isn't missing entirely — it occurs as an "other item", although only two observations in the post-harvest file use the coded category 232. There are also free-text unit descriptions mentioning cooked fufu.

This means we need to retain both `item_cd` and `item_os_cd`/`item_os` when constructing the food list.

3. How the quantity information works

For every household × food item, Section 5B contains:

Variable |	Meaning
---------|-----------
`s5bq1`	 |	Did the household consume this item in the past 7 days?
`s5bq2a` 	 |	Quantity consumed in past 7 days
`s5bq2b`	 |	Unit
`s5bq2b_os`	 |	Other unit, if applicable
`s5bq2c`	 |	Size
`s5bq2_cvn`	 |	Conversion factor: item × unit × size × zone
`s5bq3`	 |	Quantity consumed from purchases
`s5bq4`	 |	Quantity consumed from own production
`s5bq5`	 |	Quantity consumed from gifts/other sources

There is a parallel set for purchases during the previous 30 days:

`s5bq7a`
`s5bq7b`
`s5bq7c`
`s5bq7_cvn`
`s5bq8`



::: {.callout-note title="Calculating the weight of the food product consumed"}
The catalogue explicitly describes the conversion factor as "conversion factor item-unit-size-zone".

Suppose a household reports:

2 baskets of maize

The raw `s5bq2a` value is 2, but that is not 2 kg.

The survey records:

- quantity = 2
- unit = basket
- size = [reported size]
- conversion factor = X kg per reported unit

so we can derive:

quantity consumed in kg = `s5bq2a` × `s5bq2_cvn`

subject to checking the exact direction/definition of the conversion factor in the questionnaire/data documentation.

The last check should be conducted before producing the final numbers rather than assuming the factor's direction.
:::

4. We should use consumption, NOT purchases

This distinction is very important for our question.

We want:

average consumption per person per year

Therefore the appropriate variable is:

`s5bq2a` × `s5bq2_cvn`

—not the 30-day purchase variable.

The purchase variables tell us where the food came from / how much was bought.

The survey explicitly separates these sources with `s5bq3`, `s5bq4`, and `s5bq5`.

For total consumption we therefore want:

purchased consumption + own-production consumption + gifts/other consumption

which should correspond to the total reported in `s5bq2a`.

5. Converting the seven-day observation to kg/year

For each household and food item:

$$ Q_{7d,kg} = `s5bq2a` \times `s5bq2\_cvn` $$

Then:

$$ Q_{annual,kg} = Q_{7d,kg}\times\frac{365}{7} $$

But we should not calculate the annual value separately for every household and then simply take an unweighted mean, for that we need the survey weights.

6. Calculate the geopolitical-zone per-capita figure

There is an important statistical subtlety here.

Rather than averaging household per-capita values, we should calculate the zonal estimate as:

$$ \boxed{ \frac{ \sum_i w_i Q_i }{ \sum_i w_i HHSize_i } } $$

where:

\(Q_i\) = annual kg consumed by household \(i\)
\(HHSize_i\) = number of people in household \(i\)
\(w_i\) = household survey weight.

In other words:

estimated total kg consumed by the zone ÷ estimated population represented by the zone

This is conceptually the cleanest definition of kg/person/year.

It also avoids giving a five-person household and a one-person household equal influence simply because they are both households.


8. There is one major issue we need to resolve before calculating the numbers

The GHS food-consumption data are weekly household consumption observations, not a continuous annual food diary.

The post-harvest survey records:

"quantity ... consumed in the past 7 days"

and the post-planting survey does the same.

And the two visits produce noticeably different results.

For example, the report shows that the share of households consuming particular food groups changes between post-planting and post-harvest.

Therefore:

We cannot automatically call

7-day consumption × 52.14

"annual consumption".

It is better described as:

annualised consumption based on the post-harvest seven-day recall

unless we decide to combine the two seasonal observations.

This matters particularly for crops such as:

maize
cassava
yam
plantain
mango
etc.

because consumption is seasonal.

9. For our application we should actually calculate two versions

This would be much more informative.

A. Post-harvest annualised consumption
$$ Q_{PH,annual}=Q_{PH,7d}\times365/7 $$

This corresponds to the standard Wave 5 food-consumption observation used by NBS for its most recent visit.

B. Seasonal-average annualised consumption

If we use both visits:

$$ Q_{annual} = \frac{ Q_{PP,7d}+Q_{PH,7d} }{2} \times\frac{365}{7} $$

This is effectively an estimate based on two seasonal snapshots.

We should not present B as an official NBS annual consumption statistic, but it could be a much better estimate for our crop-consumption application.

The NBS report explicitly recognises the seasonal difference between post-planting and post-harvest consumption.

11. Final Zonal Dataset

I think the ideal final dataset would actually have three levels:

Zone |	Crop |	Product |	kg/person/year
---- |-------|----------|---------------
North Central |	Maize |	Maize flour |	...
North Central |	Maize |	Shelled maize |		...
North Central |	Maize |	Unshelled maize |		...
North Central |	Cassava |	Cassava roots |		...
North Central |	Cassava |	Cassava flour |		...
North Central |	Cassava |	Gari – white |		...
North Central |	Cassava |	Gari – yellow |		...
South East |	Cassava |	Cassava roots |		...
South East |	Cassava |	Gari – white |		...
...	...	...	...

and then a second table:

Zone |	Crop |	Crop-equivalent kg/person/year
---- |-------|-------------------------
North Central |	Maize |		...
North Central |	Cassava |		...
North East |	Maize |		...
...	...	...

The second table would require us to establish defensible processing conversion factors for gari, flour, etc.

One particularly useful discovery

The "other food" mechanism means we should not limit ourselves to `item_cd`.

The post-harvest data contain 503 `item_os_cd` records, including fufu, plantain flour, Bambara nut, palm kernel, locust beans, Kunu, Zobo and many other foods.

So when we actually process the microdata, I would:

 - retain the standard `item_cd`;
 - retain `item_os_cd`;
 - retain the free-text `item_os`;
 - construct a master food-code lookup;
 - identify crop/product relationships;
 - calculate consumption in kg;
 - apply household weights;
 - calculate the six-zone estimates;
 - separately identify observations that require manual classification.

That last step is important because otherwise we could silently lose things like fufu.

What I would do next

The remaining critical piece is the actual Wave 5 household weight and household-size variables. Once those are identified, we can write an R workflow that takes the NBS .dta files and produces the six-zone consumption estimates.

I would also like to inspect the NBS questionnaire/data documentation for the precise definition of `s5bq2_cvn`, rather than making an assumption about whether it is kg per reported unit or its reciprocal.

Then we can produce, for example:

Maize flour — North Central — X kg/person/year
Maize flour — North East — X kg/person/year
...
Gari white — South East — X kg/person/year
...

with the calculation fully reproducible in R.


The next step is to pin down the exact variables and conversion-factor definition, then we can build the reproducible R workflow for the six geopolitical zones.

I’ll treat the target output as:

food product → crop → geopolitical zone → kg/person/year, with the option to aggregate processed products back to crop-equivalent kg/person/year later.

I’ve now pinned down the key pieces. There is one correction to my previous answer: we should use the cross-sectional weight `wt_cross_wave5`, not the longitudinal weight, because your objective is a representative estimate of food consumption in 2023/24 rather than a longitudinal comparison. NBS explicitly says the cross-sectional weight is calibrated to the 2023 population and is intended for representative current-population estimates.

1. Files we need

For the primary calculation we need three files:

`sect5b_harvestw5.dta` — food consumption
`secta_harvestw5.dta` — household ID, zone and wt_cross_wave5
`sect1_harvestw5.dta` — household roster, from which we calculate household size

The food file contains 509,220 household × food-item observations and 26 variables.

The cover file contains 4,771 households and the three Wave 5 weight variables, including `wt_cross_wave5`.

The roster contains 28,256 individual records across the households.

2. The consumption calculation

The relevant variables are:

`s5bq1`       consumed this item in previous 7 days?
`s5bq2a`      quantity consumed
`s5bq2b`      unit
`s5bq2b_os`   other unit
`s5bq2c`      size
`s5bq2_cvn`   conversion factor

NBS describes `s5bq2_cvn` specifically as:

conversion factor item-unit-size-zone

and it ranges from 0.001 to 50 in the released data.

So the basic calculation is:

quantity_kg_7days =
    s5bq2a * s5bq2_cvn
    
But I would test this against the actual data before finalising it, because the catalogue gives us the definition and range but doesn't explicitly state the direction of the conversion factor in the metadata. We don't want to assume that 0.5, for example, means 0.5 kg/unit rather than the reverse.

3. Household size

Rather than relying on a pre-calculated household-size variable, I recommend calculating it directly from `sect1_harvestw5.dta`.

That gives us:
hh_size =
    number of household members in the Wave 5 post-harvest roster

This is preferable because we can explicitly inspect which roster records represent current household members.

. The weight

We'll join:

`sect5b_harvestw5`
        |
        | `hhid`
        v
`secta_harvestw5`
        |
        +---- `zone`
        +---- `wt_cross_wave5`

The `wt_cross_wave5` weight is the appropriate one because NBS says it was calibrated against official 2023 population projections at state level and is specifically designed for cross-sectional estimates of the 2023/24 population.

5. The resulting estimator

For a food item \(j\) in geopolitical zone \(z\), I would calculate:

$$ Q_{j,z} = \frac{ \sum_i w_i Q_{ij} }{ \sum_i w_i H_i } $$

where:

\(Q_{ij}\) = annual kg of food \(j\) consumed by household \(i\)
\(H_i\) = household size
\(w_i\) = wt_cross_wave5

This is effectively:

estimated annual kilograms consumed by the population / estimated population

which is exactly the interpretation we want for kg/person/year.

6. Annualisation

The survey asks about consumption during the previous seven days.

So for the post-harvest observation:

kg_year <- kg_7days * 365 / 7

However, I suggest that we initially report this explicitly as:

annualised post-harvest consumption

rather than implying that NBS directly measured annual consumption.

This is important because Wave 5 has two separate food-consumption observations:

post-planting: July–September 2023
post-harvest: January–March 2024.

The catalogue confirms that the post-planting food-consumption file has the same basic structure, using `s6bq2a` and `s6bq2_cvn`.

7. I think we should actually retain both seasons

For your eventual crop-consumption analysis, I would create:

Dataset A — post-harvest
`zone`
`item_cd`
`item_name`
`kg_person_year`

based on:

January–March 2024
7-day recall
annualised × 365/7

Dataset B — post-planting
`zone`
`item_cd`
`item_name`
`kg_person_year`

based on:

July–September 2023
7-day recall
annualised × 365/7

Dataset C — seasonal average

Then:

annualised =
    ((post_planting_7day + post_harvest_7day) / 2) * 365 / 7
    
I'd regard Dataset C as an analytical estimate, not an official NBS annual statistic.

For crops with strong seasonality, this will probably be much more useful than relying exclusively on the January–March observation.

8. An important additional opportunity

There is also a Section 5C – Aggregate Food Consumption file:

`sect5c_harvestw5.dta` 

It has 56,580 observations and records the number of days the household consumed each food-group item during the previous seven days.

We should not use that for the kg calculation, but it is potentially useful as a quality check.

For example, if our calculation says a household consumed 50 kg of maize flour in seven days, we can check whether that household reported consuming maize flour on some/all of the seven days. It could help identify implausible conversion-factor results.

::: {.callout-note title="Testing the conversion factor"}
The conversion factor works exactly as we hoped

For example, one household reported:

*"25 cigarette cups of local rice"*

and the conversion factor is:

0.232

giving:

25 × 0.232 = 5.8 kg

That is clearly the intended interpretation.

Another household reported:

*"2 large loaves of bread"*

with conversion factor 0.780, giving:

2 × 0.780 = 1.56 kg

So we can safely calculate:

kg_7days = `s5bq2a` * `s5bq2_cvn`

The conversion factor is therefore effectively kg per reported quantity/unit/size/zone combination.
::: 


2. The food item coding is more interesting than expected

`item_cd` is actually a labelled text field in our CSV, e.g.:

16. Maize flour
18. Cassava  flour
30. Cassava - roots
32. Gari - white
33. Gari - yellow

There is also `item_cd_alt`, which is extremely important.

For example, the standard category:
23. Other grains and flour

can have:

231. Semovita/semolina
232. Akanu/Pap/Obi/Fufu
234. Plantain flour
235. Macaroni/pasta/noodles
238. INDOMIE noodles

Similarly:

107. Other fish or seafood
can contain:
1071. Crayfish
1072. Periwinkle
1073. Stockfish

So we absolutely need to retain `item_cd_alt` when building the food dictionary.

3. There is a complication with cassava. Some respondents coded their consumption as:

30. Cassava - roots

but their free-text description says things such as:

WRAP (FUFU)

or

WRAP COOKEDFUFU

This means the NBS coding isn't perfectly clean at the product level.

For example, some households have:

`item_cd` = 30. Cassava - roots
`unit` = WRAP
`description` = WRAP FUFU

So we shouldn't blindly interpret every kilogram assigned to item 30 as raw cassava root.

This is actually a strong argument for retaining the original variables and creating our own carefully documented crop and product classification.

4. Missing conversion factors are rare—but need treatment

There are 80,702 "Yes" consumption observations.

Of those, only 210 have a missing `s5bq2_cvn`.

For the crops we're interested in, the situation is generally excellent.

For example:

Food | Consumed |	Missing conversion
-----|----------|--------------------
Maize flour |		1,106 |		0
Yam flour |		425	 |	0
Cassava flour |		886	 |	1
Maize, unshelled |		26 |		0
Maize, shelled |		444 |		0
Cassava roots |		530 |	38
Yam roots |	2,138 |		0
Gari white |		1,926 |		0
Gari yellow |		1,011 |		1
Plantains |		1,080 |		0

So for maize, yam, gari, etc., we're in very good shape.

Cassava roots require some additional cleaning because of those 38 observations, many involving locally described units such as wrap or satana.

5. Household size can also be calculated reliably

The household roster contains 28,256 records.

We found that we cannot simply count rows.

For existing household members, `s1q4` = "1. YES" identifies membership, but new household members have `NEWMEMBER` = 1 and `s1q4` is blank.

So I would define household size as:

member =
    `s1q4` == "1. YES" |
    `NEWMEMBER` == 1
    
This gives 4,716 households with a calculable household size.

The average household size in these records is about 5.44 people, with a range of 1–25.

That's a useful validation check.

6. The weight is exactly where expected

`secta_harvestw5.csv` contains:

`wt_wave5`
`wt_longpanel_wave5`
`wt_cross_wave5`

For our purpose the most appropriate is `wt_cross_wave5`, because we're estimating population consumption in 2023/24, rather than analysing changes among the longitudinal panel.

7. First calculation

Using:

- post-harvest consumption
- 7-day recall
- the NBS conversion factor
- household size from the roster
- wt_cross_wave5
- population-weighted aggregation
- annualisation by 365/7

We get the following preliminary figures.

kg/person/year

Product	| North Central	| North East| North West| South East| South South	| South West
--------|---------------|-----------|-----------|-----------|-------------|-----------
Maize flour	      | 	 31.78	| 	…		| …		| …		| …	| 	14.85
Maize – shelled		|    16.86	| 	…		| …		| …		| …	| 	15.92
Maize – unshelled	| 	 14.43	| 	…	  | …		| …	  | …	|   10.72
Cassava flour	    | 	 17.41	| 	…		| …		| …		| …	|   17.74
Cassava roots*	  | 	204.33	| 	…	  | …		| …		| …	|  375.90
Gari – white	    | 	 12.40	| 	…	  | 	…	| 	…	| …	|   19.08
Gari – yellow	    | 	 12.78	| 	…		|  …	| 	…	| …	|   22.98
Yam flour	        | 	 23.31	| 	…	  | 	…	| …		| …	|   16.51
Yam roots	        | 	134.97	| 	…	  | …		| …		| …	|   86.09
Plantains	        | 	 15.05	| 	…	  | 	…	| …		| …	|   18.63

* Do not use the cassava-root figures yet. The fufu/wrap coding issue means these need additional investigation.

These numbers are not yet the final dataset; They're shown as a check that the calculation is working, and the numbers are reasonable.

The magnitude of the yam and cassava numbers is plausible, but we need to investigate the coding before interpreting them as crop consumption.

8. Intermediate food dictionary

Rather than immediately producing a simple:

crop | zone | kg/person/year
-----|------|---------------

dataset, we should construct an intermediate food dictionary.

Something like:

GHS code	| 	GHS description	| 	Crop 	Product	| 	Include?	| 	Notes
----------|-------------------|-----------------|-------------|----------
16	      | 	Maize flour		    | Maize	Flour	  | 	Yes		| 
20	      | 	Maize unshelled		| Maize	Cob	    | 	Yes		| conversion needed for crop equivalent
22	      | 	Maize shelled	    | Maize	Grain	  | 	Yes		| 
30	      | 	Cassava roots	    | Cassava	Roots	|  Yes*		| inspect fufu descriptions
18	      | 	Cassava flour	    | Cassava	Flour	| 	Yes		| 
32	      | 	Gari white	      | Cassava	Gari	| 	Yes		| 
33	      | 	Gari yellow	      | Cassava	Gari	| 	Yes		| 
232	      | 	Akanu/Pap/Obi/Fufu| ?	Fufu	      | 	Yes  	| investigate composition
234	      | 	Plantain flour	  | Plantain Flour|   Yes		| 
31	      | 	Yam roots		      | Yam	Roots	    | 	Yes		| 
17	      | 	Yam flour	        | Yam	Flour   	| 	Yes		| 

This gives us a transparent classification that you can subsequently modify.

9. And I think we should make the final dataset slightly richer

I'd suggest the final output contain:

- `zone`
- `food_code`
- `food_product`
- `crop`
- `product_type`
- `households_consuming`
- `kg_person_7day`
- `kg_person_year`

and perhaps:

- `conversion_missing`

and a flag for problematic/custom food descriptions.

Then we can create a second aggregated table:

- `zone`
- `crop`
- `kg_person_year`

but only for products that genuinely can be aggregated without a processing conversion.

For crop-equivalent consumption, we'll subsequently need a separate processing-conversion table.

Since we now have the actual Wave 5 files, I would use the following as the first complete production script. It deliberately keeps the food-product calculation separate from any later crop-equivalent conversion.

The script:

- reads the three files;
- constructs household size from the roster;
- uses wt_cross_wave5;
- uses the 7-day consumption quantity, not purchases;
- converts reported quantities to kg using s5bq2_cvn;
- annualises the 7-day quantity;
- calculates population-weighted kg/person/year for each food product and geopolitical zone;
- retains the item_cd_alt and item_os information;
- creates a crop/product lookup;
- flags unusual observations rather than silently discarding them;
- writes CSV files that you can inspect in Excel.

### State-level estimates

State-level estimates are possible, technically calculable, but not entirely statistically reliable.

The Wave 5 sample covers all 36 states plus FCT, and the survey documentation says the original GHS design allocated PSUs across all 37 state-level areas. The NBS catalogue confirms that state is available in the Wave 5 data with codes 1–37.

The important caveat

The number of sampled households varies considerably by state. For example, in one Wave 5 household-level file:

- Abia: 2,997 observations
- Bauchi: 2,553
- Kaduna: 2,442
- Lagos: 2,553
- Nasarawa: only 222
- Kebbi: 777
- Sokoto: 666
- Zamfara: 666

The NBS catalogue shows all 37 areas and these substantial differences in sample size.

So we should not simply calculate a state mean and assume that all 37 estimates are equally precise. But the survey weights make state estimates legitimate. The GHS is designed to produce estimates at the state level, and the household data contain the state identifier and survey weights. Therefore, for a state \(s\), we can calculate:

$$ \text{kg/person/year}_{s} = \frac{ \sum_i w_i Q_i }{ \sum_i w_i H_i } $$

where:

\(Q_i\) = annualised kg consumed by household \(i\)
\(H_i\) = household size
\(w_i\) = Wave 5 cross-sectional household weight.

So the same estimator we developed for the geopolitical zones can be applied to each of the 37 state-level areas.

I would, however, add precision diagnostics

For each state × food product, the output should include at least:

State	| Product	| kg/person/year	| Sample households	| Consuming households
------|---------|-----------------|-------------------|-------------------- 
Nasarawa	| Maize flour	| X	 	|  ...	| 	...
Nasarawa	| Gari white	| X		| ...	| 	...
Kaduna	| Maize flour	| X		| ...	| 	...

And ideally:

- standard error
- coefficient of variation (CV)
- 95% confidence interval
- unweighted number of households contributing to the estimate.

That would allow us to flag estimates such as:

Nasarawa — a particular food — 8.4 kg/person/year (CV 38%)

as substantially less reliable than:

Kaduna — same food — 9.1 kg/person/year (CV 8%).

This is particularly important for relatively uncommon foods.

There is another reason to be cautious. The food-consumption file contains 509,220 household × food observations, but those are repeated food records within only about 4,700 households. We cannot treat the 509,220 observations as independent observations when calculating uncertainty.

For proper standard errors, we should account for the complex survey design — PSU, strata and weights.

So the final R workflow should use the 'survey' package, rather than just 'dplyr', for the state-level estimates and confidence intervals.

Instead of producing only:

- `geopolitical_zone`
- `food_product`
- `kg_person_year`

The main output should show:

- `geopolitical_zone`
- `state`
- `food_code`
- `food_product`
- `crop`
- `product_type`
- `n_households`
- `n_consuming`
- `kg_person_year`
- `se`
- `cv`
- `ci_lower`
- `ci_upper`

Then we can aggregate the state estimates to the six zones.
That gives much more flexibility: state-level results for our analysis, while retaining the geopolitical-zone results.

And because we have your actual Wave 5 files, I can revise the script to use the actual PSU/stratum variables in these files and calculate proper survey-weighted standard errors and confidence intervals, rather than just producing point estimates. That is the version I would recommend using for publication-quality state-level consumption estimates.

I have also checked the structure of your Wave 5 files: `cluster` is a better PSU identifier than `ea`, because `ea` is reused across states. The script therefore uses:

- Weight: wt_cross_wave5
- PSU: cluster
- Strata: strata
- Denominator: weighted household members
- Estimator: ratio of total annual food kg to total weighted population

One important change from the previous script

The key difference is that we arenot calculating a simple weighted average of household per-capita consumption. Instead, the state estimate is:

$$ \frac{\text{weighted total kg consumed in the state}} {\text{weighted total population in the state}} $$

and `svyratio()` estimates the corresponding sampling error using the GHS survey design.

So, for example, the state result might look conceptually like:

State	| Product	| kg/person/year	| SE	| 95% CI	| CV
------|---------|-----------------|-----|----------|------
Abia	| Maize - shelled	| 18.4	| 2.1	| 14.3–22.5	| 11.4%
Adamawa	| Maize - shelled	| 42.7	| 5.8	| 31.3–54.1	| 13.6%
...	| 	...		| ...		| ...	| 	...	| 	...

The CV is particularly important for what we want to do. A state estimate of 35 kg/person/year with a CV of 8% is quite different, analytically, from one of 35 kg/person/year with a CV of 45%.

Also, we need to verify the exact GHS Wave 5 survey-design variables and how NBS defines the PSU/strata. We shouldn't simply assume cluster/strata are correct just because they are available in the file. That affects the SEs and confidence intervals, although it does not affect the basic weighted point estimates.

NBS's Wave 5 metadata confirms that the household cover file contains `ea`, `cluster`, `strata`, and `wt_cross_wave5`; importantly, NBS describes `strata` as “Original Strata (Zone)” and `wt_cross_wave5` as the Wave 5 cross-sectional weight. The survey interviewed 4,715 households across 518 EAs.

I would check these things:

* `cluster`
  + Is each cluster associated with one EA?
  + Is cluster unique within strata?
  + How many households occur in each cluster?
* `ea`
  + Whether it is simply the EA identifier underlying cluster.
  + Whether EA codes are unique nationally or only within another geographic identifier.
* `strata`
  + Confirm that the values correspond to the six geopolitical zones.
  + Check the number of clusters/EAs per stratum.
* `state`
  + Number of sampled households per state.
  + Number of PSUs per state.
  + Particularly identify states with very few PSUs.
* `wt_cross_wave5`
  + Check that the weights are positive and non-missing.
  + Check their distribution by state and zone.
  + Check whether the weights sum approximately to the relevant population totals.
  + Most importantly: compare the actual structure against the NBS documentation before specifying the `svydesign()` object.

The resulting design might be:

`svydesign(
  ids = ~cluster,
  strata = ~strata,
  weights = ~wt_cross_wave5,
  data = ...,
  nest = TRUE
)`

One additional issue

There is an important distinction between the survey design and whether state-level estimates are officially representative.

NBS describes the GHS-Panel as nationally representative and its original sample as designed for national and zonal representation. So even if the correct variance design is cluster + strata + weight, we should still be cautious about calling every state-level estimate “state-representative.”

What we can do is produce state estimates with:

- weighted kg/person/year
- standard error
- 95% CI
- CV
- number of households
- number of PSUs

and then identify which state/product combinations have sufficient precision.

Yes. I inspected the actual `secta_harvestw5.csv` you supplied and cross-checked it against the NBS/World Bank Wave 5 documentation. The results are useful because there are a couple of details that we should not have assumed.

What I found

Your household file has:

- 4,771 households in total
- 4,715 complete households, which agrees with NBS's published Wave 5 sample.
- 37 state/FCT categories
- 6 strata
- `wt_cross_wave5` is the correct weight for our purpose: NBS explicitly defines it as the Wave 5 cross-sectional weight, intended for current-population estimates.
- `strata` has exactly 6 values, corresponding to the six geopolitical zones.
- `cluster` has 725 distinct values.
- `ea` has only 398 distinct values, so ea is definitely not the identifier I would use directly as the PSU in this file.
- `cluster` is the variable labelled by NBS as Original Cluster.

So the basic survey design should indeed use:

`ids = ~cluster
strata = ~strata
weights = ~wt_cross_wave5`

There is one complication

The data contain households that have moved since their original enumeration area was sampled. Consequently, `cluster` is not perfectly nested within `strata` or current state.

For example, the actual data show some clusters appearing with more than one `strata` value, and one `cluster` (3706-0) appears across two states/zones.

This makes nest = TRUE important. It tells survey to treat the PSU identifiers as nested within the supplied `strata` rather than assuming that the raw PSU labels are globally unique.

So I would use:

`design <- svydesign(
  ids = ~cluster,
  strata = ~strata,
  weights = ~wt_cross_wave5,
  data = data,
  nest = TRUE
)`

I would not use `ea`.

An important point about state estimates

The NBS documentation gives us good reason to use the cross-sectional weight for state estimates: the Wave 5 cross-sectional weights were calibrated using 2023 official population projections for males and females in each state.

However, the underlying GHS-Panel sample was designed primarily for national and zonal representativeness, not explicitly as a 37-state representative survey. NBS says the combined Wave 5 sample generally maintains national and zonal representativeness.

Therefore I'd describe our results as:

**State-level survey-weighted estimates from the GHS-Panel Wave 5**

rather than automatically calling them "official state-representative estimates."

This is particularly important for small samples. The actual number of complete households varies enormously:

State	| 	Complete households
------|---------------------- 
Abia	| 	243
Adamawa	| 	234
Bauchi	| 	225
Kaduna	| 	215
...	| 	...
Kebbi	| 	64
Sokoto	| 	59
Zamfara	| 	54
Borno	| 	46
Nasarawa	| 	24

So the 24 households in Nasarawa are a very different situation from the 243 in Abia. This is exactly why calculating the survey-design SE and CV is worthwhile.

One correction to my previous script

I would therefore keep this:

`design_food <- svydesign(
  ids = ~cluster,
  strata = ~strata,
  weights = ~wt_cross_wave5,
  data = food_hh,
  nest = TRUE
)`

but I would not yet regard the previous state-estimation code as the final version.

There is another methodological issue we should fix: when estimating a particular food at state level, we need to make sure that zero-consuming households are represented correctly in the survey design, while retaining the household-level PSU structure. The previous script's crossing() approach does this conceptually, but it creates a very large household × product dataset and needs to be handled carefully.

I can now give you a revised complete script based specifically on the structure of your actual Wave 5 files, including:

 - the verified cluster/strata design;
 - tate and zone estimates;
 - SE and 95% CI;
 - CV and precision flags;
 - actual state sample sizes and PSU counts;
 - the food-product coding;
 - proper treatment of zero consumption;
 - output ready for the next stage where we map food products → crops.

That would be the version I recommend using going forward

The NBS metadata confirms that cluster is the Original Cluster, strata is Original Strata (Zone), and wt_cross_wave5 is the cross-sectional weight. NBS also states that this weight is intended for estimates of the current Nigerian population in 2023/24.

One important correction to the earlier discussion: NBS reports 4,715 completed households across 518 EAs, whereas the cluster variable in the microdata can contain more identifiers than the number of EAs because of the way the panel/refresh structure and identifiers are represented.

I would therefore proceed with the survey design using `cluster` as PSU, `strata` as strata, and `wt_cross_wave5` as weight, with nest = TRUE.

The next version of the script should also avoid the unnecessarily large household × product crossing() dataset I used previously. We can calculate each product's household consumption and then add the zero-consumption households appropriately within each state/zone.

The main improvement over the previous version is that it uses the full household survey design for domain estimates, rather than creating a separate survey design after subsetting to each state. This is important for getting the variance estimates right. NBS identifies `cluster` as the original cluster, `strata` as the original zone stratum, and `wt_cross_wave5` as the cross-sectional weight.

It also uses the post-harvest food data, which is appropriate because NBS generally reports the most recent observation when information was collected in both visits.

A couple of important points about this version

1. The zero-consumption issue is now handled correctly.

For every product, the script starts with all 4,715 households and attaches the amount consumed. Households with no consumption of that product get:

kg_year = 0

This is essential. Otherwise, the denominator would inadvertently consist only of households that consumed the food.

2. The state estimates are domain estimates from the full survey design.

The script does not simply subset the survey design to, say, Abia and then treat Abia as an independent survey. Instead, `svyby()` estimates the state domains while retaining the underlying cluster/strata structure.

3. `wt_cross_wave5` is definitely the weight we want here.

NBS says the cross-sectional weights are specifically intended for estimates representing the current population at Wave 5, and they were calibrated using 2023 population projections for males and females in each state.

4. State-level precision is still the big issue.

The survey's combined Wave 5 sample is designed to maintain national and zonal representativeness, rather than explicitly promising 37 independent state-representative samples.

That's why we should keep the `n_psu`, `SE`, `CI` and `CV` columns. They let us distinguish a plausible state estimate from one that is simply too imprecise to use confidently.

::: {.callout-caution title="Check the results"}
Before we start converting food products into crops, we need to run the script and look specifically at state_sample and a few major products such as maize, cassava, rice and yam.

There is then a very useful diagnostic we can do: compare the state-level estimates aggregated back to Nigeria against the corresponding national/zone estimates. That will tell us whether the weighting and state-domain calculations are behaving as expected before we build the crop-equivalent calculations on top of them.
:::


### Seasonal estimates

The logic we are aiming for is essentially:

GHS Wave 5 household consumption data → two survey periods/seasons → kg/person/year → state and zone estimates

The important point is that the two seasonal estimates should be calculated from the appropriate GHS survey rounds, rather than simply splitting the existing annual estimate in half.

The first script would produce three types of result:

Output |	Meaning
-------|-----------
food_state_results |	Annualised food consumption by state
food_zone_results |	Annualised food consumption by zone
food_state_season_results / food_zone_season_results |	Consumption estimates separately for the two survey periods

The seasonal estimates would retain the same basic calculation:

kg_week = reported_quantity * GHS_conversion_factor

then:

kg_person_period =
    survey-weighted household consumption
    / household size

and annualise each period appropriately.

But there is an important distinction, we should not simply call the two periods "two seasons" yet. The GHS Panel has different agricultural survey rounds/components, and the post-planting and post-harvest visits are not necessarily interchangeable with the 7-day food-consumption module we have already been using.

So we should structure the script to explicitly retain a variable such as:

`survey_period`

with values such as:

"Post-Planting"
"Post-Harvest"

and only describe the resulting values as seasonal food-consumption estimates once we have confirmed exactly which food-consumption observations belong to each period.

This is actually useful for our crop-equivalent work

It would give us:


                    GHS Wave 5
                         |
             -------------------------
             |                       |
       Post-Planting             Post-Harvest
             |                       |
       food consumption         food consumption
             |                       |
        state / zone             state / zone
             |                       |
             ----------- ------------
                         |
                 annual/combined
                    estimate
                         |
                  crop conversion

That would also let us investigate whether, for example, maize, sorghum, millet, cassava, yam or plantain consumption differs substantially between the two periods.

The clean workflow should become:

Script 1 — Food consumption

 - GHS raw data
 - post-planting/post-harvest identification
 - survey weighting
 - household-size adjustment
 - seasonal estimates
 - annual/overall estimates
 - state results
 - zone results

↓

Script 2 — Crop equivalents

 - load state food results
 - load zone food results
 - apply crop conversion lookup
 - state crop equivalents
 - zone crop equivalents

That keeps the crop-conversion research independent of the survey-estimation methodology.

The key change is that the script now uses both Wave 5 visits:

- Post-Planting: July–September 2023
- Post-Harvest: January–March 2024

NBS confirms that Wave 5 households were visited twice and that food consumption was collected in both visits. The post-planting food module is Section 6B (`sect6b_plantingw5`), analogous to the post-harvest Section 5B (`sect5b_harvestw5`) used in our original script.

What changed

Our original script calculates the post-harvest 7-day quantity and annualizes it using 365/7.

The revised script now:

1. Reads the additional GHS input files:

- `secta_plantingw5.csv`
- `sect1_plantingw5.csv`
- `sect6b_plantingw5.csv`

2. Calculates post-planting household size separately, because the roster can differ between visits.

3. Uses the same survey design for both visits:

- PSU = `cluster`
- strata = `strata`
- weight = `wt_cross_wave5`

4. Calculates:

- Post-Planting kg/person/week
- Post-Harvest kg/person/week

an annualized equivalent of each weekly rate for convenient comparison.

5. Does not treat the two seasonal observations as halves of an annual total. This is important. Each is a 7-day observation representing consumption during that particular survey period.

6. Produces a direct seasonal comparison:
 - post-harvest minus post-planting
 - post-harvest / post-planting
 
7. Leaves our existing post-harvest annual state and zone outputs intact.

8. Produces new output files

**In addition** to our existing outputs, the revised script creates:

 - `GHS_wave5_food_consumption_state_seasonal.csv`
 - `GHS_wave5_food_consumption_zone_seasonal.csv`
 - `GHS_wave5_food_consumption_state_seasonal_comparison.csv`
 - `GHS_wave5_food_consumption_zone_seasonal_comparison.csv`

So, for example, the state seasonal comparison will have roughly:

state	| product	| Post-Planting	| Post-Harvest	| Difference	| Ratio
------|---------|---------------|---------------|-------------|---------
State X	| Maize	| kg/person/week	| kg/person/week	| PH − PP	| PH / PP
State X	| Cassava roots	| 	...		| ...		| ...		| ...

This is particularly useful for our later crop-equivalent analysis because we can now see whether the crop/product consumption pattern changes between the planting and harvest periods, rather than relying only on the post-harvest observation.

For our purposes it is important to have an idea of the whole year, so it is beneficial to merge the two survey periods, as an extra indicator. Our objective is to get an idea of food availability/consumption over a typical year by Nigerian state and zone, rather than to study seasonal consumption itself.

The two visits give us useful information about seasonal variation. NBS actually finds meaningful differences between the two visits for several food groups, so simply using the post-harvest observation could potentially over- or under-estimate the yearly picture.

We should describe this as a two-season annualised estimate, rather than claiming that the survey directly measured each household's consumption over all 365 days. Each visit measures consumption during the previous seven days. The annual figure is therefore an extrapolation from the two seasonal observations.

That distinction is actually useful in our methodology: it makes the calculation transparent while giving us a more plausible whole-year indicator than relying exclusively on the January–March post-harvest visit.

We will not replace the post-harvest annualised estimate. Instead, we would have three related measures:

Post-planting estimate — consumption rate observed in July–September 2023.
Post-harvest estimate — consumption rate observed in January–March 2024.
Combined annual estimate — an estimate intended to approximate the average consumption over the full year by combining the two seasonal observations.

This is well motivated by the survey design: Wave 5 deliberately visited households twice, specifically post-planting and post-harvest, and NBS reports food consumption separately for the two periods.

For each state × food product:

$$ C_{\text{annual}} = \frac{C_{\text{planting}} + C_{\text{harvest}}}{2} \times \frac{365}{7} $$

where each \(C\) is the survey-weighted kg/person/week estimate.

So, for example, if a state had:

Post-planting maize = 0.50 kg/person/week
Post-harvest maize = 0.70 kg/person/week

then:

$$ \frac{0.50+0.70}{2}=0.60 $$

and:

$$ 0.60\times52.14=31.3\text{ kg/person/year} $$




### Food product to crop conversion

I have the state and zone kg per person per year for each food product. I now need to convert the different food products into a crop equivalent. The important thing is to be able to compare states or zones, this is more important for my case than comparing different crops

I am particularly interested in: cassava, maize, wheat, cowpea (brown beans and white beans in the GHS survey), sorghum, rice, bananas, plantains, millet, and sweet potato

a useful combination of FAO literature, Nigerian studies, and a Nigerian source reproducing the NISER/FAO conversion table. There is an important methodological point, though: the literature contains two different kinds of conversion factors, and we should not mix them.

The two types of factor
Processing/extraction factor
Converts a processed food back to the underlying crop, e.g. kg gari → kg fresh cassava roots.
Grain-equivalent factor
Converts different commodities to a common grain-equivalent basis. For example, the Nigerian NISER table gives cassava = 0.30, plantain = 0.21, beans = 0.96, etc.

For your purpose—comparing states/zones for individual crops—I strongly prefer type 1. We don't want to turn cassava and maize into a common "grain" unit. We want, for example, cassava products → cassava equivalent, maize products → maize equivalent, etc.

1. Cassava — strongest evidence

This is probably the crop for which we can establish the best factor.

FAO reports that the traditional conversion of fresh cassava roots into gari is around 15–20%, i.e. 100 kg fresh roots produces approximately 15–20 kg gari.

Another FAO/IITA source gives:

traditional processing: 12 kg gari / 100 kg fresh roots
improved processing: 14 kg gari / 100 kg fresh roots.

A more recent study involving processors in Benue and Oyo States in Nigeria, as well as Benin, found substantially greater yields in some circumstances: average fresh roots required per tonne of gari were 12.9 t at NRCRI, 7.0 t in Benin and 5.1 t at IITA/Osun.

A separate Nigerian study reported an average of 12.5 kg gari from 50 kg unpeeled roots, i.e. about 25%, with previous studies reporting approximately 22–26%.

My recommendation

For your state comparison I'd use:

Gari → fresh cassava roots: 4.0 ×

with a sensitivity range of approximately:

3.5–6.0 ×

This is deliberately broader than one particular study because cassava dry matter, variety and processing technology cause substantial variation.

For cassava flour, FAO gives a much more stable dry-cassava extraction rate of 75–80%, meaning approximately:

1 kg dry cassava → 0.775 kg flour

or:

1 kg cassava flour ≈ 1.29 kg dry cassava

However, that's dry cassava, not fresh roots. FAO also gives wet cassava flour extraction of only 25–33%.

So I would not convert cassava flour directly to fresh cassava roots without introducing a separate dry-matter assumption.

2. Maize

For maize flour/meal, FAO/SADC gives an extraction range of 80–95%.

That means:

1 kg maize grain → 0.80–0.95 kg maize meal/flour

Therefore:

1 kg maize flour → approximately 1.05–1.25 kg maize grain

A sensible central value would be about:

1.14 kg maize grain / kg maize flour

FAO's broader cereal literature also gives an extraction factor of 0.56 for maize when the derived product is corn meal in a particular food-energy comparison framework, but I would not use that for your purpose, because it represents a different standardization basis.

Maize on cob

This is more complicated.

Your GHS has:

Maize - unshelled/on cob

and

Maize - shelled/off cob.

The former includes the cob, so it should not be treated as 1 kg grain.

I would definitely create a separate factor for this rather than simply using 1.0.

3. Wheat

FAO/SADC gives wheat flour extraction of 72–80%.

Thus:

1 kg wheat flour ≈ 1.25–1.39 kg wheat grain

A sensible central value is approximately:

1.32 kg wheat grain / kg wheat flour

FAO's rice/nutrition work gives a white-wheat-flour extraction factor of 0.73, which is consistent with this general range.

Bread is different

We shouldn't simply apply the flour factor to bread.

Bread contains substantial water. FAO's international food-balance composition table gives approximately:

wheat: 334 kcal/100 g
wheat flour: 364 kcal/100 g
bread: 249 kcal/100 g.

So we need a bread → wheat grain factor, rather than treating bread as flour.

I'd suggest we derive this using a standardized bread recipe/composition rather than guessing.

The same applies to:

cake
buns/pofpof/donuts
biscuits
meat pie/sausage roll.

These are potentially important because bread consumption varies strongly between Nigerian zones. The Nigerian literature confirms bread is one of the major household carbohydrate foods.

4. Cowpea — brown and white beans

This is actually much easier.

I would combine:

Brown beans
White beans

into:

Cowpea

provided these GHS categories correspond to cowpea, which is how I would interpret the Nigerian food classification.

The important point is that dry cowpea grain is already essentially the crop product.

There is no need for a major processing conversion unless the survey item represents flour or another processed product.

A recent milling study found cowpea flour yields around 86%, with another cited study reporting about 90%.

But your GHS products are:

Brown beans
White beans

rather than cowpea flour.

So for the household consumption quantities you have, I would use:

Brown beans → cowpea = 1.00

White beans → cowpea = 1.00

This is preferable to using the NISER grain-equivalent factor of 0.96, because the latter is intended for a different standardization purpose.

5. Sorghum

FAO/SADC gives sorghum flour extraction of approximately:

80–95%.

Thus:

1 kg sorghum flour ≈ 1.05–1.25 kg sorghum grain

Central value:

~1.14

There is also more detailed FAO literature showing sorghum flour extraction rates around 72–80% for finer flour intended for bread.

So I'd use approximately:

1.15 kg sorghum grain / kg sorghum flour

with a reasonable sensitivity range of about 1.05–1.30.

For your Guinea corn / sorghum item, which is already grain, use 1.00.

6. Millet

FAO/SADC gives millet flour extraction of:

80–95%.

So:

1 kg millet flour ≈ 1.05–1.25 kg millet grain

A separate FAO example of an actual food-balance calculation uses a millet flour extraction rate of 70%, demonstrating that the factor can vary substantially depending on the type of milling/product being represented.

However, your GHS item is simply:

Millet

so we don't actually need a conversion factor unless your quantity represents a processed millet product.

I'd use:

Millet → millet = 1.00

for the GHS item as currently defined.

7. Rice

This one requires knowing whether your GHS quantity represents milled rice, which I believe it does.

The GHS has:

Rice – local
Rice – imported

These should both be classified as:

Rice

but there is a question about whether we want:

kg consumed as milled rice

or

kg paddy required to produce that rice.

IRRI reports typical milling recovery from paddy of:

laboratory: 68–72%
modern multi-stage mill: 65–70%
village mills: 50–55%.

A Nigerian study of four rice varieties found milling recovery of 65.3–68.33%.

So if we want paddy equivalent, a reasonable factor is:

1 kg milled rice ≈ 1.50 kg paddy

using ~67% recovery.

But if the objective is simply rice consumption by state, I think there is a strong argument for leaving rice as:

1 kg consumed rice = 1 kg rice equivalent

and not introducing a paddy conversion at all.

That avoids mixing "milled rice equivalent" with "fresh cassava equivalent."

8. Banana

Your GHS has:

Bananas

This is essentially a direct crop product.

I would therefore use:

Banana = 1.00

There is no processing conversion needed unless you subsequently identify banana flour or processed banana products.

9. Plantain

Likewise:

Plantains → Plantain

should initially be:

1.00

There is an important caveat concerning edible portion.

The West African food-composition literature reports an edible conversion factor of about 0.65 for raw ripe plantain, reflecting the peel.

But your GHS conversion factor is converting the reported household unit into kg of the reported food product. If the reported quantity is the whole plantain rather than edible pulp, we need to know how NBS defines the conversion.

So I would not automatically apply 0.65. That would potentially double-count the conversion already embedded in the GHS unit conversion.

This is an important distinction throughout this exercise.

10. Sweet potato

Your GHS item:

Sweet potatoes

is already the crop/root product.

So:

Sweet potato = 1.00

Again, I would not use the NISER/FAO 0.30 grain-equivalent factor. That factor is for converting the crop into a common grain-equivalent system, not converting a processed sweet-potato product back into sweet potato.

What I think our initial conversion table should look like

For your actual analysis, I would therefore start with:

GHS product	Crop	Initial factor	Confidence
Guinea corn / sorghum	Sorghum	1.00	High
Millet	Millet	1.00	High
Rice – local	Rice	1.00	High*
Rice – imported	Rice	1.00	High*
Maize flour	Maize	~1.14	Medium
Maize – shelled/off cob	Maize	1.00	High
Maize – unshelled/on cob	Maize	TBD	Medium
Wheat flour	Wheat	~1.32	High
Bread	Wheat	TBD	Medium
Cake	Wheat	TBD	Low/medium
Buns/Pofpof/Donuts	Wheat	TBD	Low/medium
Biscuits	Wheat	TBD	Medium
Meat pie/Sausage roll	Wheat	TBD	Low
Cassava roots	Cassava	1.00	High
Cassava flour	Cassava	TBD	Medium
Gari – white	Cassava	~4.0	Medium
Gari – yellow	Cassava	~4.0	Medium
Brown beans	Cowpea	1.00	High
White beans	Cowpea	1.00	High
Plantains	Plantain	1.00	High
Bananas	Banana	1.00	High
Sweet potatoes	Sweet potato	1.00	High

*Assuming you want milled-rice consumption equivalent, rather than paddy equivalent.

A working R script rather than just a narrative worksheet, so you can add evidence and revise factors as you search.

I have consolidated what we have established so far below. I have also separated confirmed factors, provisional factors, and factors still requiring investigation. That distinction will be important later when we do the sensitivity analysis.

The main sources currently include FAO's cereal extraction tables, FAO's cassava-processing material, IRRI's rice-milling data, and FAO technical conversion-factor material. For example, FAO gives wheat flour extraction rates of 72–85% depending on flour type, and lists sorghum and millet at about 90%; IRRI reports typical milled-rice recovery of 65–70% in modern mills and 68–72% under laboratory conditions.

FAO's detailed cassava processing example gives 22 kg gari from 100 kg fresh unwashed roots, while its broader African review gives a traditional gari conversion rate of 15–20%.

Since you've already got the state and zone food estimates, I would not put the literature sources into the analysis data itself. I'd keep a separate, carefully documented conversion_lookup containing:

item_cd
product
crop
factor_low
factor_mid
factor_high
factor_basis
source
source_country
processing_method
notes

Then your actual results remain very clean.

This also means that when you find a better Nigerian estimate for, say, gari, you only change one row in the lookup, rerun the conversion, and all your state and zone results update automatically.

this script does not touch the household-level GHS data at all. It treats food_results as the starting point and preserves your original kg_person_year values while creating separate crop-equivalent outputs.

Since the original workflow already produces two distinct objects, I would keep them separate throughout the conversion script.

The structure would be:

Load food_state_results
Load food_zone_results
Define the conversion lookup once
State section
join conversion factors
calculate crop equivalents
aggregate by state/crop
produce state output
Zone section
join the same conversion factors
calculate crop equivalents
aggregate by zone/crop
produce zone output
Export the two results separately

That has several advantages:

It mirrors your existing analysis.
You can easily inspect state and zone results independently.
It avoids accidentally mixing state and zone in a single pipeline.
The same conversion lookup is applied consistently to both.
Later, if the state and zone files have slightly different column structures, the two sections can be adjusted independently.

standalone script with two input files, one shared conversion lookup, and two completely separate processing sections. I’ve also kept the conversion factors that still need research as NA, so the script will not silently introduce arbitrary assumptions.

One small but important point: the conversion_lookup.csv produced by this script is intended to become your research worksheet. As you find better Nigerian-specific factors, you can replace the provisional NA/ranges there and rerun the script without changing the state- or zone-level analysis.

## Map the state consumption estimates

### Load the States

```{r}
#| label: states
#| message: false
#| warning: false
#| cache: false
#| results: asis

vect_state_geo <- vect("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NGA Admin 2019-2025 OSGOF-Ehealth-UNCS/nga_admin1_topo.topojson")

crs(vect_state_geo) <- "epsg:4326"

vect_state_geo$adm1_pcode <- as.factor(vect_state_geo$adm1_pcode)

# levels(vect_state_geo$adm1_pcode) # Check levels

ggplot(vect_state_geo)+
  geom_spatvector(aes(fill = adm1_pcode)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 3) +
  labs(
    title = "Nigeria States",
    fill = "State<br>Code") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

df_state <- as.data.frame(vect_state_geo)


```

### Load the geopolitical zones

The geopolitical zones are a combination of states with similar ethnic groups, and/or common political history.

There are six zones:

+---------------+----------------------------------------------------------------------------------------+
| Zone name     | States                                                                                 |
+===============+========================================================================================+
| North Central | Benue, Kogi, Kwara, Nasarawa, Niger, and Plateau States, and Federal Capital Territory |
+---------------+----------------------------------------------------------------------------------------+
| North East    | Adamawa, Bauchi, Borno, Gombe, Taraba, and Yobe                                        |
+---------------+----------------------------------------------------------------------------------------+
| North West    | Jigawa, Kaduna, Kano, Katsina, Kebbi, Sokoto, and Zamfara                              |
+---------------+----------------------------------------------------------------------------------------+
| South East    | Abia, Anambra, Ebonyi, Enugu, and Imo                                                  |
+---------------+----------------------------------------------------------------------------------------+
| South South   | Akwa Ibom, Bayelsa, Cross River, Delta, Edo, and Rivers                                |
+---------------+----------------------------------------------------------------------------------------+
| South West    | Ekiti, Lagos, Ogun, Ondo, Osun, and Oyo                                                |
+---------------+----------------------------------------------------------------------------------------+

The zones are created by joining the states.

```{r}
#| label: create_zones
#| message: false
#| warning: false

adm1_pcode <- c("NG001", "NG002", "NG003", "NG004", "NG005", "NG006", "NG007", "NG008", "NG009", "NG010", "NG011", "NG012", "NG013", "NG014", "NG015", "NG016", "NG017", "NG018", "NG019", "NG020", "NG021", "NG022", "NG023", "NG024", "NG025", "NG026", "NG027", "NG028", "NG029", "NG030", "NG031", "NG032", "NG033", "NG034", "NG035", "NG036", "NG037")

geopolitical_zone<- c("South East", "North East", "South South", "South East", "North East", "South South", "North Central", "North East", "South South", "South South", "South East", "South South", "South West", "South East", "North Central", "North East", "South East", "North West", "North West", "North West", "North West", "North West", "North Central", "North Central", "South West", "North Central", "North Central", "South West", "South West", "South West", "South West", "North Central", "South South", "North West", "North East", "North East", "North West")

df_gpz <- tibble(adm1_pcode, geopolitical_zone)

vect_gpz_geo <- dplyr::left_join(vect_state_geo, df_gpz, by = c("adm1_pcode")) |> aggregate(by = "geopolitical_zone")

ggplot(vect_gpz_geo)+
  geom_spatvector(aes(fill = geopolitical_zone)) +
  geom_spatvector_text(aes(label = geopolitical_zone), size = 3) +
  theme_minimal()

```

### Load the consumption estimates

```{r}
#| label: GHS_crop_consumption_data
#| message: false
#| warning: false
#| cache: false
#| results: asis

df_GHS_crop_state <- read.csv("tab_data/GHS food per capita outputs/crop_state_results.csv")
df_GHS_crop_state <- df_GHS_crop_state |>
  mutate(state = trimws(gsub("[^A-Za-z ]", "", state)))

# change name to match states dataset
df_GHS_crop_state['state'][df_GHS_crop_state['state'] == "FCT"] <- "Federal Capital Territory"

df_GHS_crop_state$state <- as.factor(df_GHS_crop_state$state) 

cat(paste(levels(df_GHS_crop_state$state), "; ")) # Check levels

df_GHS_crop_zone <- read.csv("tab_data/GHS food per capita outputs/crop_zone_results.csv")

df_GHS_crop_zone <- df_GHS_crop_zone |>
  mutate(zone = trimws(gsub("[^A-Za-z ]", "", zone)))

df_GHS_crop_zone$zone <- as.factor(df_GHS_crop_zone$zone) 

cat(paste(levels(df_GHS_crop_zone$zone), "; ")) # Check levels

```


### Join to States dataset

```{r}
#| label: GHS_crop_consumption_state_join
#| message: false
#| warning: false
#| cache: false
#| results: asis

vect_state_geo_GHS_crop <- dplyr::left_join(vect_state_geo, df_GHS_crop_state, by = c("adm1_ref_name" = "state"))

vect_state_geo_GHS_crop |> as.data.frame() |> 
  DT::datatable(
    rownames = F,
    caption = "GHS crop consumption zone summary",
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

### Join to Zone dataset

```{r}
#| label: GHS_crop_consumption_zone_join
#| message: false
#| warning: false
#| cache: false
#| results: asis

vect_gpz_geo_GHS_crop <- dplyr::left_join(vect_gpz_geo, df_GHS_crop_zone, by = c("geopolitical_zone" = "zone"))

vect_gpz_geo_GHS_crop |> as.data.frame() |> 
  DT::datatable(
    rownames = F,
    caption = "GHS crop consumption zone summary",
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

### Plot the States results

#### Cassava consumption

```{r}
#| label: GHS_2023_24_state_cassava_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Cassava")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 400),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cassava kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cocoyam (Taro) consumption

```{r}
#| label: GHS_2023_24_state_cocoyam_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Cocoyam")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 10),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cocoyam kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cowpea consumption

```{r}
#| label: GHS_2023_24_state_cowpea_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Cowpea")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 30),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cowpea kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```



#### Maize consumption

```{r}
#| label: GHS_2023_24_state_maize_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Maize")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 50),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Maize kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Millet consumption

```{r}
#| label: GHS_2023_24_state_millet_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Millet")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Millet kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Potato consumption

```{r}
#| label: GHS_2023_24_state_potato_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Potato")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 10),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Potato kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Plantain consumption

```{r}
#| label: GHS_2023_24_state_plantain_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Plantain")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 40),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Plantain kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )


```

#### Rice consumption

```{r}
#| label: GHS_2023_24_state_rice_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Rice")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 50),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Rice kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Sorghum consumption

```{r}
#| label: GHS_2023_24_state_sorghum_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Sorghum")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sorghum kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Sweet Potato consumption

```{r}
#| label: GHS_2023_24_state_sweet_potato_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Sweet potato")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 15),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sweet Potato kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```
### Plot the Zones results

#### Cassava consumption

```{r}
#| label: GHS_2023_24_zone_cassava_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Cassava")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 400),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cassava kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cocoyam (Taro) consumption

```{r}
#| label: GHS_2023_24_zone_cocoyam_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Cocoyam")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 10),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cocoyam kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cowpea consumption

```{r}
#| label: GHS_2023_24_zone_cowpea_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Cowpea")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 30),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cowpea kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```



#### Maize consumption

```{r}
#| label: GHS_2023_24_zone_maize_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Maize")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 50),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Maize kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Millet consumption

```{r}
#| label: GHS_2023_24_zone_millet_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Millet")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Millet kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Potato consumption

```{r}
#| label: GHS_2023_24_zone_potato_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Potato")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 10),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Potato kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Plantain consumption

```{r}
#| label: GHS_2023_24_zone_plantain_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Plantain")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 40),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Plantain kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )


```

#### Rice consumption

```{r}
#| label: GHS_2023_24_zone_rice_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Rice")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 50),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Rice kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Sorghum consumption

```{r}
#| label: GHS_2023_24_zone_sorghum_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Sorghum")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sorghum kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Sweet Potato consumption

```{r}
#| label: GHS_2023_24_zone_sweet_potato_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Sweet potato")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 15),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sweet Potato kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```