---
title: "Nigeria Crop Consumption"
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    page-layout: full
    embed-resources: true
    self-contained-math: true
    toc-expand: true
    theme: united
    code-fold: true
    code-summary: "Show the code"
    css: css/style.css
bibliography: biofort_quarto_better.bib
---

```{r results='hide', warning=FALSE, echo=FALSE, message=FALSE}

library(tidyverse)
library(terra)
library(tidyterra)
library(readODS)
library(ggplot2)
library(ggtext)
library(tidyverse)
library(readODS)
library(kableExtra)
library(DT)
library(geojsonsf)
library(survey)

```
# Introduction

The global Biofortification Priority Index (BPI) [@asare-marfoPrioritizingCountriesBiofortifcation2013] uses an indicator of
consumption per capita per year for each specific crop, standardised to fit with the maximum and minimum values encountered.

These consumption values are not readily available in the reports of the 2021 NFCMS or the 2023/24 General Household Survey–Panel. An advantage of the 2023/24 General Household Survey–Panel (GHS-Panel Wave 5) is that it has a household food-consumption module publicly available from the Nigeria National Bureau of Statistics.

The 2023/24 GHS comprised two surveys of the same panel households: a Post-Planting survey which was conducted between July and September 2023, and a Post-Harvest survey conducted between January and March 2024. One of the sections of the surveys concentrated on food consumption.

For each household and food product item data were collected on:

 - quantity consumed during the past 7 days
 - quantity purchased during the past 30 days
 - expenditure on food

The dataset for each survey has over half a million food-item/household observations.

The microdata have individual food items, so it is possible investigate the actual item list and identify products such as maize, maize flour, maize meal, gari, cassava flour, fufu, etc.
From the households we are able to generate reliable estimates of weekly food product consumption for the population within each geopolitical zone and less representative but useful estimates at the state level.

With these weekly per capita estimates we can apply conversion factors to calculate the crop-equivalent consumption, per week and by merging the results from the two surveys generate a yearly estimate, consistent with that used in the global BPI.

# Methods

Our objective is to calculate per capita crop product consumption over a typical year, by Nigerian state and zone.

To capture variations in food consumption during the year we took advantage of the two survey periods in the GHS Wave 5 survey:

- Post-Planting: July–September 2023
- Post-Harvest: January–March 2024

NBS found meaningful differences between the two visits for several food groups, so simply using the post-harvest observation could potentially over- or under-estimate the yearly picture.

We describe this as a two-season annualised estimate, rather than claiming that the survey directly measured each household's consumption over all 365 days. Each visit measures consumption during the previous seven days. The annual figure is therefore an extrapolation from the two seasonal observations.

This distinction made the calculation transparent while giving us a more plausible whole-year indicator than relying exclusively on a single survey visit.

Our overall workflow was:

                    GHS Wave 5
                         |
             -------------------------
             |                       |
       Post-Planting             Post-Harvest
             |                       |
       food consumption         food consumption
             |                       |
        state / zone             state / zone
             |                       |
             ----------- ------------
                         |
                 annual/combined
                    estimate
                         |
                  crop conversion

Our workflow kept the crop-conversion research independent of the survey-estimation methodology. The workflow also allowed us to investigate whether, for example, maize, sorghum, millet, cassava, yam or plantain consumption differed substantially between the two periods.

The two main calculations were:

**Per capita food consumption calculation**

 - load GHS post-planting & post-harvest raw data
 - generate descriptive statistics
 - apply survey weighting
 - household-size adjustment
 - generate state/zone diagnostics
 - calculate seasonal estimates & precision summaries
 - combine results to give annual estimates
 - save state results
 - save zone results

         ↓

**Crop equivalent calculation**

 - load combined annual state food results
 - load combined annual zone food results
 - create a crop conversion lookup table
 - apply crop conversion lookup table
 - calcuate state crop equivalents
 - calcuate zone crop equivalents


## Per capita food consumption calculation

The first calculation determined estimates for the per capita **food product** consumption for each food product by state and zone.

The initial estimates of food product consumption were at the **household** level. These estimates were calculated separately for each survey period as:

kg_week = reported_quantity * GHS_conversion_factor

To move from household to the general population we estimated the total kg consumed by the state/zone ÷ estimated population represented by the state/zone:

$$ \boxed{ \frac{ \sum_i w_i Q_i }{ \sum_i w_i HHSize_i } } $$

where:

$Q_i$ = annual kg consumed by household $i$
$HHSize_i$ = number of people in household $i$
$w_i$ = household survey weight.

This avoided giving a five-person household and a one-person household equal influence simply because they are both households.


#### Alternative food product types

The food item coding is complicated

`item_cd` is a labelled text field in our CSV, e.g.:

16. Maize flour
18. Cassava  flour
30. Cassava - roots
32. Gari - white
33. Gari - yellow

There is also `item_cd_alt`, which is extremely important.

For example, the standard category:
23. Other grains and flour

can have:

231. Semovita/semolina
232. Akanu/Pap/Obi/Fufu
234. Plantain flour
235. Macaroni/pasta/noodles
238. INDOMIE noodles

Similarly:

107. Other fish or seafood
can contain:
1071. Crayfish
1072. Periwinkle
1073. Stockfish

So we retained `item_cd_alt` when building our food dictionary.

3. There is a complication with cassava. Some respondents coded their consumption as:

30. Cassava - roots

but their free-text description says things such as:

WRAP (FUFU)

or

WRAP COOKEDFUFU

This means the NBS coding isn't perfectly clean at the product level.

For example, some households have:

`item_cd` = 30. Cassava - roots
`unit` = WRAP
`description` = WRAP FUFU

So we shouldn't blindly interpret every kilogram assigned to item 30 as raw cassava root.

This is actually a strong argument for retaining the original variables and creating our own carefully documented crop and product classification.

The "other food" mechanism means we should not limit ourselves to `item_cd`.

The post-harvest data contain 503 `item_os_cd` records, including fufu, plantain flour, Bambara nut, palm kernel, locust beans, Kunu, Zobo and many other foods.

So when we actually process the microdata, I would:

 - retain the standard `item_cd`;
 - retain `item_os_cd`;
 - retain the free-text `item_os`;
 - construct a master food-code lookup;
 - identify crop/product relationships;
 - calculate consumption in kg;
 - apply household weights;
 - calculate the six-zone estimates;
 - separately identify observations that require manual classification.

That last step is important because otherwise we could silently lose things like fufu.

#### The consumption calculation

The relevant variables are:

`s5bq1`       consumed this item in previous 7 days?
`s5bq2a`      quantity consumed
`s5bq2b`      unit
`s5bq2b_os`   other unit
`s5bq2c`      size
`s5bq2_cvn`   conversion factor

NBS describes `s5bq2_cvn` specifically as:

conversion factor item-unit-size-zone

and it ranges from 0.001 to 50 in the released data.

So the basic calculation is:

quantity_kg_7days =
    s5bq2a * s5bq2_cvn
    

::: {.callout-note title="Testing the conversion factor"}
The conversion factor works exactly as we hoped

For example, one household reported:

*"25 cigarette cups of local rice"*

and the conversion factor is:

0.232

giving:

25 × 0.232 = 5.8 kg

That is clearly the intended interpretation.

Another household reported:

*"2 large loaves of bread"*

with conversion factor 0.780, giving:

2 × 0.780 = 1.56 kg

So we can safely calculate:

kg_7days = `s5bq2a` * `s5bq2_cvn`

The conversion factor is therefore effectively kg per reported quantity/unit/size/zone combination.
::: 

#### Survey weights

Because our objective is a representative estimate of food consumption in 2023/24 rather than a longitudinal comparison we used the cross-sectional weight `wt_cross_wave5`, rather than the longitudinal weight. NBS stated that the cross-sectional weight was calibrated to the 2023 population and was intended for representative current-population estimates.

#### Household size

We calculated the household size directly from `sect1_harvestw5.dta`.

That gives us:
hh_size =
    number of household members in the Wave 5 post-harvest roster

This allowed us to inspect which roster records represented current household members.

We joined:

`sect5b_harvestw5`
        |
        | `hhid`
        v
`secta_harvestw5`
        |
        +---- `zone`
        +---- `state`
        +---- `wt_cross_wave5`




4. Missing conversion factors are rare—but need treatment

There are 80,702 "Yes" consumption observations.

Of those, only 210 have a missing `s5bq2_cvn`.

For the crops we're interested in, the situation is generally excellent.

For example:

Food | Consumed |	Missing conversion
-----|----------|--------------------
Maize flour |		1,106 |		0
Yam flour |		425	 |	0
Cassava flour |		886	 |	1
Maize, unshelled |		26 |		0
Maize, shelled |		444 |		0
Cassava roots |		530 |	38
Yam roots |	2,138 |		0
Gari white |		1,926 |		0
Gari yellow |		1,011 |		1
Plantains |		1,080 |		0

So for maize, yam, gari, etc., we're in very good shape.

Cassava roots require some additional cleaning because of those 38 observations, many involving locally described units such as wrap or satana.

5. Household size can also be calculated reliably

The household roster contains 28,256 records.

We found that we cannot simply count rows.

For existing household members, `s1q4` = "1. YES" identifies membership, but new household members have `NEWMEMBER` = 1 and `s1q4` is blank.

So I would define household size as:

member =
    `s1q4` == "1. YES" |
    `NEWMEMBER` == 1
    
This gives 4,716 households with a calculable household size.

The average household size in these records is about 5.44 people, with a range of 1–25.

That's a useful validation check.

6. The weight is exactly where expected

`secta_harvestw5.csv` contains:

`wt_wave5`
`wt_longpanel_wave5`
`wt_cross_wave5`

For our purpose the most appropriate is `wt_cross_wave5`, because we're estimating population consumption in 2023/24, rather than analysing changes among the longitudinal panel.


#### State level estimates

The Wave 5 sample covers all 36 states plus FCT, and the survey documentation says the original GHS design allocated PSUs across all 37 state-level areas. The NBS catalogue confirms that state is available in the Wave 5 data with codes 1–37. The survey weights make state estimates legitimate. The GHS is designed to produce estimates at the state level, and the household data contain the state identifier and survey weights. Therefore, for a state \(s\), we can calculate:

$$ \text{kg/person/year}_{s} = \frac{ \sum_i w_i Q_i }{ \sum_i w_i H_i } $$

where:

\(Q_i\) = annualised kg consumed by household \(i\)
\(H_i\) = household size
\(w_i\) = Wave 5 cross-sectional household weight.

So the same estimator we developed for the geopolitical zones can be applied to each of the 37 state-level areas.

However, the number of sampled households varies considerably by state. So we should not simply calculate a state mean and assume that all 37 estimates are equally precise. As a result we added precision diagnostics, these allowed us to flag estimates such as:

Nasarawa — a particular food — 8.4 kg/person/year (CV 38%)

as substantially less reliable than:

Kaduna — same food — 9.1 kg/person/year (CV 8%).

This is particularly important for relatively uncommon foods.

The food-consumption files contained over 500,000 household × food observations, but those were repeated food records within only about 4,700 households. We could not treat the 500,000+ observations as independent observations when calculating uncertainty.

For proper standard errors, we accounted for the complex survey design — PSU, strata and weights.
So the final R workflow used the 'survey' package, rather than just 'dplyr', for the state-level estimates and confidence intervals.

The main output should show:

- `geopolitical_zone`
- `state`
- `food_code`
- `food_product`
- `crop`
- `product_type`
- `n_households`
- `n_consuming`
- `kg_person_year`
- `se`
- `cv`
- `ci_lower`
- `ci_upper`

Then we could aggregate the state estimates to the six zones.

The script uses:

- Weight: wt_cross_wave5
- PSU: cluster
- Strata: strata
- Denominator: weighted household members
- Estimator: ratio of total annual food kg to total weighted population


We did not calculate a simple weighted average of household per-capita consumption. Instead, the state estimate was:

$$ \frac{\text{weighted total kg consumed in the state}} {\text{weighted total population in the state}} $$

and `svyratio()` estimates the corresponding sampling error using the GHS survey design.

So, for example, the state result might look conceptually like:

State	| Product	| kg/person/year	| SE	| 95% CI	| CV
------|---------|-----------------|-----|----------|------
Abia	| Maize - shelled	| 18.4	| 2.1	| 14.3–22.5	| 11.4%
Adamawa	| Maize - shelled	| 42.7	| 5.8	| 31.3–54.1	| 13.6%
...	| 	...		| ...		| ...	| 	...	| 	...

The CV was particularly important. A state estimate of 35 kg/person/year with a CV of 8% is quite different, analytically, from one of 35 kg/person/year with a CV of 45%.

Also, we needed to verify the exact GHS Wave 5 survey-design variables and how NBS defined the PSU/strata. We didn't simply assume cluster/strata were correct just because they were available in the file. That affects the SEs and confidence intervals, although it does not affect the basic weighted point estimates.

NBS's Wave 5 metadata confirms that the household cover file contains `ea`, `cluster`, `strata`, and `wt_cross_wave5`; importantly, NBS describes `strata` as “Original Strata (Zone)” and `wt_cross_wave5` as the Wave 5 cross-sectional weight. The survey interviewed 4,715 households across 518 EAs.

I would check these things:

* `cluster`
  + Is each cluster associated with one EA?
  + Is cluster unique within strata?
  + How many households occur in each cluster?
* `ea`
  + Whether it is simply the EA identifier underlying cluster.
  + Whether EA codes are unique nationally or only within another geographic identifier.
* `strata`
  + Confirm that the values correspond to the six geopolitical zones.
  + Check the number of clusters/EAs per stratum.
* `state`
  + Number of sampled households per state.
  + Number of PSUs per state.
  + Particularly identify states with very few PSUs.
* `wt_cross_wave5`
  + Check that the weights are positive and non-missing.
  + Check their distribution by state and zone.
  + Check whether the weights sum approximately to the relevant population totals.
  + Most importantly: compare the actual structure against the NBS documentation before specifying the `svydesign()` object.

The resulting design might be:

`svydesign(
  ids = ~cluster,
  strata = ~strata,
  weights = ~wt_cross_wave5,
  data = ...,
  nest = TRUE
)`

One additional issue

There is an important distinction between the survey design and whether state-level estimates are officially representative.

NBS describes the GHS-Panel as nationally representative and its original sample as designed for national and zonal representation. So even if the correct variance design is cluster + strata + weight, we should still be cautious about calling every state-level estimate “state-representative.”

What we can do is produce state estimates with:

- weighted kg/person/year
- standard error
- 95% CI
- CV
- number of households
- number of PSUs

and then identify which state/product combinations have sufficient precision.

Yes. I inspected the actual `secta_harvestw5.csv` you supplied and cross-checked it against the NBS/World Bank Wave 5 documentation. The results are useful because there are a couple of details that we should not have assumed.

What I found

Your household file has:

- 4,771 households in total
- 4,715 complete households, which agrees with NBS's published Wave 5 sample.
- 37 state/FCT categories
- 6 strata
- `wt_cross_wave5` is the correct weight for our purpose: NBS explicitly defines it as the Wave 5 cross-sectional weight, intended for current-population estimates.
- `strata` has exactly 6 values, corresponding to the six geopolitical zones.
- `cluster` has 725 distinct values.
- `ea` has only 398 distinct values, so ea is definitely not the identifier I would use directly as the PSU in this file.
- `cluster` is the variable labelled by NBS as Original Cluster.

So the basic survey design should indeed use:

`ids = ~cluster
strata = ~strata
weights = ~wt_cross_wave5`

There is one complication

The data contain households that have moved since their original enumeration area was sampled. Consequently, `cluster` is not perfectly nested within `strata` or current state.

For example, the actual data show some clusters appearing with more than one `strata` value, and one `cluster` (3706-0) appears across two states/zones.

This makes nest = TRUE important. It tells survey to treat the PSU identifiers as nested within the supplied `strata` rather than assuming that the raw PSU labels are globally unique.

So I would use:

`design <- svydesign(
  ids = ~cluster,
  strata = ~strata,
  weights = ~wt_cross_wave5,
  data = data,
  nest = TRUE
)`

I would not use `ea`.

Therefore I'd describe our results as:

**State-level survey-weighted estimates from the GHS-Panel Wave 5**

rather than automatically calling them "official state-representative estimates."

This is particularly important for small samples. The actual number of complete households varies enormously:

State	| 	Complete households
------|---------------------- 
Abia	| 	243
Adamawa	| 	234
Bauchi	| 	225
Kaduna	| 	215
...	| 	...
Kebbi	| 	64
Sokoto	| 	59
Zamfara	| 	54
Borno	| 	46
Nasarawa	| 	24

So the 24 households in Nasarawa are a very different situation from the 243 in Abia. This is exactly why calculating the survey-design SE and CV is worthwhile.

One correction to my previous script

I would therefore keep this:

`design_food <- svydesign(
  ids = ~cluster,
  strata = ~strata,
  weights = ~wt_cross_wave5,
  data = food_hh,
  nest = TRUE
)`

but I would not yet regard the previous state-estimation code as the final version.

There is another methodological issue we should fix: when estimating a particular food at state level, we need to make sure that zero-consuming households are represented correctly in the survey design, while retaining the household-level PSU structure. The previous script's crossing() approach does this conceptually, but it creates a very large household × product dataset and needs to be handled carefully.

I can now give you a revised complete script based specifically on the structure of your actual Wave 5 files, including:

 - the verified cluster/strata design;
 - tate and zone estimates;
 - SE and 95% CI;
 - CV and precision flags;
 - actual state sample sizes and PSU counts;
 - the food-product coding;
 - proper treatment of zero consumption;
 - output ready for the next stage where we map food products → crops.

That would be the version I recommend using going forward

The NBS metadata confirms that cluster is the Original Cluster, strata is Original Strata (Zone), and wt_cross_wave5 is the cross-sectional weight. NBS also states that this weight is intended for estimates of the current Nigerian population in 2023/24.

One important correction to the earlier discussion: NBS reports 4,715 completed households across 518 EAs, whereas the cluster variable in the microdata can contain more identifiers than the number of EAs because of the way the panel/refresh structure and identifiers are represented.

I would therefore proceed with the survey design using `cluster` as PSU, `strata` as strata, and `wt_cross_wave5` as weight, with nest = TRUE.

The next version of the script should also avoid the unnecessarily large household × product crossing() dataset I used previously. We can calculate each product's household consumption and then add the zero-consumption households appropriately within each state/zone.

The main improvement over the previous version is that it uses the full household survey design for domain estimates, rather than creating a separate survey design after subsetting to each state. This is important for getting the variance estimates right. NBS identifies `cluster` as the original cluster, `strata` as the original zone stratum, and `wt_cross_wave5` as the cross-sectional weight.

It also uses the post-harvest food data, which is appropriate because NBS generally reports the most recent observation when information was collected in both visits.

A couple of important points about this version

1. The zero-consumption issue is now handled correctly.

For every product, the script starts with all 4,715 households and attaches the amount consumed. Households with no consumption of that product get:

kg_year = 0

This is essential. Otherwise, the denominator would inadvertently consist only of households that consumed the food.

2. The state estimates are domain estimates from the full survey design.

The script does not simply subset the survey design to, say, Abia and then treat Abia as an independent survey. Instead, `svyby()` estimates the state domains while retaining the underlying cluster/strata structure.

3. `wt_cross_wave5` is definitely the weight we want here.

NBS says the cross-sectional weights are specifically intended for estimates representing the current population at Wave 5, and they were calibrated using 2023 population projections for males and females in each state.

4. State-level precision is still the big issue.

The survey's combined Wave 5 sample is designed to maintain national and zonal representativeness, rather than explicitly promising 37 independent state-representative samples.

That's why we should keep the `n_psu`, `SE`, `CI` and `CV` columns. They let us distinguish a plausible state estimate from one that is simply too imprecise to use confidently.




We structured the calculation to explicitly retain a `survey_period` variable with values:

"Post-Planting"
"Post-Harvest"

The per capita food product consumption calculation produced six main results:

Output |	Meaning
-------|-----------
1. food_state_results |	Annualised food product consumption estimates by state for **each** survey period
2. food_zone_results |	Annualised food product consumption by zone for **each** survey period
3. food_state_season_results | Consumption estimates by state for **both** survey periods (row-wise)
4. food_zone_season_results |	Consumption estimates by zone for **both** survey periods (row-wise)
5. food_state_season_comparison_results |	Comparison of estimates by state for **both** survey periods (column-wise)
6. food_zone_season_comparison_results | Comparison of estimates by zone for **both** survey periods (column-wise)


The state results will have the estimates plus as assessment of the precision:

state	| 	zone	| 	survey_period	| 	product	| 	estimate_kg_person_week	| 	se	| 	ci_low	| 	ci_high	| 	cv	| 	n_households	| 	n_psu	| 	weighted_population	| 	median_household_size	| 	mean_household_size	| 	estimate_kg_person_year_equivalent	| 	precision_flag
---|----|----|----|----|----|----|----|----|----|----|----|----|----|----|-----
15. Gombe	| 2. North East	| Post-Harvest	| Maize flour	| 0.410829435	| 0.075329055	| 0.263184487	| 	0.56	| 0.18	| 133	| 17	| 4058199	| 6	| 6.66	| 21.42	| CV 10-20%


The zone results contain fewer variables:

survey_period	| 	product	| 	zone	| 	estimate_kg_person_week	| 	se	| 	cv	| 	ci_low	| 		ci_high	| 	estimate_kg_person_year_equivalent
---|----|----|----|----|----|----|----|----
Post-Planting	| 		Maize flour	| 		1. North Central	| 		0.38	| 		0.04	| 		0.12	| 		0.29	| 		0.46	| 		19.62

The state and zone seasonal comparison results contain the following variables:


zone | state	| product	| Post-Planting	| Post-Harvest	| Difference	| Ratio
------|------|---------|---------------|---------------|-------------|---------
Zone Y	| State X	| Maize	| kg/person/week	| kg/person/week	| PH − PP	| PH / PP
Zone Y		| State X	| Cassava roots	| 	...		| ...		| ...		| ...

This allows us to see whether the crop/product consumption pattern changes between the planting and harvest periods.

### Summary

The script:

- reads the three files;
- constructs household size from the roster;
- uses wt_cross_wave5;
- uses the 7-day consumption quantity, not purchases;
- converts reported quantities to kg using s5bq2_cvn;
- annualises the 7-day quantity;
- calculates population-weighted kg/person/year for each food product and geopolitical zone;
- retains the item_cd_alt and item_os information;
- creates a crop/product lookup;
- flags unusual observations rather than silently discarding them;
- writes CSV files that you can inspect in Excel.


## Crop equivalent calculation

For each state × food product:

$$ C_{\text{annual}} = \frac{C_{\text{planting}} + C_{\text{harvest}}}{2} \times \frac{365}{7} $$

where each \(C\) is the survey-weighted kg/person/week estimate.

So, for example, if a state had:

Post-planting maize_flour = 0.50 kg/person/week
Post-harvest maize_flour = 0.70 kg/person/week

then:

$$ \frac{0.50+0.70}{2}=0.60 $$

and:

$$ 0.60\times52.14=31.3\text{ kg/person/year} $$



::: {.callout-note title="Calculating the crop-equivalent consumption"}
For example, suppose the household records:

Gari → 10 kg consumed

We could have two possible measures:

Food-product consumption

Gari = 10 kg/person/year

or, with the appropriate processing/yield conversion:

Crop-equivalent consumption

Cassava equivalent = X kg/person/year

That would allow you to aggregate all cassava-derived products:

gari + fufu + cassava flour + other cassava foods
→ total cassava-equivalent consumption kg/person/year

Likewise:

maize grain + maize flour + maize meal + maize products
→ total maize-equivalent consumption
:::




8. Intermediate food dictionary

Rather than immediately producing a simple:

crop | zone | kg/person/year
-----|------|---------------

dataset, we should construct an intermediate food dictionary.

Something like:

GHS code	| 	GHS description	| 	Crop 	Product	| 	Include?	| 	Notes
----------|-------------------|-----------------|-------------|----------
16	      | 	Maize flour		    | Maize	Flour	  | 	Yes		| 
20	      | 	Maize unshelled		| Maize	Cob	    | 	Yes		| conversion needed for crop equivalent
22	      | 	Maize shelled	    | Maize	Grain	  | 	Yes		| 
30	      | 	Cassava roots	    | Cassava	Roots	|  Yes*		| inspect fufu descriptions
18	      | 	Cassava flour	    | Cassava	Flour	| 	Yes		| 
32	      | 	Gari white	      | Cassava	Gari	| 	Yes		| 
33	      | 	Gari yellow	      | Cassava	Gari	| 	Yes		| 
232	      | 	Akanu/Pap/Obi/Fufu| ?	Fufu	      | 	Yes  	| investigate composition
234	      | 	Plantain flour	  | Plantain Flour|   Yes		| 
31	      | 	Yam roots		      | Yam	Roots	    | 	Yes		| 
17	      | 	Yam flour	        | Yam	Flour   	| 	Yes		| 

This gives us a transparent classification that you can subsequently modify.

I have the state and zone kg per person per year for each food product. I now need to convert the different food products into a crop equivalent. The important thing is to be able to compare states or zones, this is more important for my case than comparing different crops

I am particularly interested in: cassava, maize, wheat, cowpea (brown beans and white beans in the GHS survey), sorghum, rice, bananas, plantains, millet, and sweet potato

a useful combination of FAO literature, Nigerian studies, and a Nigerian source reproducing the NISER/FAO conversion table. There is an important methodological point, though: the literature contains two different kinds of conversion factors, and we should not mix them.

The two types of factor
Processing/extraction factor
Converts a processed food back to the underlying crop, e.g. kg gari → kg fresh cassava roots.
Grain-equivalent factor
Converts different commodities to a common grain-equivalent basis. For example, the Nigerian NISER table gives cassava = 0.30, plantain = 0.21, beans = 0.96, etc.

For your purpose—comparing states/zones for individual crops—I strongly prefer type 1. We don't want to turn cassava and maize into a common "grain" unit. We want, for example, cassava products → cassava equivalent, maize products → maize equivalent, etc.

1. Cassava — strongest evidence

This is probably the crop for which we can establish the best factor.

FAO reports that the traditional conversion of fresh cassava roots into gari is around 15–20%, i.e. 100 kg fresh roots produces approximately 15–20 kg gari.

Another FAO/IITA source gives:

traditional processing: 12 kg gari / 100 kg fresh roots
improved processing: 14 kg gari / 100 kg fresh roots.

A more recent study involving processors in Benue and Oyo States in Nigeria, as well as Benin, found substantially greater yields in some circumstances: average fresh roots required per tonne of gari were 12.9 t at NRCRI, 7.0 t in Benin and 5.1 t at IITA/Osun.

A separate Nigerian study reported an average of 12.5 kg gari from 50 kg unpeeled roots, i.e. about 25%, with previous studies reporting approximately 22–26%.

My recommendation

For your state comparison I'd use:

Gari → fresh cassava roots: 4.0 ×

with a sensitivity range of approximately:

3.5–6.0 ×

This is deliberately broader than one particular study because cassava dry matter, variety and processing technology cause substantial variation.

For cassava flour, FAO gives a much more stable dry-cassava extraction rate of 75–80%, meaning approximately:

1 kg dry cassava → 0.775 kg flour

or:

1 kg cassava flour ≈ 1.29 kg dry cassava

However, that's dry cassava, not fresh roots. FAO also gives wet cassava flour extraction of only 25–33%.

So I would not convert cassava flour directly to fresh cassava roots without introducing a separate dry-matter assumption.

2. Maize

For maize flour/meal, FAO/SADC gives an extraction range of 80–95%.

That means:

1 kg maize grain → 0.80–0.95 kg maize meal/flour

Therefore:

1 kg maize flour → approximately 1.05–1.25 kg maize grain

A sensible central value would be about:

1.14 kg maize grain / kg maize flour

FAO's broader cereal literature also gives an extraction factor of 0.56 for maize when the derived product is corn meal in a particular food-energy comparison framework, but I would not use that for your purpose, because it represents a different standardization basis.

Maize on cob

This is more complicated.

Your GHS has:

Maize - unshelled/on cob

and

Maize - shelled/off cob.

The former includes the cob, so it should not be treated as 1 kg grain.

I would definitely create a separate factor for this rather than simply using 1.0.

3. Wheat

FAO/SADC gives wheat flour extraction of 72–80%.

Thus:

1 kg wheat flour ≈ 1.25–1.39 kg wheat grain

A sensible central value is approximately:

1.32 kg wheat grain / kg wheat flour

FAO's rice/nutrition work gives a white-wheat-flour extraction factor of 0.73, which is consistent with this general range.

Bread is different

We shouldn't simply apply the flour factor to bread.

Bread contains substantial water. FAO's international food-balance composition table gives approximately:

wheat: 334 kcal/100 g
wheat flour: 364 kcal/100 g
bread: 249 kcal/100 g.

So we need a bread → wheat grain factor, rather than treating bread as flour.

I'd suggest we derive this using a standardized bread recipe/composition rather than guessing.

The same applies to:

cake
buns/pofpof/donuts
biscuits
meat pie/sausage roll.

These are potentially important because bread consumption varies strongly between Nigerian zones. The Nigerian literature confirms bread is one of the major household carbohydrate foods.

4. Cowpea — brown and white beans

This is actually much easier.

I would combine:

Brown beans
White beans

into:

Cowpea

provided these GHS categories correspond to cowpea, which is how I would interpret the Nigerian food classification.

The important point is that dry cowpea grain is already essentially the crop product.

There is no need for a major processing conversion unless the survey item represents flour or another processed product.

A recent milling study found cowpea flour yields around 86%, with another cited study reporting about 90%.

But your GHS products are:

Brown beans
White beans

rather than cowpea flour.

So for the household consumption quantities you have, I would use:

Brown beans → cowpea = 1.00

White beans → cowpea = 1.00

This is preferable to using the NISER grain-equivalent factor of 0.96, because the latter is intended for a different standardization purpose.

5. Sorghum

FAO/SADC gives sorghum flour extraction of approximately:

80–95%.

Thus:

1 kg sorghum flour ≈ 1.05–1.25 kg sorghum grain

Central value:

~1.14

There is also more detailed FAO literature showing sorghum flour extraction rates around 72–80% for finer flour intended for bread.

So I'd use approximately:

1.15 kg sorghum grain / kg sorghum flour

with a reasonable sensitivity range of about 1.05–1.30.

For your Guinea corn / sorghum item, which is already grain, use 1.00.

6. Millet

FAO/SADC gives millet flour extraction of:

80–95%.

So:

1 kg millet flour ≈ 1.05–1.25 kg millet grain

A separate FAO example of an actual food-balance calculation uses a millet flour extraction rate of 70%, demonstrating that the factor can vary substantially depending on the type of milling/product being represented.

However, your GHS item is simply:

Millet

so we don't actually need a conversion factor unless your quantity represents a processed millet product.

I'd use:

Millet → millet = 1.00

for the GHS item as currently defined.

7. Rice

This one requires knowing whether your GHS quantity represents milled rice, which I believe it does.

The GHS has:

Rice – local
Rice – imported

These should both be classified as:

Rice

but there is a question about whether we want:

kg consumed as milled rice

or

kg paddy required to produce that rice.

IRRI reports typical milling recovery from paddy of:

laboratory: 68–72%
modern multi-stage mill: 65–70%
village mills: 50–55%.

A Nigerian study of four rice varieties found milling recovery of 65.3–68.33%.

So if we want paddy equivalent, a reasonable factor is:

1 kg milled rice ≈ 1.50 kg paddy

using ~67% recovery.

But if the objective is simply rice consumption by state, I think there is a strong argument for leaving rice as:

1 kg consumed rice = 1 kg rice equivalent

and not introducing a paddy conversion at all.

That avoids mixing "milled rice equivalent" with "fresh cassava equivalent."

8. Banana

Your GHS has:

Bananas

This is essentially a direct crop product.

I would therefore use:

Banana = 1.00

There is no processing conversion needed unless you subsequently identify banana flour or processed banana products.

9. Plantain

Likewise:

Plantains → Plantain

should initially be:

1.00

There is an important caveat concerning edible portion.

The West African food-composition literature reports an edible conversion factor of about 0.65 for raw ripe plantain, reflecting the peel.

But your GHS conversion factor is converting the reported household unit into kg of the reported food product. If the reported quantity is the whole plantain rather than edible pulp, we need to know how NBS defines the conversion.

So I would not automatically apply 0.65. That would potentially double-count the conversion already embedded in the GHS unit conversion.

This is an important distinction throughout this exercise.

10. Sweet potato

Your GHS item:

Sweet potatoes

is already the crop/root product.

So:

Sweet potato = 1.00

Again, I would not use the NISER/FAO 0.30 grain-equivalent factor. That factor is for converting the crop into a common grain-equivalent system, not converting a processed sweet-potato product back into sweet potato.

What I think our initial conversion table should look like

For your actual analysis, I would therefore start with:

GHS product	Crop	Initial factor	Confidence
Guinea corn / sorghum	Sorghum	1.00	High
Millet	Millet	1.00	High
Rice – local	Rice	1.00	High*
Rice – imported	Rice	1.00	High*
Maize flour	Maize	~1.14	Medium
Maize – shelled/off cob	Maize	1.00	High
Maize – unshelled/on cob	Maize	TBD	Medium
Wheat flour	Wheat	~1.32	High
Bread	Wheat	TBD	Medium
Cake	Wheat	TBD	Low/medium
Buns/Pofpof/Donuts	Wheat	TBD	Low/medium
Biscuits	Wheat	TBD	Medium
Meat pie/Sausage roll	Wheat	TBD	Low
Cassava roots	Cassava	1.00	High
Cassava flour	Cassava	TBD	Medium
Gari – white	Cassava	~4.0	Medium
Gari – yellow	Cassava	~4.0	Medium
Brown beans	Cowpea	1.00	High
White beans	Cowpea	1.00	High
Plantains	Plantain	1.00	High
Bananas	Banana	1.00	High
Sweet potatoes	Sweet potato	1.00	High

*Assuming you want milled-rice consumption equivalent, rather than paddy equivalent.

We separated confirmed factors, provisional factors, and factors still requiring investigation. That distinction will be important later for a sensitivity analysis.

The main sources currently include FAO's cereal extraction tables, FAO's cassava-processing material, IRRI's rice-milling data, and FAO technical conversion-factor material. For example, FAO gives wheat flour extraction rates of 72–85% depending on flour type, and lists sorghum and millet at about 90%; IRRI reports typical milled-rice recovery of 65–70% in modern mills and 68–72% under laboratory conditions.

FAO's detailed cassava processing example gives 22 kg gari from 100 kg fresh unwashed roots, while its broader African review gives a traditional gari conversion rate of 15–20%.

We decided to put the literature sources in a separate, carefully documented conversion_lookup containing:

item_cd
product
crop
factor_low
factor_mid
factor_high
factor_basis
source
source_country
processing_method
notes

This also meant that if we found a better Nigerian estimate for, say, gari, we only needed to change one row in the lookup, rerun the conversion, and all our state and zone results would update automatically.

The crop equivalent conversion calculation did not touch the household-level GHS data at all. It treated the results of the per capita food product consumption as the starting point and preserved our original kg_person_week values.


## Household level GHS data

The raw data files are available for use if necessary, the summaries per geopolitical zone are sufficient for some aspects of the atlas, but other indicators - such as the consumption of individual crop food products - require access to the raw household level survey data.

::: {.callout-note title="Relevant Datasets in the GHS"}

1. Files we need

For the primary calculation we need three files:

`sect5b_harvestw5.dta` — food consumption
`secta_harvestw5.dta` — household ID, zone and wt_cross_wave5
`sect1_harvestw5.dta` — household roster, from which we calculate household size

The food file contains 509,220 household × food-item observations and 26 variables.

The cover file contains 4,771 households and the three Wave 5 weight variables, including `wt_cross_wave5`.

The roster contains 28,256 individual records across the households.



There are five GHS datasets that are relevant for the theme of crop food product consumption:

*sect6b_plantingw5.csv*\
Data collected through Post-Planting Household Questionnaire Section 6B (Food Consumption). This section records household-food item level information on the quantity and value of food consumed within the household during the past 7 days, as well as the quantity and value of food purchased during the past 30 days.

*sect6c_plantingw5.csv*\
Data collected through Post-Planting Household Questionnaire Section 6C (Aggregate Food Consumption). This section records household-food item level information on the number of days the household consumed each food group item during the last 7 days.

*sect5b_harvestw5.csv*\
Data collected through Post- Harvest Household Questionnaire Section 5B (Food Consumption). This section records household-food item level information on the quantity and value of food consumed within the household during the past 7 days, as well as the quantity and value of food purchased during the past 30 days.

*sect5c_harvestw5.csv*\
Data collected through Post-Harvest Household Questionnaire Section 5C (Aggregate Food Consumption). This section records household-food item level information on the number of days the household consumed each food group item during the last 7 days.

*sect3d_harvestw5.csv*\
Data collected through Post-Harvest Household Questionnaire Section 3D (Women Dietary Diversity). This section records women-level information on women’s nutrition diet for the previous day.


:::

### Number of households


```{r}
#| label: GHS_household_data
#| message: false
#| warning: false
#| cache: false
#| results: asis

# 1. Post planting survey

df_GHS_secta_plantingw5 <- read.csv(
  "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Planting Wave 5/Household/secta_plantingw5.csv")

zone_sample_pp <- df_GHS_secta_plantingw5 |>
  group_by(
    zone
  ) |>
  summarise(
    n_households_pp = n(),
    .groups = "drop"
  ) |>
  arrange(zone)

state_sample_pp <- df_GHS_secta_plantingw5 |>
  group_by(
    zone,
    state
  ) |>
  summarise(
    n_households_pp = n(),
    .groups = "drop"
  ) |> 
  arrange(zone, state)


# 2. Post harvest survey

df_GHS_secta_harvestw5 <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Harvest Wave 5/Household/secta_harvestw5.csv")

zone_sample_ph <- df_GHS_secta_harvestw5 |>
  group_by(
    zone
  ) |>
  summarise(
    n_households_ph = n(),
    .groups = "drop"
  ) |>
  arrange(zone)

state_sample_ph <- df_GHS_secta_harvestw5 |>
  group_by(
    zone,
    state
  ) |>
  summarise(
    n_households_ph = n(),
    .groups = "drop"
  ) |>
  arrange(zone, state)

```

There were 5,067 households in the post planting survey, and 4771 households in the post harvest survey.

These were distributed more or less equally among the zones (@tbl-zone), but there were large differences between states. The state with the highest number of households was Abia, with 258 households in the post planting survey and 244 in the post harvest survey (@tbl-state). The state with the fewest number of households surveyed was Nasarawa, with only 24 households in both surveys.

```{r}
#| label: tbl-zone
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "hh_zone"

zone_sample <- zone_sample_pp |> left_join(zone_sample_ph, by = "zone")

zone_sample |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "# of households (post planting survey)", "# of households (post harvest survey)"),
    caption = "Households per zone",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

```{r}
#| label: tbl-state
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "hh_state"

state_sample <- state_sample_pp |> left_join(select(state_sample_ph, -zone), by = "state")

state_sample |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "State", "# of households (post planting survey)", "# of households (post harvest survey)"),
    caption = "Households per state",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(0, 'asc'), list(2, 'asc')),
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

### Household composition

We are interested in calculating the food consumption of different crops per person per year. The food consumption in the survey is at the household level. We therefore need to know the number of people per household so that the food quantities can be shared among the household and thereafter extrapolated to the population.

The composition of the household is contained within a 'roster' dataset. We remove household members in the roster who are no longer residing in the household (the GHS is a panel survey), and add any new members.

```{r}
#| label: GHS_roster_data
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ----1. post planting household size----

df_GHS_sect1_plantingw5 <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Planting Wave 5/Household/sect1_plantingw5.csv") |>
  mutate(
    household_member = s1q4 == "1. YES" |
      NEWMEMBER == 1
  )


# get size of each household

df_hh_size_pp <- df_GHS_sect1_plantingw5 |>
  filter(household_member) |>
  group_by(hhid) |>
  mutate(hh_size = n()) |>
  select(zone, state, hhid, hh_size) |>
  distinct()


# Get zone samples of hh_size

zone_sample_hh_size_pp <- df_hh_size_pp |>
  group_by(
    zone
  ) |>
  summarise(
    median_hh_size_pp = median(hh_size),
    max_hh_size_pp = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone)


# Get state samples of hh_size

state_sample_hh_size_pp <- df_hh_size_pp |>
  group_by(
    zone, state
  ) |>
  summarise(
    median_hh_size_pp = median(hh_size),
    max_hh_size_pp = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone, state)


# ----2. post harvest household size----

df_GHS_sect1_harvestw5 <- read.csv("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Harvest Wave 5/Household/sect1_harvestw5.csv") |>
  mutate(
    household_member = s1q4 == "1. YES" |
      NEWMEMBER == 1
  )

# get size of each household

df_hh_size_ph <- df_GHS_sect1_harvestw5 |>
  filter(household_member) |>
  group_by(hhid) |>
  mutate(hh_size = n()) |>
  select(zone, state, hhid, hh_size) |>
  distinct()


# Get zone samples of hh_size

zone_sample_hh_size_ph <- df_hh_size_ph |>
  group_by(
    zone
  ) |>
  summarise(
    median_hh_size_ph = median(hh_size),
    max_hh_size_ph = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone)


# Get state samples of hh_size

state_sample_hh_size_ph <- df_hh_size_ph |>
  group_by(
    zone, state
  ) |>
  summarise(
    median_hh_size_ph = median(hh_size),
    max_hh_size_ph = max(hh_size),
    .groups = "drop"
  ) |>
  arrange(zone, state)

```

The average (median) household size per zone varies between 4 and 8 in the post-planting survey, with the three northern zones displaying higher averages than the three southern zones (@tbl-hh_size_zone). The zone with the largest individual household size was North West zone.

This pattern is mirrored in the post-harvest survey.

```{r}
#| label: tbl-hh_size_zone
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "Household size per zone"

zone_sample_hh_size <- zone_sample_hh_size_pp |> left_join(zone_sample_hh_size_ph, by = "zone")

zone_sample_hh_size |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "Median household size (post planting survey)", "Max household size (post planting survey)", "Median household size (post harvest survey)", "Max household size (post harvest survey)"),
    caption = "Household size per zone",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 1),
      # lengthMenu = c(15, 40),
      pageLength = 6
    ))

```

At the state level the (median) household size per zone varies between 3 in Cross River state and 9 in Borno state in the post-planting survey. The state that had the highest individual household was Kaduna (@tbl-hh_size_state).

In the post-harvest survey there were five southern states with a median of 3 households, and four northern states with a median of eight members per household.

```{r}
#| label: tbl-hh_size_state
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "Household size per state"

state_sample_hh_size <- state_sample_hh_size_pp |> left_join(select(state_sample_hh_size_ph, -zone), by = "state")

state_sample_hh_size |> 
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "State", "Median household size (post planting survey)", "Max household size (post planting survey)", "Median household size (post harvest survey)", "Max household size (post harvest survey)"),
    caption = "Household size per state",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(0, 'asc'), list(2, 'desc')),
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 1),
      # lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

### Valid households for our analysis

Not all of the households in the surveys are valid for our analysis. Some households in the panel were not able to be contacted or the interview was incomplete.
Those households are removed to leave a consistent set of 4,715 households from the two surveys.

```{r}
#| label: GHS_analysis_households
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ----1. post planting survey ----

df_hh_analysis_pp <- df_GHS_secta_plantingw5 |>
  filter(
    interview_result == "1. COMPLETE",
    !is.na(wt_cross_wave5),
    wt_cross_wave5 > 0
  ) |>
  select(
    hhid,
    zone,
    state,
    lga,
    sector,
    ea,
    cluster,
    strata,
    wt_cross_wave5
  ) |>
  left_join(
    select(df_hh_size_pp, -c(zone, state)),
    by = "hhid"
  ) |>
  filter(
    !is.na(hh_size),
    hh_size > 0
  )

# ----2. post harvest survey ----

df_hh_analysis_ph <- df_GHS_secta_harvestw5 |>
  filter(
    interview_result == "1. COMPLETE",
    !is.na(wt_cross_wave5),
    wt_cross_wave5 > 0
  ) |>
  select(
    hhid,
    zone,
    state,
    lga,
    sector,
    ea,
    cluster,
    strata,
    wt_cross_wave5
  ) |>
  left_join(
    select(df_hh_size_ph, -c(zone, state)),
    by = "hhid"
  ) |>
  filter(
    !is.na(hh_size),
    hh_size > 0
  )

```


```{r}
#| label: GHS_sample_analysis_households
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ----1. post planting survey ----

zone_sample_analysis_pp <- df_hh_analysis_pp |>
  group_by(
    zone
  ) |>
  summarise(
    n_households = n(),
    n_psu = n_distinct(cluster),
    weighted_population = sum(
      wt_cross_wave5 * hh_size,
      na.rm = TRUE
    ),
    median_household_size = median(
      hh_size,
      na.rm = TRUE
    ),
    mean_household_size = weighted.mean(
      hh_size,
      wt_cross_wave5,
      na.rm = TRUE
    ),
    max_household_size = max(
      hh_size,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  arrange(zone)

state_sample_analysis_pp <- df_hh_analysis_pp |>
  group_by(
    zone,
    state
  ) |>
  summarise(
    n_households = n(),
    n_psu = n_distinct(cluster),
    weighted_population = sum(
      wt_cross_wave5 * hh_size,
      na.rm = TRUE
    ),
    median_household_size = median(
      hh_size,
      na.rm = TRUE
    ),
    mean_household_size = weighted.mean(
      hh_size,
      wt_cross_wave5,
      na.rm = TRUE
    ),
    max_household_size = max(
      hh_size,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  arrange(zone, state)

# ----2. post harvest survey ----

zone_sample_analysis_ph <- df_hh_analysis_ph |>
  group_by(
    zone
  ) |>
  summarise(
    n_households = n(),
    n_psu = n_distinct(cluster),
    weighted_population = sum(
      wt_cross_wave5 * hh_size,
      na.rm = TRUE
    ),
    median_household_size = median(
      hh_size,
      na.rm = TRUE
    ),
    mean_household_size = weighted.mean(
      hh_size,
      wt_cross_wave5,
      na.rm = TRUE
    ),
    max_household_size = max(
      hh_size,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  arrange(zone)

state_sample_analysis_ph <- df_hh_analysis_ph |>
  group_by(
    zone,
    state
  ) |>
  summarise(
    n_households = n(),
    n_psu = n_distinct(cluster),
    weighted_population = sum(
      wt_cross_wave5 * hh_size,
      na.rm = TRUE
    ),
    median_household_size = median(
      hh_size,
      na.rm = TRUE
    ),
    mean_household_size = weighted.mean(
      hh_size,
      wt_cross_wave5,
      na.rm = TRUE
    ),
    max_household_size = max(
      hh_size,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  arrange(zone, state)
```

More households were dropped from the post-planting survey than from the post-harvest survey.

The pattern of the distribution of households among the zones is similar to the full set of households. North East zone has most households in both surveys, while South West has the fewest (@tbl-analysis_zone).

```{r}
#| label: tbl-analysis_zone
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "hh_analysis_zone"

zone_sample_analysis <- zone_sample_analysis_pp |> left_join(zone_sample_analysis_ph, by = "zone")
  
zone_sample_analysis[, c(1,2,8,3,9,4,10,5,11,6,12,7,13)] |> 
  mutate(
    weighted_population.x = prettyunits::pretty_round(weighted_population.x, digits = 0),
    weighted_population.y = prettyunits::pretty_round(weighted_population.y, digits = 0),
    mean_household_size.x = prettyunits::pretty_round(mean_household_size.x, digits = 1),
    mean_household_size.y = prettyunits::pretty_round(mean_household_size.y, digits = 1)
  )|>
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "# of households (post planting survey)", "# of households (post harvest survey)","# of clusters (post planting survey)","# of clusters (post harvest survey)","weighted population (post planting survey)","weighted population (post harvest survey)","median household size (post planting survey)","median household size (post harvest survey)","mean household size (post planting survey)","mean household size (post harvest survey)","max household size (post planting survey)","max household size (post harvest survey)"),
    caption = "Households per zone",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

At the state level the changes are greater, especially for the post planting survey, with states like Zamfara experiencing a reduced number of households for our analysis. However, the effect on the post-harvest number of households is less severe (@tbl-analysis_state), and the number of households per state is now almost the same when comparing the post planting and post harvest surveys.


```{r}
#| label: tbl-analysis_state
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "hh_analysis_state"

state_sample_analysis <- state_sample_analysis_pp |> left_join(select(state_sample_analysis_ph, -zone), by = "state")

state_sample_analysis[, c(1,2,3,9,4,10,5,11,6,12,7,13,8,14)] |> 
  mutate(
    weighted_population.x = prettyunits::pretty_round(weighted_population.x, digits = 0),
    weighted_population.y = prettyunits::pretty_round(weighted_population.y, digits = 0),
    mean_household_size.x = prettyunits::pretty_round(mean_household_size.x, digits = 1),
    mean_household_size.y = prettyunits::pretty_round(mean_household_size.y, digits = 1)
  )|>
  DT::datatable(
    rownames = F,
    colnames=c("Zone", "State","# of households (post planting survey)", "# of households (post harvest survey)","# of clusters (post planting survey)","# of clusters (post harvest survey)","weighted population (post planting survey)","weighted population (post harvest survey)","median household size (post planting survey)","median household size (post harvest survey)","mean household size (post planting survey)","mean household size (post harvest survey)","max household size (post planting survey)","max household size (post harvest survey)"),
    caption = "Households per state",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

<!-- ```{r} -->
<!-- #| label: GHS_roster_data_analysis -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| cache: false -->
<!-- #| results: asis -->

<!-- # ----1. post planting household size---- -->


<!-- # filter the roster using the valid hh_analysis hhid -->

<!-- df_hh_analysis_size_pp <- df_hh_size_pp |> right_join(select(df_hh_analysis_pp, hhid), by = "hhid") -->

<!-- # Get zone samples of hh_size -->

<!-- zone_sample_hh_analysis_size_pp <- df_hh_analysis_size_pp |> -->
<!--   group_by( -->
<!--     zone -->
<!--   ) |> -->
<!--   summarise( -->
<!--     median_hh_size_pp = median(hh_size), -->
<!--     max_hh_size_pp = max(hh_size), -->
<!--     .groups = "drop" -->
<!--   ) |> -->
<!--   arrange(zone) -->


<!-- # Get state samples of hh_size -->

<!-- state_sample_hh_analysis_size_pp <- df_hh_analysis_size_pp |> -->
<!--   group_by( -->
<!--     zone, state -->
<!--   ) |> -->
<!--   summarise( -->
<!--     median_hh_size_pp = median(hh_size), -->
<!--     max_hh_size_pp = max(hh_size), -->
<!--     .groups = "drop" -->
<!--   ) |> -->
<!--   arrange(zone, state) -->

<!-- # ----2. post harvest household size---- -->
<!-- # filter the roster using the valid hh_analysis hhid -->

<!-- df_hh_analysis_size_ph <- df_hh_size_ph |> right_join(select(df_hh_analysis_ph, hhid), by = "hhid") -->


<!-- # Get zone samples of hh_size -->

<!-- zone_sample_hh_analysis_size_ph <- df_hh_analysis_size_ph |> -->
<!--   group_by( -->
<!--     zone -->
<!--   ) |> -->
<!--   summarise( -->
<!--     median_hh_size_ph = median(hh_size), -->
<!--     max_hh_size_ph = max(hh_size), -->
<!--     .groups = "drop" -->
<!--   ) |> -->
<!--   arrange(zone) -->


<!-- # Get state samples of hh_size -->

<!-- state_sample_hh_analysis_size_ph <- df_hh_analysis_size_ph |> -->
<!--   group_by( -->
<!--     zone, state -->
<!--   ) |> -->
<!--   summarise( -->
<!--     median_hh_size_ph = median(hh_size), -->
<!--     max_hh_size_ph = max(hh_size), -->
<!--     .groups = "drop" -->
<!--   ) |> -->
<!--   arrange(zone, state) -->

<!-- ``` -->

<!-- The average sizes of households in the analysis do not change, either for the zonal averages (@tbl-hh_analysis_size_zone) or the state averages (@tbl-hh_analysis_size_state). -->

<!-- ```{r} -->
<!-- #| label: tbl-hh_analysis_size_zone -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| cache: false -->
<!-- #| results: asis -->
<!-- #| tab-cap: "Households in analysis size per zone" -->

<!-- zone_sample_hh_analysis_size <- zone_sample_hh_analysis_size_pp |> left_join(zone_sample_hh_analysis_size_ph, by = "zone") -->

<!-- zone_sample_hh_analysis_size |>  -->
<!--   DT::datatable( -->
<!--     rownames = F, -->
<!--     colnames=c("Zone", "Median household size (post planting survey)", "Max household size (post planting survey)", "Median household size (post harvest survey)", "Max household size (post harvest survey)"), -->
<!--     caption = "Household size per zone", -->
<!--     selection = list(mode = "none"), -->
<!--     editable = FALSE, -->
<!--     options = list( -->
<!--       digits = 0, -->
<!--       scrollX = TRUE, -->
<!--       scrollCollapse = TRUE, -->
<!--       # autoWidth = TRUE, -->
<!--       fixedColumns = list(leftColumns = 1), -->
<!--       # lengthMenu = c(15, 40), -->
<!--       pageLength = 6 -->
<!--     )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| label: tbl-hh_analysis_size_state -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| cache: false -->
<!-- #| results: asis -->
<!-- #| tab-cap: "Households in analysis size per state" -->

<!-- state_sample_hh_analysis_size <- state_sample_hh_analysis_size_pp |> left_join(select(state_sample_hh_analysis_size_ph, -zone), by = "state") -->

<!-- state_sample_hh_analysis_size |>  -->
<!--   DT::datatable( -->
<!--     rownames = F, -->
<!--     colnames=c("Zone", "State", "Median household size (post planting survey)", "Max household size (post planting survey)", "Median household size (post harvest survey)", "Max household size (post harvest survey)"), -->
<!--     caption = "Household size per state", -->
<!--     selection = list(mode = "none"), -->
<!--     editable = FALSE, -->
<!--     options = list( -->
<!--       order = list(list(0, 'asc'), list(2, 'desc')), -->
<!--       digits = 0, -->
<!--       scrollX = TRUE, -->
<!--       scrollCollapse = TRUE, -->
<!--       # autoWidth = TRUE, -->
<!--       fixedColumns = list(leftColumns = 1), -->
<!--       # lengthMenu = c(15, 40), -->
<!--       pageLength = 15 -->
<!--     )) -->
<!-- ``` -->

<!-- ### Extrapolating from the sample to the population -->

<!-- Since our objective is a representative estimate of food consumption in 2023/24 (rather than a longitudinal comparison) we should use the cross-sectional weight `wt_cross_wave5`. NBS explicitly says the cross-sectional weight is calibrated to the 2023 population and is intended for representative current-population estimates. -->

<!-- ```{r} -->
<!-- #| label: GHS_analysis_weights -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| cache: false -->
<!-- #| results: asis -->

<!-- # Check -->

<!-- summary(df_hh_analysis_pp$wt_cross_wave5) -->
<!-- summary(df_hh_analysis_ph$wt_cross_wave5) -->

<!-- # Get zone samples of hh_size -->

<!-- zone_sample_wt_cross_pp <- df_hh_analysis_pp |> -->
<!--   group_by( -->
<!--     zone -->
<!--   ) |> -->
<!--   summarise( -->
<!--     mean_wt_cross_pp = mean(wt_cross_wave5), -->
<!--     .groups = "drop" -->
<!--   ) |> -->
<!--   arrange(zone) -->

<!-- ``` -->

<!-- We will use the cross sectional weight in further calculations. -->


### Food product consumption

There are two datasets that record household-food item level information on the quantity and value of food consumed within the household during the past 7 days. The surveys are representative of the six geopolitical zones, so zonal estimates are an intended use of the data.

For the post-harvest 2023/24 visit, the key file is:

`sect5b_harvestw5.dta` — Section 5B: Food Consumption

It contains 509,220 household-food observations and 26 variables. Each household has a record for the food items in the food roster, whether or not that particular item was consumed.

The corresponding post-planting file is:

`sect6b_plantingw5.dta`

with 529,470 observations, and essentially the same food-consumption structure.

There are over 100 different food products that can be recorded, and as a consequence of our samples of over 4,000 households, there are over half a million records for both the post-planting and post-harvest datasets.

To calculate household food consumption the relevant variables are:

Variable: Post Planting Survey | Variable: Post Harvest Survey | Description | Resposne Data Type
------------------------------|------------------------------|--------------|-----------
`s6bq1` | `s5bq1`  | consumed this item in previous 7 days? | "1. Yes" or "2. No"
`s6bq2a` | `s5bq2a` | quantity consumed | numeric
`s6bq2b` | `s5bq2b` | unit | text
`s6bq2b_os` | `s5bq2b_os` | other unit | text
`s6bq2c` | `s5bq2c` | size | text
`s6bq2_cvn` | `s5bq2_cvn` | conversion factor | numeric


::: {.callout-note title="Calculating the weight of the food product consumed"}
The NBS catalogue explicitly describes `s6bq2_cvn` and `s5bq2_cvn` as "conversion factor item-unit-size-zone", whose values range from 0.001 to 50 in the released data.

Suppose a household reports:

2 baskets of maize

The raw `s5bq2a` value is 2, but that is not 2 kg.

The survey records:

- quantity = 2
- unit = basket
- size = [reported size]
- conversion factor = X kg per reported unit

so we can derive:

quantity consumed in kg per week = `s5bq2a` × `s5bq2_cvn` or `s6bq2a`  x `s6bq2_cvn`

subject to checking the exact direction/definition of the conversion factor in the questionnaire/data documentation.

The last check should be conducted before producing the final numbers rather than assuming the factor's direction.
:::

However, a number of food products items in the surveys do not have a conversion factor (@tbl-GHS_food_consumed_hh_analysis_pp and @tbl-GHS_food_consumed_hh_analysis_ph). The food products with the highest number of missing conversion factors are okra and cassava roots.
The reasons for the missing conversion factors for okra are unclear, because other households with the same unit and size do have a conversion unit.
For cassava, the reason appears to be a mis-match between the principal product name (cassava roots) and the alternative name and unit, e.g. wrap of fufu.
Given these mis-matches we will not attempt to add a conversion unit.


```{r}
#| label: tbl-GHS_food_consumed_hh_analysis_pp
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "Number of consumed food product items without a conversion factor in the post planting survey"


# ----1. post planting household food consumption----

df_GHS_sect6b_plantingw5 <- read.csv(
  "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Planting Wave 5/Household/sect6b_plantingw5.csv")

df_food_pp <- df_GHS_sect6b_plantingw5 |>
  mutate(
    s6bq2a = as.numeric(s6bq2a),
    s6bq2_cvn = as.numeric(s6bq2_cvn),
    item_cd_int = as.numeric(stringi::stri_extract_first_regex(
      item_cd,
      "[0-9]+"
    ))
  ) 

df_food_pp_consumed <- df_food_pp |>
  filter(
    s6bq1 == "1. Yes",
    !is.na(s6bq2a),
    s6bq2a >= 0
  ) |>
  mutate(
    kg_week = s6bq2a * s6bq2_cvn) 


df_missing_conversion_pp <- df_food_pp_consumed |>
  filter(
    is.na(s6bq2_cvn)
  ) |>
  count(
    item_cd_int,
    item_cd,
    item_cd_alt,
    item_os,
    sort = TRUE
  )

df_missing_conversion_pp |> 
  DT::datatable(
    rownames = F,
    colnames=c("Item code", "Food Product Name", "alternative name", "alternative name", "Number without conversion factor"),
    caption = "Post planting missing conversion factors",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(4, 'desc'), list(0, 'asc')),
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 1),
      # lengthMenu = c(15, 40),
      pageLength = 15
    ))

df_food_pp_consumed <- df_food_pp_consumed |> filter(!is.na(kg_week))

df_food_item_pp_consumed <- df_food_pp_consumed |>
  group_by(
    item_cd_int
  ) |>
  summarise(
    mean_kg = mean(kg_week),
    .groups = "drop"
  ) |>
  arrange(item_cd_int)



```

```{r}
#| label: tbl-GHS_food_consumed_hh_analysis_ph
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "Number of consumed food product items without a conversion factor in the post harvest survey"

# ----2. post harvest household food consumption----

df_GHS_sect5b_plantingw5 <- read.csv(
  "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024/Post Harvest Wave 5/Household/sect5b_harvestw5.csv")

df_food_ph <- df_GHS_sect5b_plantingw5 |>
  mutate(
    s5bq2a = as.numeric(s5bq2a),
    s5bq2_cvn = as.numeric(s5bq2_cvn),
    item_cd_int = as.numeric(stringi::stri_extract_first_regex(
      item_cd,
      "[0-9]+"
    ))
  ) 

df_food_ph_consumed <- df_food_ph |>
  filter(
    s5bq1 == "1. Yes",
    !is.na(s5bq2a),
    s5bq2a >= 0
  ) |>
  mutate(
    kg_week = s5bq2a * s5bq2_cvn,
  ) 

df_missing_conversion_ph <- df_food_ph_consumed |>
  filter(
    is.na(s5bq2_cvn)
  ) |>
  count(
    item_cd_int,
    item_cd,
    item_cd_alt,
    item_os,
    sort = TRUE
  )

df_missing_conversion_ph |> 
  DT::datatable(
    rownames = F,
    colnames=c("Item code", "Food Product Name", "alternative name", "alternative name", "Number without conversion factor"),
    caption = "Post harvest missing conversion factors",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(4, 'desc'), list(0, 'asc')),
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      # autoWidth = TRUE,
      fixedColumns = list(leftColumns = 1),
      # lengthMenu = c(15, 40),
      pageLength = 15
    ))

df_food_ph_consumed <- df_food_ph_consumed |>  filter(!is.na(kg_week))

df_food_item_ph_consumed <- df_food_ph_consumed |>
  group_by(
    item_cd_int
  ) |>
  summarise(
    mean_kg = mean(kg_week),
    .groups = "drop"
  ) |>
  arrange(item_cd_int)
```

```{r}
#| label: GHS_food_product_lookup
#| message: false
#| warning: false
#| cache: false
#| results: asis


food_lookup <- tribble(
  ~item_cd_int , ~product                     ,

        10 , "Guinea corn / sorghum"      ,
        11 , "Millet"                     ,
        13 , "Rice - local"               ,
        14 , "Rice - imported"            ,
        16 , "Maize flour"                ,
        17 , "Yam flour"                  ,
        18 , "Cassava flour"              ,
        19 , "Wheat flour"                ,
        20 , "Maize - unshelled/on cob"   ,
        22 , "Maize - shelled/off cob"    ,
        23 , "Other grains/flour"         ,

        25 , "Bread"                      ,
        26 , "Cake"                       ,
        27 , "Buns/Pofpof/Donuts"         ,
        28 , "Biscuits"                   ,
        29 , "Meat pie/Sausage roll"      ,

        30 , "Cassava roots"              ,
        31 , "Yam roots"                  ,
        32 , "Gari - white"               ,
        33 , "Gari - yellow"              ,
        34 , "Cocoyam"                    ,
        35 , "Plantains"                  ,
        36 , "Sweet potatoes"             ,
        37 , "Potatoes"                   ,
        38 , "Other roots/tubers"         ,

        40 , "Soya beans"                 ,
        41 , "Brown beans"                ,
        42 , "White beans"                ,
        43 , "Groundnuts - unshelled"     ,
        44 , "Groundnuts - shelled"       ,
        45 , "Other nuts/seeds/pulses"    ,
        46 , "Coconut"                    ,
        47 , "Kola nut"                   ,
        48 , "Cashew nut"                 ,

        50 , "Palm oil"                   ,
        51 , "Butter/Margarine"           ,
        52 , "Groundnut oil"              ,
        53 , "Other oil/fat"              ,
        54 , "Shea butter"                ,
        56 , "Animal fat"                 ,

        60 , "Bananas"                    ,
        61 , "Orange/Tangerine"           ,
        62 , "Mangoes"                    ,
        63 , "Avocado pear"               ,
        64 , "Pineapples"                 ,
        66 , "Other fruits"               ,
        67 , "Pawpaw"                     ,
        68 , "Watermelon"                 ,
        69 , "Apples"                     ,

        70 , "Tomatoes"                   ,
        71 , "Tomato puree - canned"      ,
        72 , "Onions"                     ,
        73 , "Garden eggs/Eggplant"       ,
        74 , "Okra - fresh"               ,
        75 , "Okra - dried"               ,
        76 , "Fresh pepper"               ,
        77 , "Dry pepper"                 ,
        78 , "Leaves"                     ,
        79 , "Other vegetables"           ,

        80 , "Chicken"                    ,
        81 , "Duck"                       ,
        82 , "Other domestic poultry"     ,
        83 , "Agricultural eggs"          ,
        84 , "Local eggs"                 ,
        85 , "Other eggs"                 ,

        90 , "Beef"                       ,
        91 , "Mutton"                     ,
        92 , "Pork"                       ,
        93 , "Goat"                       ,
        94 , "Wild game/Bush meat"        ,
        95 , "Canned beef"                ,
        96 , "Other meat"                 ,

       100 , "Fish - fresh"               ,
       101 , "Fish - frozen"              ,
       102 , "Fish - smoked"              ,
       103 , "Fish - dried"               ,
       104 , "Snails"                     ,
       105 , "Seafood"                    ,
       106 , "Canned fish/seafood"        ,
       107 , "Other fish/seafood"         ,

       110 , "Fresh milk"                 ,
       111 , "Milk powder"                ,
       112 , "Baby milk powder"           ,
       113 , "Milk - tinned"              ,
       114 , "Cheese/wara"                ,
       115 , "Other milk products"        ,

       120 , "Coffee"                     ,
       121 , "Chocolate drinks"           ,
       122 , "Tea"                        ,

       130 , "Sugar"                      ,
       132 , "Honey"                      ,
       133 , "Other sweets"               ,

       141 , "Salt"                       ,
       142 , "Ogbono - unground"          ,
       143 , "Ogbono - ground"            ,
       144 , "Ground pepper"              ,
       145 , "Melon - shelled"            ,
       146 , "Melon - unshelled"          ,
       147 , "Melon - ground"             ,
       148 , "Other spices"               ,

       150 , "Bottled water"              ,
       151 , "Sachet water"               ,
       152 , "Malt drinks"                ,
       153 , "Soft drinks"                ,
       154 , "Fruit juice"                ,
       155 , "Other non-alcoholic drinks" ,

       160 , "Beer"                       ,
       161 , "Palm wine"                  ,
       162 , "Pito"                       ,
       163 , "Gin"                        ,
       164 , "Other alcoholic beverages"  ,

       601 , "Guava"
)


```

The conversion factor is used to calculate the household consumption of the food product item.
The crop products with the highest weight are cassava roots and yam roots (@tbl-food_kg_week), but it should be remembered that this table only considers the households that actually consumed the food product in the previous seven days, so these values cannot yet be used to calculate the per capita consumption.

```{r}
#| label: tbl-food_kg_week
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "food_kg_week"


df_food_item_consumed <- df_food_item_pp_consumed |>
  merge(df_food_item_ph_consumed, by = "item_cd_int", all = T) |>
left_join(food_lookup, by = "item_cd_int")  |>
  mutate(
    kg_week_pp = prettyunits::pretty_round(mean_kg.x, digits = 2),
    kg_week_ph = prettyunits::pretty_round(mean_kg.y, digits = 2)
  )|>
  select(item_cd_int, product, kg_week_pp, kg_week_ph)

df_food_item_consumed |> 
  DT::datatable(
    rownames = F,
    colnames=c("ID", "Food Product", "Mean kg per household per week (post planting survey)", "Mean kg per household per week (post harvest survey)"),
    caption = "Weekly Food Product Consumption",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(2, 'desc'), list(3, 'asc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```



```{r}
#| label: tbl-food_kg_week_agg
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "food_kg_week_pp"


# ------------------------------------------------------------
# JOIN Product name
# ------------------------------------------------------------

df_food_pp_consumed_standard <- df_food_pp_consumed |>
left_join(food_lookup, by = "item_cd_int") 

df_food_ph_consumed_standard <- df_food_ph_consumed |>
left_join(food_lookup, by = "item_cd_int") 

# ------------------------------------------------------------
# AGGREGATE FOOD TO HOUSEHOLD × PRODUCT
# A household can potentially have more than one observation
# contributing to a product.
# ------------------------------------------------------------

df_food_pp_consumed_standard_hh <- df_food_pp_consumed_standard |>
  group_by(hhid, product) |>
  summarise(kg_week = sum(kg_week, na.rm = TRUE), .groups = "drop")

df_food_ph_consumed_standard_hh <- df_food_ph_consumed_standard |>
  group_by(hhid, product) |>
  summarise(kg_week = sum(kg_week, na.rm = TRUE), .groups = "drop")
# ------------------------------------------------------------

```



2. What food items are actually available?

The standard `item_cd` roster contains the following crop-related items, among others:

Code |	Food item
-----|-----------
10	| Guinea corn/sorghum
11	| 	Millet
13	| 	Rice – local
14	| 	Rice – imported
16	| 	Maize flour
17	| 	Yam flour
18	| 	Cassava flour
20	| 	Maize – unshelled/on the cob
22	| 	Maize – shelled/off the cob
23	| 	Other grains and flour
30	| 	Cassava – roots
31	| 	Yam – roots
32	| 	Gari – white
33	| 	Gari – yellow
34	| 	Cocoyam
35	| 	Plantains
36	| 	Sweet potatoes
37	| 	Potatoes
38	| 	Other roots and tubers
40	| 	Soya beans
41	| 	Brown beans
42	| 	White beans
43	| 	Groundnuts – unshelled
44	| 	Groundnuts – shelled
45	| 	Other nuts/seeds/pulses
46	| 	Coconut
47	| 	Kola nut
48	| 	Cashew nut
50	| 	Palm oil
52	| 	Groundnut oil
60	| 	Bananas
61	| 	Orange/tangerine
62	| 	Mangoes
63	| 	Avocado pear
64	| 	Pineapples
67	| 	Pawpaw
68	| 	Watermelon
70	| 	Tomatoes
72	| 	Onions
73	| 	Garden eggs/eggplant
74	| 	Okra – fresh
75	| 	Okra – dried
76	| 	Fresh pepper
77	| 	Dry pepper
78	| 	Leaves
79	| 	Other vegetables

There are also numerous non-crop foods. The complete catalogue runs from item code 10 through 601, with gaps between food groups.

But there is an important second level

The variable `item_os_cd` captures additional foods when the respondent selected "other item".

This is particularly interesting for our purposes because it contains:

Code |	Food item
-----|-----------
231	|  Semovita/semolina
232	|  Akanu/Pap/Obi/Fufu
234	|  Plantain flour
235	|  Macaroni/pasta/noodles
237	|  Acha/fonio
381	|  Yam – second variety
453	|  Bambara nut
454	|  Palm nut
455	|  Palm kernel
456	|  Locust beans
661	|  Agbalumo/Udara/star apple
662	|  Cashew fruit
790	|  Jute leaves
793	|  Kuka/Baobab
798	|  Pumpkin leaves
1071	|  Crayfish
1152	|  Yoghurt
1331	|  Sugar cane
1551	|  Zobo
1552	|  Kunu
etc.

There were 503 valid `item_os_cd` observations in the post-harvest file.

So fufu isn't missing entirely — it occurs as an "other item", although only two observations in the post-harvest file use the coded category 232. There are also free-text unit descriptions mentioning cooked fufu.

This means we need to retain both `item_cd` and `item_os_cd`/`item_os` when constructing the food list.

3. How the quantity information works

For every household × food item, Section 5B contains:

Variable |	Meaning
---------|-----------
`s5bq1`	 |	Did the household consume this item in the past 7 days?
`s5bq2a` 	 |	Quantity consumed in past 7 days
`s5bq2b`	 |	Unit
`s5bq2b_os`	 |	Other unit, if applicable
`s5bq2c`	 |	Size
`s5bq2_cvn`	 |	Conversion factor: item × unit × size × zone
`s5bq3`	 |	Quantity consumed from purchases
`s5bq4`	 |	Quantity consumed from own production
`s5bq5`	 |	Quantity consumed from gifts/other sources

There is a parallel set for purchases during the previous 30 days:

`s5bq7a`
`s5bq7b`
`s5bq7c`
`s5bq7_cvn`
`s5bq8`


4. We should use consumption, NOT purchases

This distinction is very important for our question.

We want:

average consumption per person per year

Therefore the appropriate variable is:

`s5bq2a` × `s5bq2_cvn`

—not the 30-day purchase variable.

The purchase variables tell us where the food came from / how much was bought.

The survey explicitly separates these sources with `s5bq3`, `s5bq4`, and `s5bq5`.

For total consumption we therefore want:

purchased consumption + own-production consumption + gifts/other consumption

which should correspond to the total reported in `s5bq2a`.




# Results

## Household Food consumption calculation

```{r}
#| label: consumption_diagnostics_function
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ------------------------------------------------------------
# CONSUMPTION DIAGNOSTICS
#
# Calculates, for each geographic unit and food product:
#
# 1. % of households consuming
# 2. % of population living in consuming households
# 3. kg/person/week among consumers
# 4. number of consuming households
# 5. total number of eligible households
#
# geographic variable can be:
#   "state"
#   "zone"
# ------------------------------------------------------------

# lonely.psu = "adjust" prevents failures when a domain

# contains only one PSU within a stratum.

# ------------------------------------------------------------

options(
  survey.lonely.psu = "adjust"
)

make_consumption_diagnostics <- function(
  hh_analysis,
  food_standard,
  survey_period,
  geographic_variable = "state"
) {
  products <- sort(
    unique(food_standard$product)
  )

  results <- vector(
    "list",
    length(products)
  )

  for (i in seq_along(products)) {
    p <- products[i]

    message(
      survey_period,
      " - ",
      geographic_variable,
      " diagnostics: ",
      p,
      " (",
      i,
      "/",
      length(products),
      ")"
    )

    # --------------------------------------------------------
    # ALL ELIGIBLE HOUSEHOLDS
    # --------------------------------------------------------

    hh <- hh_analysis |>
      select(
        hhid,
        zone,
        state,
        cluster,
        strata,
        wt_cross_wave5,
        hh_size
      ) |>

      left_join(
        food_standard |>
          filter(
            product == p
          ) |>
          group_by(
            hhid
          ) |>
          summarise(
            kg_week = sum(
              kg_week,
              na.rm = TRUE
            ),
            .groups = "drop"
          ),
        by = "hhid"
      ) |>

      mutate(
        # A household is a consumer if it has a food record
        # for this product.
        consumed = !is.na(kg_week),

        # Non-consumers have zero consumption.
        kg_week = replace_na(
          kg_week,
          0
        )
      )

    # --------------------------------------------------------
    # SURVEY DESIGN
    # --------------------------------------------------------

    d <- svydesign(
      ids = ~cluster,
      strata = ~strata,
      weights = ~wt_cross_wave5,
      data = hh,
      nest = TRUE
    )

    # --------------------------------------------------------
    # HOUSEHOLD CONSUMPTION PREVALENCE
    # --------------------------------------------------------

    f_hh <- as.formula(
      paste0(
        "~consumed"
      )
    )

    f_geo <- as.formula(
      paste0(
        "~",
        geographic_variable
      )
    )

    hh_prev <- svyby(
      f_hh,
      f_geo,
      d,
      svymean,
      na.rm = TRUE
    )

    # Find the estimate column rather than assuming its name
    estimate_col <- setdiff(
      names(hh_prev),
      geographic_variable
    )[1]

    hh_prev <- hh_prev |>
      rename(
        hh_consumption_prevalence = all_of(estimate_col)
      ) |>
      mutate(
        hh_consumption_percent = 100 * hh_consumption_prevalence
      )

    # --------------------------------------------------------
    # POPULATION CONSUMPTION PREVALENCE
    #
    # A household contributes its household size to the
    # population denominator.
    # --------------------------------------------------------

    population_design <- svydesign(
      ids = ~cluster,
      strata = ~strata,
      weights = ~wt_cross_wave5,
      data = hh,
      nest = TRUE
    )

    # Use a population-weighted numerator and denominator
    # explicitly rather than changing the survey weights.
    hh <- hh |>
      mutate(
        consuming_persons = hh_size * as.numeric(consumed),

        total_persons = hh_size
      )

    d_population <- svydesign(
      ids = ~cluster,
      strata = ~strata,
      weights = ~wt_cross_wave5,
      data = hh,
      nest = TRUE
    )

    pop_ratio <- svyby(
      ~consuming_persons,
      f_geo,
      d_population,
      svytotal,
      na.rm = TRUE
    )

    pop_total <- svyby(
      ~total_persons,
      f_geo,
      d_population,
      svytotal,
      na.rm = TRUE
    )

    pop_prev <- pop_ratio |>
      select(
        all_of(geographic_variable),
        consuming_persons
      ) |>
      left_join(
        pop_total |>
          select(
            all_of(geographic_variable),
            total_persons
          ),
        by = geographic_variable
      ) |>
      mutate(
        population_consumption_prevalence = consuming_persons /
          total_persons,

        population_consumption_percent = 100 *
          population_consumption_prevalence
      )

    # --------------------------------------------------------
    # NUMBER OF CONSUMING HOUSEHOLDS
    #
    # These are UNWEIGHTED sample counts and are useful as a
    # diagnostic for small sample sizes.
    # --------------------------------------------------------

    sample_counts <- hh |>
      group_by(
        .data[[geographic_variable]]
      ) |>
      summarise(
        n_households = n(),

        n_consuming_households = sum(
          consumed,
          na.rm = TRUE
        ),

        .groups = "drop"
      ) |>
      rename(
        geographic_unit = all_of(geographic_variable)
      )

    # --------------------------------------------------------
    # CONSUMPTION AMONG CONSUMERS
    #
    # kg/person/week among households that actually consumed
    # the product.
    # --------------------------------------------------------

    consumer_hh <- hh |>
      filter(
        consumed
      )

    if (nrow(consumer_hh) > 0) {
      d_consumer <- svydesign(
        ids = ~cluster,
        strata = ~strata,
        weights = ~wt_cross_wave5,
        data = consumer_hh,
        nest = TRUE
      )

      consumer_results <- lapply(
        sort(
          unique(
            consumer_hh[[geographic_variable]]
          )
        ),
        function(g) {
          dc <- subset(
            d_consumer,
            get(geographic_variable) == g
          )

          if (nrow(dc) == 0) {
            return(
              tibble(
                geographic_unit = g,
                consumer_kg_person_week = NA_real_,
                consumer_se = NA_real_,
                consumer_cv = NA_real_
              )
            )
          }

          r <- svyratio(
            ~kg_week,
            ~hh_size,
            dc,
            na.rm = TRUE
          )

          estimate <- as.numeric(
            coef(r)
          )

          se <- as.numeric(
            SE(r)
          )

          tibble(
            geographic_unit = g,
            consumer_kg_person_week = estimate,
            consumer_se = se,
            consumer_cv = if_else(
              estimate > 0,
              100 * se / estimate,
              NA_real_
            )
          )
        }
      ) |>
        bind_rows()
    } else {
      consumer_results <- tibble(
        geographic_unit = character(),
        consumer_kg_person_week = numeric(),
        consumer_se = numeric(),
        consumer_cv = numeric()
      )
    }

    # --------------------------------------------------------
    # COMBINE
    # --------------------------------------------------------

    results[[i]] <-
      hh_prev |>

      select(
        all_of(geographic_variable),
        hh_consumption_prevalence,
        hh_consumption_percent
      ) |>

      rename(
        geographic_unit = all_of(geographic_variable)
      ) |>

      left_join(
        pop_prev |>
          select(
            geographic_unit = all_of(geographic_variable),
            population_consumption_prevalence,
            population_consumption_percent
          ),
        by = "geographic_unit"
      ) |>

      left_join(
        consumer_results,
        by = "geographic_unit"
      ) |>

      left_join(
        sample_counts,
        by = "geographic_unit"
      ) |>

      mutate(
        product = p,
        survey_period = survey_period,
        geographic_level = geographic_variable,
        .before = 1
      )
  }

  bind_rows(
    results
  )
}

```

```{r}
#| label: consumption_diagnostics_state
#| message: false
#| warning: false
#| cache: false
#| results: asis
# ------------------------------------------------------------
# 14B. CALCULATE DIAGNOSTICS FOR BOTH VISITS
# ------------------------------------------------------------

df_state_consumption_diagnostics_pp <-
  make_consumption_diagnostics(
    df_hh_analysis_pp,
    df_food_pp_consumed_standard,
    "Post-Harvest"
  )


df_state_consumption_diagnostics_ph <-
  make_consumption_diagnostics(
    df_hh_analysis_ph,
    df_food_ph_consumed_standard,
    "Post-Planting"
  )


df_state_consumption_diagnostics <-
  bind_rows(
    df_state_consumption_diagnostics_pp,
    df_state_consumption_diagnostics_ph
  )

```

As an intermediate step we check on the number of households that consume specific food product items and also assess the population for which these households represent, and the overall average consumption per week.

There are too many products and geographic units to show all records but @tbl-pop_sample_food_kg_week shows the number of households surveyed in Kaduna and Nasarawa states, and the consumption of Millet in both surveys


```{r}
#| label: tbl-pop_sample_food_kg_week
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tab-cap: "pop_sample_food_kg_week"

df_state_consumption_diagnostics_sample <- df_state_consumption_diagnostics |>
  filter(product == "Millet" & (geographic_unit == "18. Kaduna" | geographic_unit == "25. Nasarawa")) |>
  select(-population_consumption_prevalence, -hh_consumption_prevalence, -hh_consumption_percent, -consumer_se, -consumer_cv) |>
    mutate(
      population_consumption_percent = prettyunits::pretty_round(
        population_consumption_percent,
        digits = 2
      ),
      consumer_kg_person_week = prettyunits::pretty_round(
        consumer_kg_person_week,
        digits = 2
      )
    )

df_state_consumption_diagnostics_sample |> 
  DT::datatable(
    rownames = F,
    colnames=c("Food Product", "Survey period","Geographic Level", "Geographic Unit", "Mean kg per capita per week", "households sampled", "households consuming"),
    caption = "Weekly Food Product Consumption",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(2, 'desc'), list(3, 'asc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))



```

We start with ALL households and then attach the food quantity for the selected product.

Therefore:

consumer household -> measured kg/week
non-consumer        -> 0 kg/week

The denominator remains household size for EVERY household.

This avoids the serious bias that would occur if we only
analysed households that consumed the product.


```{r}
#| label: GHS_per_capita_food_consumption_calculation_functions
#| message: false
#| warning: false
#| cache: false
#| results: asis


# 15. FUNCTION FOR ONE PRODUCT
# ------------------------------------------------------------

make_product_design <- function(product_name, hh_analysis, food_hh) {
  product_data <- hh_analysis |>
    select(
      hhid,
      zone,
      state,
      lga,
      sector,
      ea,
      cluster,
      strata,
      wt_cross_wave5,
      hh_size
    ) |>
    left_join(
      food_hh |>
        filter(
          product == product_name
        ) |>
        select(
          hhid,
          kg_week
        ),
      by = "hhid"
    ) |>
    mutate(
      kg_week = replace_na(
        kg_week,
        0
      )
    )

  svydesign(
    ids = ~cluster,
    strata = ~strata,
    weights = ~wt_cross_wave5,
    data = product_data,
    nest = TRUE
  )
}

# ------------------------------------------------------------
# 16. FUNCTION TO EXTRACT RATIO ESTIMATE
#
# Ratio:
#
# weighted total kg/week
# ---------------------
# weighted total persons
#
# This gives kg/person/week.
# ------------------------------------------------------------

extract_ratio <- function(
  design,
  by_variable,
  denominator = ~hh_size
) {
  result <- svyby(
    formula = ~kg_week,
    by = by_variable,
    design = design,
    FUN = svyratio,
    denominator = denominator,
    na.rm = TRUE,
    keep.names = FALSE,
    vartype = c("se", "cv")
  )

  names(result) <- c(
    all.vars(by_variable),
    "estimate_kg_person_week",
    "se",
    "cv"
  )

  result |>
    mutate(
      ci_low = pmax(
        0,
        estimate_kg_person_week - 1.96 * se
      ),
      ci_high = estimate_kg_person_week + 1.96 * se,
      # Annualized equivalent is included only as a convenient
      # comparison scale. It is NOT an observed annual total.
      estimate_kg_person_year_equivalent = estimate_kg_person_week * 365 / 7
    )
}
```


```{r}
#| label: GHS_per_capita_food_consumption_calculation_function_products
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ------------------------------------------------------------
# 17. PRODUCTS TO ANALYSE
# ------------------------------------------------------------

products_pp <- sort(
  unique(
    df_food_pp_consumed_standard_hh$product
  )
)

cat("\n\nPost Planting Survey Number of products:",length(products_pp),"\n")
products_pp

products_ph <- sort(
  unique(
    df_food_ph_consumed_standard_hh$product
  )
)

cat("\n\nPost Harvest Survey Number of products:",length(products_ph),"\n")
products_ph


```


```{r}
#| label: GHS_per_capita_food_consumption_state_calculations
#| message: false
#| warning: false
#| cache: false
#| results: asis


# 18. STATE-LEVEL RESULTS
# ------------------------------------------------------------

# ------------------------------------------------------------
# -------------POST PLANTING------------------
# ------------------------------------------------------------

state_results_list_pp <- vector(
  "list",
  length(products_pp)
)

for (i in seq_along(products_pp)) {
  p_pp <- products_pp[i]

  message(
    "State estimates: ",
    p_pp,
    " (",
    i,
    "/",
    length(products_pp),
    ")"
  )

  d_pp <- make_product_design(p_pp, df_hh_analysis_pp, df_food_pp_consumed_standard_hh)

  result_pp <- extract_ratio(
    d_pp,
    ~state
  )

  result_pp <- extract_ratio(d_pp, ~state) |>
    mutate(survey_period = "Post-Planting", product = p_pp, .before = 1)

  state_results_list_pp[[i]] <- result_pp
}

state_results_pp <- bind_rows(
  state_results_list_pp
)

# ------------------------------------------------------------
# 19. ADD 95% CONFIDENCE INTERVALS
# ------------------------------------------------------------

state_results_pp <- state_results_pp |>
  mutate(
    ci_low = pmax(
      0,
      estimate_kg_person_week -
        1.96 * se
    ),

    ci_high = estimate_kg_person_week +
      1.96 * se
  )

# ------------------------------------------------------------
# 20. ADD STATE SAMPLE INFORMATION
# ------------------------------------------------------------

state_results_pp <- state_results_pp |>
  left_join(
    state_sample_analysis_pp,
    by = "state"
  ) |>
  select(
    state,
    zone,
    survey_period,
    product,
    estimate_kg_person_week,
    se,
    ci_low,
    ci_high,
    cv,
    n_households,
    n_psu,
    weighted_population,
    median_household_size,
    mean_household_size,
    estimate_kg_person_year_equivalent
  ) |>
  arrange(
    state,
    product
  )

# -----------------------------------------------------------
# 21. PRECISION FLAG
#
# These are analytical flags, not NBS thresholds.
# ------------------------------------------------------------

state_results_pp <- state_results_pp |>
  mutate(
    precision_flag = case_when(
      is.na(cv) ~        "No estimate",
        cv < .10 ~        "CV < 10%",
        cv < .20 ~        "CV 10–20%",
        cv < .30 ~        "CV 20–30% — caution",
        TRUE ~        "CV > 30% — low precision"
    )
  )

# ------------------------------------------------------------
# -------------POST HARVEST------------------
# ------------------------------------------------------------

state_results_list_ph <- vector(
  "list",
  length(products_ph)
)

for (i in seq_along(products_ph)) {
  p_ph <- products_ph[i]

  message(
    "State estimates: ",
    p_ph,
    " (",
    i,
    "/",
    length(products_ph),
    ")"
  )

  d_ph <- make_product_design(p_ph, df_hh_analysis_ph, df_food_ph_consumed_standard_hh)

  result_ph <- extract_ratio(
    d_ph,
    ~state
  )

  result_ph <- extract_ratio(d_ph, ~state) |>
    mutate(survey_period = "Post-Harvest", product = p_ph, .before = 1)

  state_results_list_ph[[i]] <- result_ph
}

state_results_ph <- bind_rows(
  state_results_list_ph
)

# ------------------------------------------------------------
# 19. ADD 95% CONFIDENCE INTERVALS
# ------------------------------------------------------------

state_results_ph <- state_results_ph |>
  mutate(
    ci_low = pmax(
      0,
      estimate_kg_person_week -
        1.96 * se
    ),

    ci_high = estimate_kg_person_week +
      1.96 * se
  )

# ------------------------------------------------------------
# 20. ADD STATE SAMPLE INFORMATION
# ------------------------------------------------------------

state_results_ph <- state_results_ph |>
  left_join(
    state_sample_analysis_ph,
    by = "state"
  ) |>
  select(
    state,
    zone,
    survey_period,
    product,
    estimate_kg_person_week,
    se,
    ci_low,
    ci_high,
    cv,
    n_households,
    n_psu,
    weighted_population,
    median_household_size,
    mean_household_size,
    estimate_kg_person_year_equivalent
  ) |>
  arrange(
    state,
    product
  )

# ------------------------------------------------------------
# 21. PRECISION FLAG
#
# These are analytical flags, not NBS thresholds.
# ------------------------------------------------------------

state_results_ph <- state_results_ph |>
  mutate(
    precision_flag = case_when(
      is.na(cv) ~        "No estimate",
        cv < .10 ~        "CV < 10%",
        cv < .20 ~        "CV 10–20%",
        cv < .30 ~        "CV 20–30% — caution",
        TRUE ~        "CV > 30% — low precision"
    )
  )
```


```{r}
#| label: GHS_per_capita_food_consumption_zone_calculations
#| message: false
#| warning: false
#| cache: false
#| results: asis


# ------------------------------------------------------------
# 22. ZONE-LEVEL RESULTS
# ------------------------------------------------------------

# ------------------------------------------------------------
# -------------POST PLANTING------------------
# ------------------------------------------------------------

zone_results_list_pp <- vector(
  "list",
  length(products_pp)
)

for (i in seq_along(products_pp)) {
  p_pp <- products_pp[i]

  message(
    "Zone estimates: ",
    p_pp,
    " (",
    i,
    "/",
    length(products_pp),
    ")"
  )

  d_pp <- make_product_design(p_pp, df_hh_analysis_pp, df_food_pp_consumed_standard_hh) 

  result_pp <- extract_ratio(d_pp, ~zone) |>
    mutate(survey_period = "Post-Planting", product = p_pp, .before = 1)

  zone_results_list_pp[[i]] <- result_pp
}

zone_results_pp <- bind_rows(
  zone_results_list_pp
) |>
  mutate(
    ci_low = pmax(
      0,
      estimate_kg_person_week -
        1.96 * se
    ),
    ci_high = estimate_kg_person_week +
      1.96 * se
  ) |>
  arrange(
    zone,
    product
  )

# ------------------------------------------------------------
# -------------POST HARVEST------------------
# ------------------------------------------------------------

zone_results_list_ph <- vector(
  "list",
  length(products_ph)
)

for (i in seq_along(products_ph)) {
  p_ph <- products_ph[i]

  message(
    "Zone estimates: ",
    p_ph,
    " (",
    i,
    "/",
    length(products_ph),
    ")"
  )

  d_ph <- make_product_design(p_ph, df_hh_analysis_ph, df_food_ph_consumed_standard_hh)

  result_ph <- extract_ratio(d_ph, ~zone) |>
    mutate(survey_period = "Post-Planting", product = p_ph, .before = 1)

  zone_results_list_ph[[i]] <- result_ph
}

zone_results_ph <- bind_rows(
  zone_results_list_ph
) |>
  mutate(
    ci_low = pmax(
      0,
      estimate_kg_person_week -
        1.96 * se
    ),
    ci_high = estimate_kg_person_week +
      1.96 * se
  ) |>
  arrange(
    zone,
    product
  )

```


```{r}
#| label: tbl-GHS_per_capita_food_consumption_check_results_maize
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ------------------------------------------------------------
# 23. CHECK RESULTS
# ------------------------------------------------------------

# state_results_pp
# state_results_ph
# zone_results_pp
# zone_results_ph

# ------------------------------------------------------------
# 24. EXAMPLE:
# MAIZE PRODUCTS
# ------------------------------------------------------------

maize_results_state_pp <- state_results_pp |>
  filter(
    product %in%
      c(
        "Maize flour",
        "Maize - unshelled/on cob",
        "Maize - shelled/off cob"
      )
  )

maize_results_state_pp |> select(product, state, zone, survey_period, estimate_kg_person_week, se, ci_low, ci_high, cv, precision_flag ) |>   
  mutate(
    estimate_kg_person_week = prettyunits::pretty_round(estimate_kg_person_week, digits = 2),
    se = prettyunits::pretty_round(se, digits = 3),
    ci_low = prettyunits::pretty_round(ci_low, digits = 3),
    ci_high = prettyunits::pretty_round(ci_high, digits = 3),
    cv = prettyunits::pretty_round(cv, digits = 3)
  ) |>
  DT::datatable(
    rownames = F,
    colnames=c("State", "Zone", "Survey period", "Food Product","Estimated kg per person per week", "se", "ci low", "ci high", "cv", "precision flag"),
    caption = "Maize Food Product Consumption",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(4, 'desc'), list(5, 'desc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(10, 40),
      pageLength = 10
    ))

```


```{r}
#| label: tbl-GHS_per_capita_food_consumption_check_results_cassava
#| message: false
#| warning: false
#| cache: false
#| results: asis


# ------------------------------------------------------------
# 25. EXAMPLE:
# CASSAVA PRODUCTS
#
# DO NOT add these together yet.
#
# They represent different processed/raw products.
# ------------------------------------------------------------

cassava_results_state_ph <- state_results_ph |>
  filter(
    product %in%
      c(
        "Cassava roots",
        "Cassava flour",
        "Gari - white",
        "Gari - yellow"
      )
  )

cassava_results_state_ph |> select(product, state, zone, survey_period, estimate_kg_person_week, se, ci_low, ci_high, cv, precision_flag ) |>   
  mutate(
    estimate_kg_person_week = prettyunits::pretty_round(estimate_kg_person_week, digits = 2),
    se = prettyunits::pretty_round(se, digits = 3),
    ci_low = prettyunits::pretty_round(ci_low, digits = 3),
    ci_high = prettyunits::pretty_round(ci_high, digits = 3),
    cv = prettyunits::pretty_round(cv, digits = 3)
  ) |>
  DT::datatable(
    rownames = F,
    colnames=c("State", "Zone", "Survey period", "Food Product","Estimated kg per person per week", "se", "ci low", "ci high", "cv", "precision flag"),
    caption = "Cassava Food Product Consumption",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(4, 'desc'), list(5, 'desc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(10, 40),
      pageLength = 10
    ))

```


```{r}
#| label: tbl-GHS_per_capita_food_consumption_precision_summary_pp
#| message: false
#| warning: false
#| cache: false
#| results: asis


# ------------------------------------------------------------
# 26. STATE PRECISION SUMMARY
# ------------------------------------------------------------

# ------------------------------------------------------------
# -------------POST PLANTING------------------
# ------------------------------------------------------------


state_precision_summary_pp <- state_results_pp |>
  group_by(
    state
  ) |>
  summarise(
    n_products = n(),

    n_cv_gt_30 = sum(
      cv > 0.3,
      na.rm = TRUE
    ),

    median_cv = median(
      cv,
      na.rm = TRUE
    ),

    .groups = "drop"
  ) |>
  arrange(
    desc(n_cv_gt_30)
  )

state_precision_summary_pp |> 
  mutate( median_cv = prettyunits::pretty_round(median_cv, digits = 2)) |> 
  DT::datatable(
    rownames = F,
    colnames=c("State", "Number of Products","Number of Products CV > 30%", "Median CV"),
    caption = "State Precision Summary - Post Planting survey",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(2, 'desc'), list(3, 'desc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```


```{r}
#| label: tbl-GHS_per_capita_food_consumption_precision_summary_ph
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| 

# ------------------------------------------------------------
# -------------POST HARVEST------------------
# ------------------------------------------------------------

state_precision_summary_ph <- state_results_ph |>
  group_by(
    state
  ) |>
  summarise(
    n_products = n(),

    n_cv_gt_30 = sum(
      cv > 0.3,
      na.rm = TRUE
    ),

    median_cv = median(
      cv,
      na.rm = TRUE
    ),

    .groups = "drop"
  ) |>
  arrange(
    desc(n_cv_gt_30)
  )

state_precision_summary_ph |> 
  mutate( median_cv = prettyunits::pretty_round(median_cv, digits = 2)) |> 
  DT::datatable(
    rownames = F,
    colnames=c("State", "Number of Products","Number of Products CV > 30%", "Median CV"),
    caption = "State Precision Summary - Post Harvest survey",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(2, 'desc'), list(3, 'desc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

```{r}
#| label: GHS_per_capita_food_consumption_seasonal_comparison
#| message: false
#| warning: false
#| cache: false
#| results: asis

# ------------------------------------------------------------
# 28. COMBINE THE TWO SEASONAL OBSERVATIONS
#
# To make the two visits directly comparable, the post-harvest
# annual estimate is expressed back as kg/person/week.
# The original annual post-harvest output is left unchanged.
# ------------------------------------------------------------

state_seasonal_results <- bind_rows(
  state_results_pp |>
    transmute(
      survey_period,
      state,
      zone,
      product,
      estimate_kg_person_week,
      se,
      cv,
      ci_low,
      ci_high,
      estimate_kg_person_year_equivalent,
      precision_flag,
      n_households,
      n_psu,
      weighted_population,
      median_household_size
    ),

  state_results_ph |>
    transmute(
      survey_period = "Post-Harvest",
      state,
      zone,
      product,
      estimate_kg_person_week,
      se,
      cv,
      ci_low ,
      ci_high ,
      estimate_kg_person_year_equivalent,
      precision_flag,
      n_households,
      n_psu,
      weighted_population,
      median_household_size
    )
) |>
  arrange(
    state,
    product,
    survey_period
  )

zone_seasonal_results <- bind_rows(
  zone_results_pp |>
    transmute(
      survey_period,
      zone,
      product,
      estimate_kg_person_week,
      se,
      cv,
      ci_low,
      ci_high,
      estimate_kg_person_year_equivalent
    ),

  zone_results_ph |>
    transmute(
      survey_period = "Post-Harvest",
      zone,
      product,
      estimate_kg_person_week,
      se ,
      cv,
      ci_low ,
      ci_high ,
      estimate_kg_person_year_equivalent
    )
) |>
  arrange(
    zone,
    product,
    survey_period
  )

# Direct state-level seasonal comparison
state_seasonal_comparison <- state_seasonal_results |>
  select(
    survey_period,
    state,
    zone,
    product,
    estimate_kg_person_week
  ) |>
  pivot_wider(
    names_from = survey_period,
    values_from = estimate_kg_person_week
  ) |>
  mutate(
    difference_post_harvest_minus_post_planting = `Post-Harvest` -
      `Post-Planting`,
    ratio_post_harvest_to_post_planting = if_else(
      `Post-Planting` > 0,
      `Post-Harvest` / `Post-Planting`,
      NA_real_
    )
  ) |>
  arrange(
    state,
    product
  )

# Direct zone-level seasonal comparison
zone_seasonal_comparison <- zone_seasonal_results |>
  select(
    survey_period,
    zone,
    product,
    estimate_kg_person_week
  ) |>
  pivot_wider(
    names_from = survey_period,
    values_from = estimate_kg_person_week
  ) |>
  mutate(
    difference_post_harvest_minus_post_planting = `Post-Harvest` -
      `Post-Planting`,
    ratio_post_harvest_to_post_planting = if_else(
      `Post-Planting` > 0,
      `Post-Harvest` / `Post-Planting`,
      NA_real_
    )
  ) |>
  arrange(
    zone,
    product
  )

```


```{r}
#| label: GHS_per_capita_food_consumption_save_data
#| message: false
#| warning: false
#| cache: false
#| results: asis


# ------------------------------------------------------------
# 29. SAVE OUTPUTS
# ------------------------------------------------------------

# All files

data_dir <- "D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/LSMS-GHS 2023-2024"

output_dir <- file.path(
  data_dir,
  "outputs"
)

dir.create(
  output_dir,
  showWarnings = FALSE,
  recursive = TRUE
)

#----post-planting----

# State × product

write_csv(
  state_results_pp,
  file.path(
    output_dir,
    "GHS_wave5_food_consumption_state_planting.csv"
  )
)

# Zone × product

write_csv(
  zone_results_pp,
  file.path(
    output_dir,
    "GHS_wave5_food_consumption_zone_planting.csv"
  )
)

# State sample information

write_csv(
  state_sample_pp,
  file.path(
    output_dir,
    "GHS_wave5_state_sample_diagnostics_planting.csv"
  )
)

# Precision summary

write_csv(
  state_precision_summary_pp,
  file.path(
    output_dir,
    "GHS_wave5_state_precision_summary_planting.csv"
  )
)

# # Alternative codes
# 
# write_csv(
#   alternative_codes_planting,
#   file.path(
#     output_dir,
#     "GHS_wave5_alternative_food_codes_planting.csv"
#   )
# )

# Missing conversion factors

write_csv(
  df_missing_conversion_pp,
  file.path(
    output_dir,
    "GHS_wave5_missing_conversion_factors_planting.csv"
  )
)

#----post-harvest----

# State × product

write_csv(
  state_results_ph,
  file.path(
    output_dir,
    "GHS_wave5_food_consumption_state_harvest.csv"
  )
)

# Zone × product

write_csv(
  zone_results_ph,
  file.path(
    output_dir,
    "GHS_wave5_food_consumption_zone_harvest.csv"
  )
)

# State sample information

write_csv(
  state_sample_ph,
  file.path(
    output_dir,
    "GHS_wave5_state_sample_diagnostics_harvest.csv"
  )
)

# Precision summary

write_csv(
  state_precision_summary_ph,
  file.path(
    output_dir,
    "GHS_wave5_state_precision_summary_harvest.csv"
  )
)

# # Alternative codes
# 
# write_csv(
#   alternative_codes_planting,
#   file.path(
#     output_dir,
#     "GHS_wave5_alternative_food_codes_planting.csv"
#   )
# )

# Missing conversion factors

write_csv(
  df_missing_conversion_ph,
  file.path(
    output_dir,
    "GHS_wave5_missing_conversion_factors_planting.csv"
  )
)
#-----Comparison-----

# State × product

write_csv(
  state_seasonal_results,
  file.path(
    output_dir,
    "GHS_wave5_food_consumption_state_seasonal.csv"
  )
)

write_csv(
  state_seasonal_comparison,
  file.path(
    output_dir,
    "GHS_wave5_food_consumption_state_seasonal_comparison.csv"
  )
)

# Zone × product

write_csv(
  zone_seasonal_results,
  file.path(
    output_dir,
    "GHS_wave5_food_consumption_zone_seasonal.csv"
  )
)

write_csv(
  zone_seasonal_comparison,
  file.path(
    output_dir,
    "GHS_wave5_food_consumption_zone_seasonal_comparison.csv"
  )
)

```

## Conversion to Crop consumption


```{r}
#| label: GHS_crop_conversion_load_data
#| message: false
#| warning: false
#| cache: false
#| results: asis


# ============================================================
# 2. INPUT FILES
# ============================================================
#
# Change these paths if necessary.
#
# If the CSV files are in the same directory as this script,
# the filenames alone are sufficient.
#
# ============================================================

state_file <- "tab_data/GHS food per capita outputs/GHS_wave5_food_consumption_state_seasonal_comparison.csv"

zone_file <- "tab_data/GHS food per capita outputs/GHS_wave5_food_consumption_zone_seasonal_comparison.csv"


# ============================================================
# 3. LOAD STATE-LEVEL RESULTS
# ============================================================

df_food_state_results_seasonal <- read_csv(
  state_file,
  show_col_types = FALSE
)

# convert to kg per year

df_food_state_results <- df_food_state_results_seasonal |>
  mutate(
    estimate_kg_person_year = (((`Post-Harvest` + `Post-Planting`) / 2) *
      (365 / 7))
  )


# ============================================================
# 4. LOAD ZONE-LEVEL RESULTS
# ============================================================

df_food_zone_results_seasonal <- read_csv(
  zone_file,
  show_col_types = FALSE
)

# convert to kg per year

df_food_zone_results <- df_food_zone_results_seasonal |>
  mutate(
    estimate_kg_person_year = (((`Post-Harvest` + `Post-Planting`) / 2) *
      (365 / 7))
  )


# ============================================================
# 5. CHECK INPUT FILES
# ============================================================

required_columns <- c(
  "product",
  "estimate_kg_person_year"
)


missing_state_columns <-
  setdiff(
    required_columns,
    names(df_food_state_results)
  )

missing_zone_columns <-
  setdiff(
    required_columns,
    names(df_food_zone_results)
  )


if (length(missing_state_columns) > 0) {
  stop(
    "The state-level file is missing: ",
    paste(
      missing_state_columns,
      collapse = ", "
    )
  )
}


if (length(missing_zone_columns) > 0) {
  stop(
    "The zone-level file is missing: ",
    paste(
      missing_zone_columns,
      collapse = ", "
    )
  )
}

```

```{r}
#| label: tbl-GHS_crop_conversion_lookup_table
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tbl-cap: "Crop Conversion Lookup"

# ============================================================
# 6. CROP CONVERSION LOOKUP
# ============================================================
#
# factor_mid = principal factor used for the analysis
#
# factor_low / factor_high =
# approximate sensitivity range where appropriate
#
# factor = crop-equivalent kg / food-product kg
#
# For a direct crop:
#
#       1 kg maize grain
#       =
#       1 kg maize crop product
#
# therefore factor = 1.00.
#
# ============================================================

conversion_lookup <- tribble(
  ~product, ~crop, ~factor_low , ~factor_mid , ~factor_high , ~status, ~basis, ~source, ~source_url , ~notes,
  "Guinea corn / sorghum", "Sorghum", 1.00, 1.00, 1.00,"provisional" , "direct grain",
  "GHS item: Guinea corn / sorghum", NA_character_ ,"Direct grain product."                       ,
  "Millet" , "Millet", 1.00, 1.00, 1.00, "provisional", "direct grain" ,
  "GHS item: Millet",  NA_character_ ,"Direct grain product."                                     ,
  "Rice - local", "Rice", 1.00, 1.00 , 1.00, "provisional", "direct milled product",
  "GHS item: Rice - local", NA_character_ , "Treated as rice as consumed."                        ,
  "Rice - imported" , "Rice", 1.00, 1.00, 1.00, "provisional", "direct milled product",
  "GHS item: Rice - imported", NA_character_ , "Treated as rice as consumed."                    ,
  "Maize flour", "Maize", 1.05, 1.14, 1.25, "provisional", "maize flour extraction",
  "FAO — Quantification of Root Crops in National Food Balance Sheets",
  "https://www.fao.org/4/y9422e/y9422e04.htm",  "Maize meal extraction approximately 80–95%."      ,
  "Maize - unshelled/on cob", "Maize", 0.75, 0.8, 0.85, "provisional", "on-cob to grain conversion", "Omnicalculator; SeedCo Zambia" ,
  "https://www.omnicalculator.com/biology/grain-conversion ; https://www.google.com/url?sa=i&source=web&rct=j&url=/goto?url%3DCAESaQHrOzAVLcavph2gJWT-NV7fTKMrDluyjBrI2pyPhWp3IandGoIPqjS2sYekF9De8QtlOVlzv2Xe2bvni7TGH_uPHvmLA2_l9vjUACTUnbdFNVwj9uE-hg-TJs6a_5P8h8sXO2dDdv1z4g&ved=2ahUKEwiMhOvYnOWWAxWUmYQIHTnDHR8QqYcPegoIAggACAAIGBAC&opi=89978449&cd&psig=AOvVaw3eZ4Hlxm_4fQiMTrWARoRx&ust=1789171319579000", "Uses shelling/grain recovery factor."       ,
 "Maize - shelled/off cob", "Maize", 1.00, 1.00, 1.00, "provisional", "direct grain",
  "GHS item: Maize - shelled/off cob", NA_character_ , "Direct maize grain."                     ,
 "Yam flour" , "Yam", 3.33, 3.63, 4, "provisional", "yam flour conversion",  
 "YouTube recipe for making flour", 
  "https://www.google.com/url?sa=i&source=web&rct=j&url=/goto?url%3DCAESYwHrOzAVZUTP58XwFDZod6mRyVmIcjQ6GoaeDmZYcQyomnpy1wrq-k2fLaiu5RWI5dkz_jJLG6MrYywXLunhFPKiEn052A7Bwkr7rjg9xfxIPFgtl9ECpcAGaLDNq1osZQjDfg&ved=2ahUKEwjorOuon-WWAxXRfTABHTXnFv8QqYcPegoIAggACAAIIhAC&opi=89978449&cd&psig=AOvVaw0Zrk8R1FSd2vh1SOUSvEOX&ust=1789172024227000", "Uses conversion from yam flour to fresh yam equivalent."
 ,
  "Cassava flour", "Cassava", 3.75, 4, 4.25, "provisional", "cassava flour conversion",
  "J Aristizabal  2017 ",
  "https://www.google.com/url?sa=i&source=web&rct=j&url=/goto?url%3DCAESigEB6zswFUvEIp_eTOLnYnvzde0i8HcEv4OJeq1WhLLblzOYqSyw5D42rQpZpyo2_a4w_MKmO6HplUkj9Y5vDOorkS_7W3waZwHECZZpyZbggtDMrm-IvmH8gj6hugcmLnrrsoLif5AsdFi2Hai-68CigySlygguhSosMcgWLRRp2TdgQAUUVj9xop0&ved=2ahUKEwjHyqW9oOWWAxXAgoQIHX85MMUQqYcPegoIAggACAAIFhAC&opi=89978449&cd&psig=AOvVaw2R1uvRmx7M9PlmXlUonnkT&ust=1789172335560000",  "Uses conversion from cassava flour to appropriate cassava basis."                            ,
  "Cassava roots", "Cassava", 1.00, 1.00, 1.00, "provisional", "direct roots",
  "GHS item: Cassava roots", NA_character_ , "Check GHS free-text units for prepared fufu/wrapped products."                                                                                        ,
  "Yam roots", "Yam", 1.00 , 1.00 , 1.00 , "provisional", "direct roots",
  "GHS item: Yam roots", NA_character_ , "Direct yam roots."                                      ,
  "Gari - white", "Cassava", 3.5, 4.5, 6.0, "provisional", "gari to fresh cassava",
  "FAO — cassava processing literature", "https://www.fao.org/4/x5458e/x5458e05.htm" ,
  "Highly variable processing yield."                                                              ,
  "Gari - yellow", "Cassava", 3.5, 4.5, 6.0, "provisional", "gari to fresh cassava",
  "FAO — cassava processing literature",
  "https://www.fao.org/4/x5458e/x5458e05.htm",
  "Highly variable processing yield."                                                               ,
  "Cocoyam", "Cocoyam", 1.00, 1.00, 1.00, "provisional", "direct roots",
  "GHS item: Cocoyam", NA_character_ , "Direct crop product."                                      ,
  "Plantains", "Plantain", 1.00, 1.00, 1.00, "provisional", "direct crop product",
  "GHS item: Plantains", NA_character_ , "Direct crop product."                                     ,
  "Sweet potatoes", "Sweet potato", 1.00, 1.00, 1.00, "provisional", "direct roots",
  "GHS item: Sweet potatoes", NA_character_  , "Direct crop product."                              ,
  "Potatoes", "Potato", 1.00, 1.00, 1.00, "provisional", "direct roots",
  "GHS item: Potatoes", NA_character_ , "Direct crop product."                                     ,
  "Soya beans" , "Soybean", 1.00, 1.00, 1.00, "provisional", "direct grain",
  "GHS item: Soya beans", NA_character_ , "Direct soybean grain."                                  ,
  "Brown beans", "Cowpea", 1.00, 1.00, 1.00, "provisional", "direct dry grain",
  "GHS item: Brown beans", NA_character_ , "Brown beans treated as cowpea."                        ,
  "White beans", "Cowpea", 1.00, 1.00, 1.00, "provisional", "direct dry grain",
  "GHS item: White beans", NA_character_ , "White beans treated as cowpea."                        ,
  "Groundnuts - unshelled", "Groundnut", 0.65, 0.7, 0.75, "research", "unshelled to shelled conversion"  , "Nzambani Agrifarms",
  "https://www.google.com/url?sa=i&source=web&rct=j&url=/goto?url%3DCAESzAEB6zswFe6qsuL50Ak61sGftkrGIMYIVr54zPI-HAofFPziIemt0JWX3POqA6VJZW96ySFklK2p4XVkWSAVHSZ6oO9FAeo95RMDDy0_iBcD1aLk_iKWcrQyzK81oIqzylBdQySItDxcfwfXwg1UxGRJaoULSLdQREL6_XRoq7nlct9Q20drUtZIRwj0N7UNr0ljpxMeekGfpsEJTLdQneHM3XiF_1eB1YqamBIua4jWQCwv33x9iyFs-_Tz31NvpqTCAle01rktfq0T7xY&ved=2ahUKEwid88n7oeWWAxWvTDABHTrLBKMQqYcPegoIAggACAAIHhAC&opi=89978449&cd&psig=AOvVaw0jqGOvF_ai0Vgg6rLT2nQP&ust=1789172734614000" ,
  "Uses shelling/outturn factor."                                                                ,
  "Groundnuts - shelled", "Groundnut", 1.00, 1.00, 1.00, "provisional" , "direct kernel",
  "GHS item: Groundnuts - shelled", NA_character_ , "Direct groundnut kernel."                   ,
  "Palm oil", "Oil palm", NA, NA, NA, "research", "palm oil to crop conversion",
  "Further literature required", NA_character_ , "Need to define crop endpoint."                 ,
  "Groundnut oil", "Groundnut", NA, NA, NA, "research", "groundnut oil to crop conversion" ,
  "Further literature required", NA_character_ , "Requires oil extraction rate."                  ,
  "Wheat flour", "Wheat", 1.18 , 1.32 , 1.39 , "provisional", "wheat flour extraction" ,
  "FAO — Cereals", "https://www.fao.org/4/X5557E/x5557e04.htm" ,
  "Wheat flour extraction approximately 72–85%."                                                  ,
  "Bread", "Wheat", NA, NA, NA, "research", "bread recipe",
  "Further literature required", NA_character_ , "Requires wheat flour proportion."                ,
  "Cake", "Wheat", NA, NA, NA, "research", "cake recipe",
  "Further literature required", NA_character_ , "Requires wheat flour proportion."               ,
  "Buns/Pofpof/Donuts", "Wheat", NA, NA, NA, "research", "recipe conversion",
  "Further literature required", NA_character_ , "Requires wheat flour proportion."                ,
  "Biscuits", "Wheat", NA, NA, NA, "research", "recipe conversion",
  "Further literature required", NA_character_ , "Requires wheat flour proportion."               ,
  "Meat pie/Sausage roll", "Wheat" , NA, NA, NA, "research", "recipe conversion",
  "Further literature required", NA_character_ , "Requires wheat flour proportion."               ,
  "Tomato puree - canned", "Tomato", NA, NA, NA, "research", "tomato puree conversion",
  "Further literature required", NA_character_ , "Requires tomato solids/yield conversion."      ,
  "Sugar", "Sugar", NA, NA, NA, "research", "sugar to sugarcane", 
 "Further literature required", NA_character_ , "Requires sugar recovery from sugarcane."
)

# ============================================================
# 7. SAVE CONVERSION LOOKUP
# ============================================================
#
# The lookup is effectively a research
# worksheet. We can edit/add factors as literature is found.
#
# ============================================================

write_csv(
  conversion_lookup,
  "tab_data/GHS food per capita outputs/conversion_lookup.csv"
)

conversion_lookup |> select(product, crop, factor_low , factor_mid , factor_high , status, basis, notes, source, source_url) |>
  DT::datatable(
    rownames = F,
    colnames = c("Food product", "Crop","Conversion factor (low)", "Conversion factor (mid)", "Conversion factor (high)", "Status", "Basis", "Notes", "Source", "Source url"),
    caption = "Food Product to Crop conversion lookup",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(1, 'desc'), list(0, 'desc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(10, 40),
      pageLength = 10
    ))

```

Not all food products are relevant for the Biofortification Prioritisation Index.
These products do not currently have a conversion lookup.

```{r}
#| label: tbl-GHS_crop_conversion_state_missing
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tbl-cap: "State Missing Crop Conversion Lookup"

# ============================================================
# 8a. CHECK FOR FOOD ITEMS WITH NO LOOKUP
# ============================================================

df_state_items <- df_food_state_results |>
  distinct(product)


df_missing_state_lookup <- df_state_items |>
  anti_join(conversion_lookup |>
      select(product),by = "product")

if (nrow(df_missing_state_lookup) > 0) {
 df_missing_state_lookup  |> 
  DT::datatable(
    rownames = F,
    colnames = c("Food product"),
    caption = "State Results — items without a conversion lookup",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(0, 'asc'), list(0, 'asc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(10, 40),
      pageLength = 10
    ))
}

```

```{r}
#| label: tbl-GHS_crop_conversion_zone_missing
#| message: false
#| warning: false
#| cache: false
#| results: asis
#| tbl-cap: "Zone Crop Missing Conversion Lookup"

# ============================================================
# 8b. CHECK FOR FOOD ITEMS WITH NO LOOKUP
# ============================================================

df_zone_items <- df_food_zone_results |>
  distinct(product)

df_missing_zone_lookup <- df_zone_items |>
  anti_join(conversion_lookup |>
      select(product),by = "product"
  )

if (nrow(df_missing_zone_lookup) > 0) {
  
 df_missing_zone_lookup  |> 
  DT::datatable(
    rownames = F,
    colnames = c("Food product"),
    caption = "Zone Results — items without a conversion lookup",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      order = list(list(0, 'asc'), list(0, 'asc')),
      digits = 2,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      #autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(10, 40),
      pageLength = 10
    ))
}
```


```{r}

# ============================================================
# 9. STATE-LEVEL CONVERSION
# ============================================================

df_food_state_crop <- df_food_state_results %>%

  left_join(
    conversion_lookup,
    by = "product"
  ) %>%

  mutate(
    crop_equivalent_low = estimate_kg_person_year * factor_low,
    crop_equivalent_mid = estimate_kg_person_year * factor_mid,
    crop_equivalent_high = estimate_kg_person_year * factor_high
  )

df_crop_state_results <- df_food_state_crop %>%

  filter(
    !is.na(crop),
    !is.na(factor_mid)
  ) %>%

  group_by(
    state,
    crop
  ) %>%

  summarise(
    crop_equivalent_low = sum(crop_equivalent_low, na.rm = TRUE),
    crop_equivalent_kg_person_year = sum(crop_equivalent_mid, na.rm = TRUE),
    crop_equivalent_high = sum(crop_equivalent_high, na.rm = TRUE),
    .groups = "drop"
  )

write_csv(
  df_crop_state_results,
  "tab_data/GHS food per capita outputs/crop_state_results.csv"
)
```

```{r}
# ============================================================
# 10. ZONE-LEVEL CONVERSION
# ============================================================

df_food_zone_crop <- df_food_zone_results %>%

  left_join(
    conversion_lookup,
    by = "product"
  ) %>%

  mutate(
    crop_equivalent_low = estimate_kg_person_year * factor_low,
    crop_equivalent_mid = estimate_kg_person_year * factor_mid,
    crop_equivalent_high = estimate_kg_person_year * factor_high
  )


df_crop_zone_results <- df_food_zone_crop %>%

  filter(
    !is.na(crop),
    !is.na(factor_mid)
  ) %>%

  group_by(
    zone,
    crop
  ) %>%

  summarise(
    crop_equivalent_low = sum(crop_equivalent_low, na.rm = TRUE),
    crop_equivalent_kg_person_year = sum(crop_equivalent_mid, na.rm = TRUE),
    crop_equivalent_high = sum(crop_equivalent_high, na.rm = TRUE),

    .groups = "drop"
  )

write_csv(
  df_crop_zone_results,
  "tab_data/GHS food per capita outputs/crop_zone_results.csv"
)
```








## Map the crop consumption estimates

### Load the States

```{r}
#| label: states
#| message: false
#| warning: false
#| cache: false
#| results: asis

vect_state_geo <- vect("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NGA Admin 2019-2025 OSGOF-Ehealth-UNCS/nga_admin1_topo.topojson")

crs(vect_state_geo) <- "epsg:4326"

vect_state_geo$adm1_pcode <- as.factor(vect_state_geo$adm1_pcode)

# levels(vect_state_geo$adm1_pcode) # Check levels

ggplot(vect_state_geo)+
  geom_spatvector(aes(fill = adm1_pcode)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 3) +
  labs(
    title = "Nigeria States",
    fill = "State<br>Code") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

df_state <- as.data.frame(vect_state_geo)


```

### Load the geopolitical zones

The geopolitical zones are a combination of states with similar ethnic groups, and/or common political history.

There are six zones:

+---------------+----------------------------------------------------------------------------------------+
| Zone name     | States                                                                                 |
+===============+========================================================================================+
| North Central | Benue, Kogi, Kwara, Nasarawa, Niger, and Plateau States, and Federal Capital Territory |
+---------------+----------------------------------------------------------------------------------------+
| North East    | Adamawa, Bauchi, Borno, Gombe, Taraba, and Yobe                                        |
+---------------+----------------------------------------------------------------------------------------+
| North West    | Jigawa, Kaduna, Kano, Katsina, Kebbi, Sokoto, and Zamfara                              |
+---------------+----------------------------------------------------------------------------------------+
| South East    | Abia, Anambra, Ebonyi, Enugu, and Imo                                                  |
+---------------+----------------------------------------------------------------------------------------+
| South South   | Akwa Ibom, Bayelsa, Cross River, Delta, Edo, and Rivers                                |
+---------------+----------------------------------------------------------------------------------------+
| South West    | Ekiti, Lagos, Ogun, Ondo, Osun, and Oyo                                                |
+---------------+----------------------------------------------------------------------------------------+

The zones are created by joining the states.

```{r}
#| label: create_zones
#| message: false
#| warning: false

adm1_pcode <- c("NG001", "NG002", "NG003", "NG004", "NG005", "NG006", "NG007", "NG008", "NG009", "NG010", "NG011", "NG012", "NG013", "NG014", "NG015", "NG016", "NG017", "NG018", "NG019", "NG020", "NG021", "NG022", "NG023", "NG024", "NG025", "NG026", "NG027", "NG028", "NG029", "NG030", "NG031", "NG032", "NG033", "NG034", "NG035", "NG036", "NG037")

geopolitical_zone<- c("South East", "North East", "South South", "South East", "North East", "South South", "North Central", "North East", "South South", "South South", "South East", "South South", "South West", "South East", "North Central", "North East", "South East", "North West", "North West", "North West", "North West", "North West", "North Central", "North Central", "South West", "North Central", "North Central", "South West", "South West", "South West", "South West", "North Central", "South South", "North West", "North East", "North East", "North West")

df_gpz <- tibble(adm1_pcode, geopolitical_zone)

vect_gpz_geo <- dplyr::left_join(vect_state_geo, df_gpz, by = c("adm1_pcode")) |> aggregate(by = "geopolitical_zone")

ggplot(vect_gpz_geo)+
  geom_spatvector(aes(fill = geopolitical_zone)) +
  geom_spatvector_text(aes(label = geopolitical_zone), size = 3) +
  theme_minimal()

```

### Load the consumption estimates

```{r}
#| label: GHS_crop_consumption_data
#| message: false
#| warning: false
#| cache: false
#| results: asis

df_GHS_crop_state <- read.csv("tab_data/GHS food per capita outputs/crop_state_results.csv")
df_GHS_crop_state <- df_GHS_crop_state |>
  mutate(state = trimws(gsub("[^A-Za-z ]", "", state)))

# change name to match states dataset
df_GHS_crop_state['state'][df_GHS_crop_state['state'] == "FCT"] <- "Federal Capital Territory"

df_GHS_crop_state$state <- as.factor(df_GHS_crop_state$state) 

cat(paste(levels(df_GHS_crop_state$state), "; ")) # Check levels

df_GHS_crop_zone <- read.csv("tab_data/GHS food per capita outputs/crop_zone_results.csv")

df_GHS_crop_zone <- df_GHS_crop_zone |>
  mutate(zone = trimws(gsub("[^A-Za-z ]", "", zone)))

df_GHS_crop_zone$zone <- as.factor(df_GHS_crop_zone$zone) 

cat(paste(levels(df_GHS_crop_zone$zone), "; ")) # Check levels

```


### Join to States dataset

```{r}
#| label: GHS_crop_consumption_state_join
#| message: false
#| warning: false
#| cache: false
#| results: asis

vect_state_geo_GHS_crop <- dplyr::left_join(vect_state_geo, df_GHS_crop_state, by = c("adm1_ref_name" = "state"))

vect_state_geo_GHS_crop |> as.data.frame() |> 
  DT::datatable(
    rownames = F,
    caption = "GHS crop consumption zone summary",
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

### Join to Zone dataset

```{r}
#| label: GHS_crop_consumption_zone_join
#| message: false
#| warning: false
#| cache: false
#| results: asis

vect_gpz_geo_GHS_crop <- dplyr::left_join(vect_gpz_geo, df_GHS_crop_zone, by = c("geopolitical_zone" = "zone"))

vect_gpz_geo_GHS_crop |> as.data.frame() |> 
  DT::datatable(
    rownames = F,
    caption = "GHS crop consumption zone summary",
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

### Plot the States results

#### Cassava consumption

```{r}
#| label: GHS_2023_24_state_cassava_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Cassava")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 400),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cassava kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cocoyam (Taro) consumption

```{r}
#| label: GHS_2023_24_state_cocoyam_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Cocoyam")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 10),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cocoyam kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cowpea consumption

```{r}
#| label: GHS_2023_24_state_cowpea_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Cowpea")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 30),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cowpea kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```



#### Maize consumption

```{r}
#| label: GHS_2023_24_state_maize_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Maize")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 50),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Maize kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Millet consumption

```{r}
#| label: GHS_2023_24_state_millet_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Millet")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Millet kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Potato consumption

```{r}
#| label: GHS_2023_24_state_potato_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Potato")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 10),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Potato kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Plantain consumption

```{r}
#| label: GHS_2023_24_state_plantain_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Plantain")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 40),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Plantain kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )


```

#### Rice consumption

```{r}
#| label: GHS_2023_24_state_rice_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Rice")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 50),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Rice kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Sorghum consumption

```{r}
#| label: GHS_2023_24_state_sorghum_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Sorghum")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sorghum kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Sweet Potato consumption

```{r}
#| label: GHS_2023_24_state_sweet_potato_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_state_geo_GHS_crop, crop == "Sweet potato")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 15),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sweet Potato kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```
### Plot the Zones results

#### Cassava consumption

```{r}
#| label: GHS_2023_24_zone_cassava_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Cassava")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 400),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cassava kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cocoyam (Taro) consumption

```{r}
#| label: GHS_2023_24_zone_cocoyam_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Cocoyam")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 10),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cocoyam kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Cowpea consumption

```{r}
#| label: GHS_2023_24_zone_cowpea_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Cowpea")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 30),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cowpea kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```



#### Maize consumption

```{r}
#| label: GHS_2023_24_zone_maize_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Maize")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 50),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Maize kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Millet consumption

```{r}
#| label: GHS_2023_24_zone_millet_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Millet")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Millet kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Potato consumption

```{r}
#| label: GHS_2023_24_zone_potato_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Potato")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 10),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Potato kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Plantain consumption

```{r}
#| label: GHS_2023_24_zone_plantain_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Plantain")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 40),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Plantain kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )


```

#### Rice consumption

```{r}
#| label: GHS_2023_24_zone_rice_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Rice")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 50),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Rice kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```


#### Sorghum consumption

```{r}
#| label: GHS_2023_24_zone_sorghum_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Sorghum")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sorghum kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

#### Sweet Potato consumption

```{r}
#| label: GHS_2023_24_zone_sweet_potato_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(tidyterra::filter(vect_gpz_geo_GHS_crop, crop == "Sweet potato")) +
  geom_spatvector(aes(fill = crop_equivalent_kg_person_year)) +
  geom_spatvector_text(aes(label = round(crop_equivalent_kg_person_year)), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 15),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sweet Potato kg per person per year", fill = "kg") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

# Appendix 1: Potential extra checks


::: {.callout-note title="Number of days of consumption"}

Section 5C – Aggregate Food Consumption file (`sect5c_harvestw5.dta`) represents an additional opportunity to check the validity of our results:

It has 56,580 observations and records the number of days the household consumed each food-group item during the previous seven days.

We did not use that for the kg calculation, but it was potentially useful as a quality check.

For example, if our calculation says a household consumed 50 kg of maize flour in seven days, we can check whether that household reported consuming maize flour on some/all of the seven days. It could help identify implausible conversion-factor results.
::: 


::: {.callout-caution title="Check the results"}
Before we start converting food products into crops, we need to run the script and look specifically at state_sample and a few major products such as maize, cassava, rice and yam.

There is then a very useful diagnostic we can do: compare the state-level estimates aggregated back to Nigeria against the corresponding national/zone estimates. That will tell us whether the weighting and state-domain calculations are behaving as expected before we build the crop-equivalent calculations on top of them.
:::


::: {.callout-note title="Sensitivity Analysis"}

We did not use the confidence intervals, coefficient of variation, or the upper and lower food to crop conversion factors in our analysis.

These could potentially be combined to show a range of results which could be incoporated in the sub national BPI for Nigeria.

::: 
  


