---
title: "Nigeria Micro-nutrient Deficiencies"
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    page-layout: full
    embed-resources: true
    self-contained-math: true
    toc-expand: true
    theme: united
    code-fold: true
    code-summary: "Show the code"
    css: css/style.css
bibliography: biofort_quarto_better.bib
---

```{r results='hide', warning=FALSE, echo=FALSE, message=FALSE}

library(tidyverse)
library(terra)
library(tidyterra)
library(readODS)
library(ggplot2)
library(ggtext)
library(tidyverse)
library(readODS)
library(kableExtra)
library(DT)
library(geojsonsf)

```

# Introduction


# Nigeria micro-nutrient deficiency data sources

There are number of sources of food consumption and micro-nutrient
deficiency data. The three main sources identified for Nigeria have
different dates, and differing number of deficiencies and populations (
@tbl-MNSources ).

+--------------+--------------+------------+----------------+-------------+
| Source       | Spatial      | Most       | Deficiencies   | Population  |
|              | Resolution   | recent     |                |             |
|              |              | data       |                |             |
+==============+==============+===========:+================+=============+
| National     | Geopolitical | 2021       | - serum        | - women     |
| Food         | Zone         |            |   retinol      |   15–49     |
| Consumption  |              |            |   (vitamin A)  | - children  |
| and Micro    |              |            |                |   6–59      |
| nutrient     |              |            |                |   months    |
| Survey       |              |            |                |             |
| (NFCMS) 2021 |              |            |                |             |
+--------------+--------------+------------+----------------+-------------+
| NFCMS        | "            | "          | - serum        | - women     |
|              |              |            |   ferritin /   |   15–49     |
|              |              |            |   haemoglobin  | - children  |
|              |              |            |                |   6–59      |
|              |              |            |                |   months    |
+--------------+--------------+------------+----------------+-------------+
| NFCMS        | "            | "          | - serum zinc   | - women     |
|              |              |            |                |   15–49     |
|              |              |            |                |             |
|              |              |            |                | - children  |
|              |              |            |                |   6–59      |
|              |              |            |                |   months    |
+--------------+--------------+------------+----------------+-------------+
| Nigeria      | State        | 2023-24    | - haemoglobin  | - children  |
| Demographic  |              |            |   /anaemia     |   6–59      |
| and Health   |              |            |                |   months    |
| Survey       |              |            |                |             |
| (NDHS)       |              |            |                | - women     |
| 2023-24      |              |            |                |   15–49     |
+--------------+--------------+------------+----------------+-------------+
| Modelled     | LGA          | 2018-19    | Apparent       | - household |
| Data from    |              |            | micro-nutrient |             |
| 2018/19 Food |              |            | intake         |             |
| consumption  |              |            | estimated for: |             |
| (Ishida et   |              |            |                |             |
| al., 2026    |              |            | - folate       |             |
| (submitted)) |              |            |                |             |
|              |              |            | - vitamin B12  |             |
|              |              |            |                |             |
|              |              |            | - iron         |             |
+--------------+--------------+------------+----------------+-------------+

:Micro-nutrient deficiency data sources for Nigeria {#tbl-MNSources} 

[@federalgovernmentofnigeriafgonNationalFoodConsumption2024]
[@federalministryofhealthandsocialwelfareofnigeriafmohswNigeriaDemographicHealth2025]
[@ishidaMappingSubnationalVulnerability2026submitted]

The data at state level need to be joined to the Nigeria States spatial
dataset, and the data at zone level need to be joined to the
geopolitical zones.

The stages are:

1.  Load the States dataset
2.  Load the geopolitical zone
3.  Load each health database
4.  For each indicator in the database join to the state or zone spatial
    dataset and plot

# Load the States

The source of the states data is from OSGOF-Ehealth-UNCS. These have
previously been cleaned to remove sliver polygons and gaps, and were
saved as a topojson format.

```{r}
#| label: states
#| message: false
#| warning: false
#| cache: true
#| results: asis

vect_state_geo <- vect("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NGA Admin 2019-2025 OSGOF-Ehealth-UNCS/nga_admin1_topo.topojson")

crs(vect_state_geo) <- "epsg:4326"

vect_state_geo$adm1_pcode <- as.factor(vect_state_geo$adm1_pcode)

# levels(vect_state_geo$adm1_pcode) # Check levels

ggplot(vect_state_geo)+
  geom_spatvector(aes(fill = adm1_pcode)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 3) +
  labs(
    title = "Nigeria States",
    fill = "State<br>Code") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

df_state <- as.data.frame(vect_state_geo)


```

# Load the geopolitical zones

The geopolitical zones are a combination of states with similar ethnic
groups, and/or common political history.

There are six zones:

+---------------+---------------------------------------+
| Zone name     | States                                |
+===============+=======================================+
| North Central | Benue, Kogi, Kwara, Nasarawa, Niger,  |
|               | and Plateau States, and Federal       |
|               | Capital Territory                     |
+---------------+---------------------------------------+
| North East    | Adamawa, Bauchi, Borno, Gombe,        |
|               | Taraba, and Yobe                      |
+---------------+---------------------------------------+
| North West    | Jigawa, Kaduna, Kano, Katsina, Kebbi, |
|               | Sokoto, and Zamfara                   |
+---------------+---------------------------------------+
| South East    | Abia, Anambra, Ebonyi, Enugu, and Imo |
+---------------+---------------------------------------+
| South South   | Akwa Ibom, Bayelsa, Cross River,      |
|               | Delta, Edo, and Rivers                |
+---------------+---------------------------------------+
| South West    | Ekiti, Lagos, Ogun, Ondo, Osun, and   |
|               | Oyo                                   |
+---------------+---------------------------------------+

The zones are created by joining the states.

```{r}
#| label: create_zones
#| message: false
#| warning: false
#| cache: true
#| results: asis

adm1_pcode <- c("NG001", "NG002", "NG003", "NG004", "NG005", "NG006",
"NG007", "NG008", "NG009", "NG010", "NG011", "NG012", "NG013", "NG014",
"NG015", "NG016", "NG017", "NG018", "NG019", "NG020", "NG021", "NG022",
"NG023", "NG024", "NG025", "NG026", "NG027", "NG028", "NG029", "NG030",
"NG031", "NG032", "NG033", "NG034", "NG035", "NG036", "NG037")

geopolitical_zone <- c("South East", "North East", "South South", 
"South East", "North East", "South South", "North Central", "North East",
"South South", "South South", "South East", "South South", "South West",
"South East", "North Central", "North East", "South East", "North West",
"North West", "North West", "North West", "North West", "North Central",
"North Central", "South West", "North Central", "North Central", 
"South West", "South West", "South West", "South West", "North Central", 
"South South", "North West", "North East", "North East", "North West")

df_gpz <- tibble(adm1_pcode, geopolitical_zone)

vect_gpz_geo <- dplyr::left_join(vect_state_geo, df_gpz, by =
c("adm1_pcode")) |> aggregate(by = "geopolitical_zone")

ggplot(vect_gpz_geo)+
  geom_spatvector(aes(fill = geopolitical_zone)) +
  geom_spatvector_text(aes(label = geopolitical_zone), size = 3) +
  labs(
    title = "Nigeria Geopolitical Zones",
    fill = "Zone<br>Name") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

# NFCMS 2021

## Load the database

The NFCMS data are loaded and there are data available for the following
zones:

```{r}
#| label: NFCMS_deficiency_data
#| message: false
#| warning: false
#| cache: true
#| results: asis

df_NFCMS <- read.csv("E:/repos/Biofort_Atlas_Nigeria/tab_data/NFCMS_deficiencies.csv")

df_NFCMS$geopolitical_zone <- as.factor(df_NFCMS$geopolitical_zone) 

cat(paste(levels(df_NFCMS$geopolitical_zone), "; ")) # Check levels


```

## Join the NFCMS database to the zone dataset

The NFCMS dataset does not contain codes for the zones - only names.
Therefore we need to join the two data frames using the name of the
zone - and checking that all zones have data.

```{r}
#| label: NFCMS_join
#| message: false
#| warning: false
#| cache: true
#| results: asis

vect_gpz_geo_NFCMS <- dplyr::left_join(vect_gpz_geo, df_NFCMS, by = c("geopolitical_zone"))
vect_gpz_geo_NFCMS |> as.data.frame() |> 
  DT::datatable(
    rownames = F,
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))


```

# NDHS 2023-24

## Load the databases

The NDHS state data are loaded and there are data available for the
following states:

```{r}
#| label: NDHS_state_deficiency_data
#| message: false
#| warning: false
#| cache: true
#| results: asis

df_NDHS_state <- read.csv("E:/repos/Biofort_Atlas_Nigeria/tab_data/NDHS_state_deficiencies.csv")
# change name to match states dataset
df_NDHS_state['State'][df_NDHS_state['State'] == "FCT-Abuja"] <- "Federal Capital Territory"

df_NDHS_state$State <- as.factor(df_NDHS_state$State) 

cat(paste(levels(df_NDHS_state$State), "; ")) # Check levels

```

The NDHS zone data are loaded and there are data available for the
following zones:

```{r}
#| label: NDHS_zone_deficiency_data
#| message: false
#| warning: false
#| cache: true
#| results: asis

df_NDHS_zone <- read.csv("E:/repos/Biofort_Atlas_Nigeria/tab_data/NDHS_zone_deficiencies.csv")

df_NDHS_zone$Zone <- as.factor(df_NDHS_zone$Zone) 

cat(paste(levels(df_NDHS_zone$Zone), "; ")) # Check levels

```

## Join the NDHS database to the state dataset

The NDHS dataset does not contain codes for the states - only names.
Therefore we need to join the two data frames using the name of the
state - and checking that all states have data.

```{r}
#| label: NDHS_state_join
#| message: false
#| warning: false
#| cache: true
#| results: asis

vect_state_geo_NDHS <- dplyr::left_join(vect_state_geo, df_NDHS_state, by = c("adm1_ref_name"  = "State"))
vect_state_geo_NDHS |> as.data.frame() |> 
  DT::datatable(
    rownames = F,
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))


```

## Join the NDHS database to the zone dataset

The NDHS dataset does not contain codes for the zones - only names.
Therefore we need to join the two data frames using the name of the
zone - and checking that all zones have data.

```{r}
#| label: NDHS_zone_join
#| message: false
#| warning: false
#| cache: true
#| results: asis

vect_zone_gpz_NDHS <- dplyr::left_join(vect_gpz_geo, df_NDHS_zone, by = c("geopolitical_zone"  = "Zone"))
vect_zone_gpz_NDHS |> as.data.frame() |> 
  DT::datatable(
    rownames = F,
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))


```

# Modelled Data from 2018/19 Food consumption

It was decided not to use the data from this study, for two reasons.
Firstly the data are now quite old and the other sources are more
recent. Secondly, the deficiency values are imputed from dietary intake
rather than from biomarkers.

# Compare data sources

To help select the best data source it is necessary to compare the
statistics for each indicator from the two sources.

## Vitamin A deficiency

Vitamin A deficiency is only available from the NFCMS data, and
therefore there is no spatial variation below the geopolitical zone
level.

There are data for women of reproductive age and children between 6 and
59 months.

### Vitamin A deficiency in women of reproductive age (NFCMS, 2021)

```{r}
#| label: NFCMS_vit_A_deficiency_WRA
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_gpz_geo_NFCMS) +
  geom_spatvector(aes(fill = WRA_def_vit_A_pc_246)) +
  geom_spatvector_text(aes(label = WRA_def_vit_A_pc_246), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of vitamin A deficiency in women of reproductive age", fill = " %") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The highest values for vitamin A deficiency among women are in North
West zone with 22.2%. These values are almost double the next highest
value in South South. The lowest values are in South East with 3.4% of
women with a deficiency.

### Vitamin A deficiency in children (NFCMS, 2021)

```{r}
#| label: NFCMS_vit_A_deficiency_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_gpz_geo_NFCMS) +
  geom_spatvector(aes(fill = CH_def_vit_A_pc_236)) +
  geom_spatvector_text(aes(label = CH_def_vit_A_pc_236), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of vitamin A deficiency in children", fill = " %") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The pattern of vitamin A deficiency among children is similar, with the
highest value again in North West zone with 50.6%. These values are 20%
greater than in North East - the zone with next highest value. The
lowest values were again in South East with 6.1% of children with a
vitamin A deficiency.

## Iron deficiency

### Iron deficiency in women of reproductive age (NFCMS, 2021)

Iron deficiency is defined as serum ferritin concentration \<15 µg/L
corrected for inflammation.

```{r}
#| label: NFCMS_iron_deficiency
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_gpz_geo_NFCMS) +
  geom_spatvector(aes(fill = WRA_def_iron_pc_245)) +
  geom_spatvector_text(aes(label = WRA_def_iron_pc_245), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of iron deficiency in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

Iron deficiency in women of reproductive age is highest in North East
geopolitical zone, followed by South South and South East - all three
had values higher than 10%. The zone with lowest value was South West.

### Anaemia in women of reproductive age (NFCMS, 2021)

Iron deficiency anaemia defined as serum ferritin concentration \<15
µg/L, adjusted for inflammation and adjusted haemoglobin \<12 g/dL
Estimates calculated using weights that account for survey design and
non-response.

```{r}
#| label: NFCMS_anaemia
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_gpz_geo_NFCMS) +
  geom_spatvector(aes(fill = WRA_def_anaemia_pc_245)) +
  geom_spatvector_text(aes(label = WRA_def_anaemia_pc_245), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The variation among zones of iron deficiency anaemia was small, varying
between 4.7% and 5.5%, except in South West where the percentage of
anaemic women was only 1.6%. The North East zone once again had the
highest rate.

### Anaemia in women of reproductive age (NDHS, 2023-24)

The Nigeria Demographic and Health Survey of 2023-24 gives data for four
levels of anaemia: (1) Any; (2) Mild; (3) Moderate, and; (4) Severe
anaemia.

**Any anaemia** is when haemoglobin level \<12.0 g/dl

```{r}
#| label: NDHS_zone_anaemia_any
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_zone_gpz_NDHS) +
  geom_spatvector(aes(fill = WRA_anaemia_any)) +
  geom_spatvector_text(aes(label = WRA_anaemia_any), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (any) in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The zone with the highest percentage of women with anaemia was South
East, with over 50% of women. The other zones all had values between 40%
and 50%. The zone with the lowest levels was North Central.

```{r}
#| label: NDHS_anaemia_any
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_state_geo_NDHS) +
  geom_spatvector(aes(fill = WRA_anaemia_any)) +
  geom_spatvector_text(aes(label = WRA_anaemia_any), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (any) in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The individual states show a more mixed pattern. The state with the
highest value is Ebonyi with 71.8% of women anaemic. Other states in the
same zone (South East) have lower values, such as Imo with 41.7%.

In North West zone there are instances of negative spatial
auto-correlation where neighbouring states have very different values -
a clear example is the difference between Sokoto (67.7%) and Kebbi
(21.4%). States in South West zone, in contrast, have more similar
values.

**Mild anaemia** when haemoglobin level 11.0–11.9 g/dl

```{r}
#| label: NDHS_zone_anaemia_mild
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_zone_gpz_NDHS) +
  geom_spatvector(aes(fill = WRA_anaemia_mild)) +
  geom_spatvector_text(aes(label = WRA_anaemia_mild), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (mild) in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The zone with the highest percentage of women with mild anaemia was
South East, with over 30% of women. The other zones all had values
between 20% and 30%. The zone with the lowest levels was North West

```{r}
#| label: NDHS_anaemia_mild
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_state_geo_NDHS) +
  geom_spatvector(aes(fill = WRA_anaemia_mild)) +
  geom_spatvector_text(aes(label = WRA_anaemia_mild), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (mild) in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

For mild anaemia Ebonyi is no longer the state with the highest values.
There are numerous southern states with values around 30%. In the North
East zone the values per state are very similar, whereas in North West
there is again a big difference between Kebbi and its neighbouring
states.

**Moderate anaemia** when haemoglobin level 8.0–10.9 g/dl

```{r}
#| label: NDHS_zone_anaemia_mod
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_zone_gpz_NDHS) +
  geom_spatvector(aes(fill = WRA_anaemia_mod)) +
  geom_spatvector_text(aes(label = WRA_anaemia_mod), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (moderate) in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The zone with the highest percentage of women with moderate anaemia was
again South East, with over 25% of women. The other zones varied between
17% and 23%. The zone with the lowest levels was North Central.

```{r}
#| label: NDHS_anaemia_mod
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_state_geo_NDHS) +
  geom_spatvector(aes(fill = WRA_anaemia_mod)) +
  geom_spatvector_text(aes(label = WRA_anaemia_mod), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (moderate) in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The state with the highest value for moderate anaemia is Sokoto, with
44% of women - a higher value than for mild anaemia. A similar situation
is encountered in Katsina, Adamawa, Abia, Ebonyi, and Akwa Ibom states.

Once again the state with the lowest percentage of women with anaemia is
Kebbi.

**Severe anaemia** when haemoglobin level \<8.0 g/dl

```{r}
#| label: NDHS_zone_anaemia_sev
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_zone_gpz_NDHS) +
  geom_spatvector(aes(fill = WRA_anaemia_sev)) +
  geom_spatvector_text(aes(label = WRA_anaemia_sev), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (severe) in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

There was a different spatial pattern for severe anaemia. The zone with
the highest percentage of women with severe anaemia was North East, 1.7%
of women anaemic. The other zones varied between 0.4% and 1.2%. The zone
with the lowest levels was South West.

```{r}
#| label: NDHS_anaemia_sev
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_state_geo_NDHS) +
  geom_spatvector(aes(fill = WRA_anaemia_sev)) +
  geom_spatvector_text(aes(label = WRA_anaemia_sev), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (severe) in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

Levels of severe anaemia show less obvious differences within
geopolitical zones, except in North East where Bauchi has the highest
value for any state with 3.3% but neighbouring Yobe has a value of 0.6%.
South West zone has consistently low vales for severe anaemia.

### Compare sources

The only micro nutrient for which there are data in different sources is
for iron. However, there are some differences in the indicators used for
iron deficiency anaemia in the two sources.

The NDHS has a finer spatial resolution, and the data were collected
more recently. The only issue would be to decide which of the four
anaemia indicators to use.

### Iron deficiency in children between 6 and 59 months (NFCMS, 2021)

Iron deficiency is defined as serum ferritin concentration \<15 µg/L
corrected for inflammation.

```{r}
#| label: NFCMS_iron_deficiency_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_gpz_geo_NFCMS) +
  geom_spatvector(aes(fill = CH_def_iron_pc_235)) +
  geom_spatvector_text(aes(label = CH_def_iron_pc_235), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of iron deficiency in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

Iron deficiency among children in 2021 was far higher in the northern
geopolitical zones than in the south. North East zone had the highest
values while South South had the lowest. This was a differing pattern to
that for women of reproductive age.

### Anaemia in children between 6 and 59 months (NFCMS, 2021)

```{r}
#| label: NFCMS_anaemia_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_gpz_geo_NFCMS) +
  geom_spatvector(aes(fill = CH_def_anaemia_pc_235)) +
  geom_spatvector_text(aes(label = CH_def_anaemia_pc_235), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

Anaemia among young children is highest in North West zone, with far
lower figures in the southern zones. The pattern is slightly different
to iron deficiency with bigger differences among the northern states

### Anaemia in children between 6 and 59 months (NDHS, 2023-24)

The Nigeria Demographic and Health Survey of 2023-24 gives data for four
levels of anaemia: (1) Any; (2) Mild; (3) Moderate, and; (4) Severe
anaemia.

**Any anaemia** is when haemoglobin level \<12.0 g/dl

```{r}
#| label: NDHS_zone_anaemia_any_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_zone_gpz_NDHS) +
  geom_spatvector(aes(fill = CH_anaemia_any)) +
  geom_spatvector_text(aes(label = CH_anaemia_any), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (any) in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

Levels of child anaemia are high throughout Nigeria with five out of six
zones showing rates above 50%. The zone with the highest percentage is
South East with 63.7%, while the lowest is in South West with 45.5%.

```{r}
#| label: NDHS_anaemia_any_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_state_geo_NDHS) +
  geom_spatvector(aes(fill = CH_anaemia_any)) +
  geom_spatvector_text(aes(label = CH_anaemia_any), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (any) in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The individual states show a more mixed pattern. The state with the
highest value is Ebonyi with 84.8% of children anaemic, the next highest
value is 74.5% in Akwa Ibom state in South South zone.

The state with the lowest value is Delta state with 33.4%.

**Mild anaemia** when haemoglobin level 11.0–11.9 g/dl

```{r}
#| label: NDHS_zone_anaemia_mild_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_zone_gpz_NDHS) +
  geom_spatvector(aes(fill = CH_anaemia_mild)) +
  geom_spatvector_text(aes(label = CH_anaemia_mild), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (mild) in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The zone with the highest percentage of women with mild anaemia was
South South, with over 30% of children However all zones had values
between 26% and 32%. The zone with the lowest levels was South West.

```{r}
#| label: NDHS_anaemia_mild_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_state_geo_NDHS) +
  geom_spatvector(aes(fill = CH_anaemia_mild)) +
  geom_spatvector_text(aes(label = CH_anaemia_mild), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (mild) in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

For mild anaemia in children the differences among states are not great,
with the highest at 38.4% in Abia (South East) and the lowest at 19.8%
in Ondo (South West).

**Moderate anaemia** when haemoglobin level 8.0–10.9 g/dl

```{r}
#| label: NDHS_zone_anaemia_mod_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_zone_gpz_NDHS) +
  geom_spatvector(aes(fill = CH_anaemia_mod)) +
  geom_spatvector_text(aes(label = CH_anaemia_mod), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (moderate) in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The zone with the highest percentage of children with moderate anaemia
was again South East, with over 32% of children The other zones varied
between 19% and 31% - a bigger range than for mild anaemia. The zone
with the lowest levels was South West.

```{r}
#| label: NDHS_anaemia_mod_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_state_geo_NDHS) +
  geom_spatvector(aes(fill = CH_anaemia_mod)) +
  geom_spatvector_text(aes(label = CH_anaemia_mod), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (moderate) in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

There are some very big differences in the levels of moderate anaemia
among the states - with a low of 9% in the Federal Capital Territory and
Delta states, while both Akwa Ibom and Zamfara had levels above 40%.

These are patterns which were masked in the zone map.

**Severe anaemia** when haemoglobin level \<8.0 g/dl

```{r}
#| label: NDHS_zone_anaemia_sev_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_zone_gpz_NDHS) +
  geom_spatvector(aes(fill = CH_anaemia_sev)) +
  geom_spatvector_text(aes(label = CH_anaemia_sev), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (severe) in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

There was a different spatial pattern for severe anaemia. The zone with
the highest percentage of children with severe anaemia was North East,
2.5% of children anaemic. The other zones varied between 0.1% and 1.8%.
The zone with the lowest levels was South West. This showed a similar
pattern to the values for women of reproductive age.

```{r}
#| label: NDHS_anaemia_sev_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis

ggplot(vect_state_geo_NDHS) +
  geom_spatvector(aes(fill = CH_anaemia_sev)) +
  geom_spatvector_text(aes(label = CH_anaemia_sev), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of anaemia (severe) in children between 6 and 59 months", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

Levels of severe anaemia show less obvious differences within
geopolitical zones, except in North East where Adamawa has the highest
value for any state with 4.6% but neighbouring Borno has a value of
0.8%. North West zone also has large differences among states, while
states in South West zone have consistently low vales for severe
anaemia.

### Compare sources

The only micro nutrient for which there are data in different sources is
for iron. However, there are some differences in the indicators used for
iron deficiency anaemia in the two sources.

The NDHS has a finer spatial resolution, and the data were collected
more recently. The only issue would be to decide which of the four
anaemia indicators to use.

## Zinc deficiency

### Zinc deficiency in women of reproductive age (NFCMS, 2021)

```{r}
#| label: NFCMS_zinc_deficiency_WRA
#| message: false
#| warning: false
#| cache: true
#| results: asis


ggplot(vect_gpz_geo_NFCMS) +
  geom_spatvector(aes(fill = WRA_def_zinc_pc_252)) +
  geom_spatvector_text(aes(label = WRA_def_zinc_pc_252), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of zinc deficiency in women of reproductive age", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

There are large differences in the percentage of women of reproductive
age with zinc deficiency among geopolitical zones in Nigeria. The zone
with the lowest values is South East, with 15.5%, whereas in North West
over 60% of women are deficient in zinc.

### Zinc deficiency in children (NFCMS, 2021)

```{r}
#| label: NFCMS_zinc_deficiency_CH
#| message: false
#| warning: false
#| cache: true
#| results: asis


ggplot(vect_gpz_geo_NFCMS) +
  geom_spatvector(aes(fill = CH_def_zinc_pc_238)) +
  geom_spatvector_text(aes(label = CH_def_zinc_pc_238), size = 2) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 100),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Prevalence of zinc deficiency in children", fill = "%") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

The patterns of zinc deficiency for children are almost exactly the same
for children. The zone with the lowest values is again South East, with
11.8%, whereas in North West over 55% of young children are zinc
deficient.
