---
title: "Nigeria Subnational BPI 2024"
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    page-layout: full
    embed-resources: true
    self-contained-math: true
    toc-expand: true
    theme: united
    code-fold: true
    code-summary: "Show the code"
    css: css/style.css
bibliography: biofort_quarto_better.bib
---

```{r results='hide', warning=FALSE, echo=FALSE, message=FALSE}

library(tidyverse)
library(terra)
library(tidyterra)
library(readODS)
library(ggplot2)
library(ggtext)
library(kableExtra)
library(DT)
library(geojsonsf)
library(DT)
library(spdep)

```

# Introduction

The Biofortification Priority Indices (BPI) have been used by HarvestPlus for targeting at the country level for many years [@asare-marfoPrioritizingCountriesBiofortifcation2013], and has a well developed portal <https://bpi.harvestplus.org/>. But as @lividiniIdentificationOptimalInvestments2017 recognise, the national level BPIs can hide important differences within countries. As a result, a number of efforts have been made to develop sub-national biofortification priority indices [e.g. @funesSpatialPerspectiveIntroducing2015], including for Nigeria [@herringtonSubnationalPrioritizationBiofortification2018].

The Biofortification Atlas cannot afford to ignore the work already carried out and with new data available there is the opportunity to update the sub national BPI for Nigeria, and to analyse the changes since 2018. Re-creating the BPIs from 2018 would involve a considerable time effort, in addition the authors were unable to obtain indicators for zinc.

The specific activities will be:

- Review 2018 biofortification priority indices, for instance assessing how well the consumption data relate to food baskets and diets for target populations. 
- Smooth 2021 consumption and deficiency data to state level (following method in Herrington et al., 2018) 
- Create updated BPIs weighted by land-area and population 
- Compare state level Target Area Classification for each crop

# Review 2018 Subnational Prioritisation for Biofortification Interventions in Nigeria

The objectives of this short review are to:

(1) identify and assess the data sources used in the 2018 study
(2) seek more recent equivalent data, or data for alternative indicators
(3) assess how well the consumption data relate to food baskets and diets of the target populations

## Identify and assess the data sources used in the 2018 study

The 2018 study constructs and analyses a subnational version of the Biofortification Priority Index (BPI) [@asare-marfoPrioritizingCountriesBiofortifcation2013]. The basic components of the BPI are: (i) the intensity of supply of a specific crop, (ii) the intensity of demand for that crop or its food products, and (iii) the deficiency levels for the micronutrient with which the crop can be enriched, all in a given geographical unit.

The indicators for these components will differ depending on the scale of the analysis and the data available. The Global BPI uses country-level data that are standardised across countries, while the challenge for subnational indices is to find the same or similar variables for which there are spatial data varying within a single country.


### Production Index

The global BPI uses three variables in the production index (@tbl-global_prod_index). These consider the percentage of the harvested area that is dedicated to the crop, the harvested area weighted by population, and the share of production which is exported.

| Variable Name | Variable Explanation | Main Data Source and Year |
|------------------|------------------------------------|------------------|
| Share of area harvested (%) | Total area harvested of a specific crop (ha)/total agricultural land of a country (ha) | FAO 2010 |
| Per-capita area harvested (ha) | Total area harvested of a specific crop (ha)/total population in the country | FAO 2010 |
| Export share (%) | If Production greater than 0: Export share =Exports/(Production+Imports); otherwise export share is 0 % | FAO 2010 |

: Variables Used for the Calculation of the Production Index of the Global BPIs {#tbl-global_prod_index}

::: {.callout-note title="Rescaling"}
All variables are rescaled to a range between 0 and 1 by applying the following formula in @eq-rescale

$$V_r = \frac{V_a – V_{min}} {V_{max} – V_{min}}$$ {#eq-rescale} Where: - $V_r$ is the rescaled value - $V_a$ is the actual value - $V_{min}$ is the minimum value - $V_{max}$ is the maximum value
:::

At the subnational level (@tbl-national_prod_index), the export share is not considered explicitly in the index as a separate variable - instead, states exporting a large percentage of production can be targeted with interventions that focus on the market incentives to produce biofortified crops (e.g. agronomic practices, marketing or processing linkages).

| Variable Name | Variable Explanation | Main Data Source and Year |
|------------------|------------------------------------|------------------|
| Share of area harvested (%) | Total area harvested of a specific crop (ha) / total agricultural land of a state (ha) | (NAERLS and FDAE, 2014);(HarvestChoice, 2014); (FAOSTAT 2014) |
| Per-capita area harvested (ha) | Total area harvested of a specific crop (ha) / total population in the state | FAO 2010 |

: Variables Used for the Calculation of the Production Index of the Subnational BPIs {#tbl-national_prod_index}

The percentage land area allocated to a given crop of interest, of total agricultural land area cultivated in a state was estimated by summing all crop land areas outlined in the 2014 Agricultural Performance Survey report. The authors acknowledge that the percentages may be biased upwards since not all crops were included in the report. However, total land area cultivated at the state-level was not available.

To calculate per capita land area, state-level population was obtained from the 2016 Annual Abstract of Statistics Volume 1 Report, produced by the National Bureau of Statistics in Nigeria. To match the timeframe of production estimates, 2014 population figures were used.

### Consumption Index

The consumption index consists of two variables (@tbl-global_cons_index) in the global BPI, and a single variable in the subnational BPI (@tbl-national_cons_index).

| Variable Name | Variable Explanation | Main Data Source and Year |
|------------------|------------------------------------|------------------|
| Per-capita food consumption | Per-capita food consumption (kg/year) | FAO 2010 |
| Import Dependency Ratio=Import Share (%) | If Production greater than 0: Import share =Imports/(Production+Imports-Exports); otherwise, import share is 10 % | FAO 2010 |

: Variables Used for the Calculation of the Consumption Index of the Global BPIs {#tbl-global_cons_index}

The import dependency ratio is not collected for states in Nigeria. Instead states importing a large proportion of the volumes of food products consumed can be targeted with different interventions.

| Variable Name | Variable Explanation | Main Data Source and Year |
|----|----|----|
| Per-capita food consumption | Per-capita food consumption (kg/year) | World Bank Living Standards Measurement Studies – General Household Surveys. Specifically Wave 3 data, 2015-2016 |

: Variables Used for the Calculation of the Consumption Index of the Subnational BPIs {#tbl-national_cons_index}

### Micronutrient Deficiency Index

Global BPI (@tbl-global_def_index) uses two variables each for Vitamin A, Zinc and Iron deficiencies.
These variables are available for individual nations from the WHO Global Databases on
Vitamin A Deficiency and Anaemia, as well as other sources. @asare-marfoPrioritizingCountriesBiofortifcation2013 recognise that there is correlation between the variables for each micronutrient, but prefer to use two indicators to account for a lack of reliability in either one of the variables.

| Variable Name | Variable Explanation | Main Data Source and Year |
|------------------|------------------------------------|------------------|
| SerumRetinol07 | Proportion of preschool-age children with retinol \< 0.70 umol/l | WHO 2009 |
| DALYs Vitamin A Deficiency (VAD) | Age-standardised DALYs lost per 100,000 inhabitants to VAD | WHO 2009 |
| InadequateZinc | Percentage of population at risk of inadequate intake of zinc | Hotz and Brown 2004 |
| Stunting | Prevalence of stunting among children age 6-59 months | WHO 2008 |
| DALYs Iron Deficiency Anaemia (IDA | Age-standardized DALYs lost per 100,000 inhabitants to IDA | WHO 2002 |
| HbLess110 | Proportion of preschool-age children with Hb \< 110 g/l | ICF International 2012; WHO 2008 |

: Variables Used for the Calculation of the Micronutrient Deficiency Index of the Global BPIs {#tbl-global_def_index}

For their subnational BPIs @herringtonSubnationalPrioritizationBiofortification2018, do not include any variables for zinc deficiency. Their indicators for vitamin A and iron are also slightly different to those for the Global BPIs. Proxies were used for some indicators - notably  women's prevalence of aneamia - leading to a lack of representation for a portion of the HarvestPlus target population (women of reproductive age who are not pregnant or who have not been pregnant). 


| Variable Name | Variable Explanation | Main Data Source and Year |
|------------------|------------------------------------|------------------|
| VitARich | Children consuming foods rich in vitamin A |  2013 (DHS) 2015 (MIS) |
| VitASupl | Children given vitamin A supplements | 2013 (DHS) 2015 (MIS) |
| DiarhChil | Children with diarrhea | 2013 (DHS) 2015 (MIS) |
| IronSupl | Women iron syrup supplements during pregnancy| 2013 (DHS) 2015 (MIS) |
| IDAChil | Children with anemia | 2013 (DHS) 2015 (MIS) |
| IronRich | Children consuming foods rich in iron | 2013 (DHS) 2015 (MIS) |


: Variables Used for the Calculation of the Micronutrient Deficiency Index of the Subnational BPIs {#tbl-national_def_index}

## Seek more recent equivalent data, or data for alternative indicators

### Production Index


### Consumption Index

The consumption index consisted of a single variable in the subnational BPI (@tbl-national_cons_index).

This was updated using the Wave 5 2023-24 General Household Surveys

| Variable Name | Variable Explanation | Main Data Source and Year |
|----|----|----|
| Per-capita food consumption | Per-capita crop consumption (kg/year) | World Bank Living Standards Measurement Studies – General Household Surveys. Specifically Wave 3 data, 2023-2024 |

: Variables available for the Calculation of the Consumption Index of the Subnational BPIs {#tbl-national_cons_index}

We calculated crop consumption per capita per year at the state and the zone level. The survey sampling strategy is not representative at the state level, but the survey deign allowed us to estimate direct, design-weighted state estimates. For some states the overall number of households was small, and for some food products there were often few responses and as a consequence a large coefficient of variation associated with the estimates.





### Micronutrient Deficiency Index

- Serum retinol (vitamin A) for women of reproductive age and children between 6 and 59 months is available from 2021 (this was not available for 2018 study) 

- Age standardised DALYs Vitamin A Deficiency (VAD) available at subnational level for 2023

- Percentage of population at risk of inadequate intake of zinc. This is now available at the subnational level from the 2021 NFCMS as *womens's prevalence of inadequacy of zinc intakes*. A further indicator is also available:  *prevalence of women of reproductive age with zinc deficiency* and *prevalence of children with zinc deficiency*
These are only available at the geopolitical zone level in Nigeria, but are potentially an improvement on both the global and 2018 subnational indicators

- Prevalence of stunting among children age 6-59 months available at subnational level for 2023-24

- Age-standardized DALYs lost per 100,000 inhabitants to IDA. The IMHE GBD database includes Age standardised DALYs Dietary iron Deficiency - a slightly different indicator, which focuses on the dietary component of anaemia, but is available at the state level

- Proportion of preschool-age children with Hb \< 110 g/l - the NFCMS includes Iron deficiency in children defined as serum ferritin concentration \<15 µg/L corrected for inflammation. Both the NFCMS and NDHS include anaemia for pre-school children - for the NDHS at the state level, and for NFCMS at the zone level. The definition of any anaemia in the NDHS is is when haemoglobin level is <12.0 g/dl. A closer level of Haemoglobin to the global BPI is for moderate and severe anaemia, which combined have an upper limit of 11 g/dl.


## Assess relationship between consumption data and food baskets 


# Rescale spatially smoothed benchmarked state-level crop consumption estimates

For each crop we take the minimum and maximum kg per capita per year


$$V_r = \frac{V_a – V_{min}} {V_{max} – V_{min}}$$ {#eq-rescale} Where: - $V_r$ is the rescaled value - $V_a$ is the actual value - $V_{min}$ is the minimum value - $V_{max}$ is the maximum value

```{r}
#| label: GHS_crop_consumption_rescale_function
#| message: false
#| warning: false
#| cache: false
#| results: asis

# custom function to implement min max scaling
minMax <- function(x) {
  (x - min(x,
      na.rm = T)) / (max(x,
      na.rm = T) - min(x,
      na.rm = T))
}

df_GHS_pop_crop_state <- read.csv("tab_data/GHS food per capita outputs/crop_state_results_benchmarked_smoothed.csv")


df_GHS_pop_crop_state_wide <- df_GHS_pop_crop_state |> 
  select(state, crop, benchmarked_crop_equivalent_kg_person_year_state) |> 
  pivot_wider(names_from = crop, values_from = benchmarked_crop_equivalent_kg_person_year_state)

df_normalised <- as.data.frame(lapply(select(df_GHS_pop_crop_state_wide, -state), minMax))
head(df_normalised)

df_GHS_pop_crop_state_rescale <- bind_cols(select(df_GHS_pop_crop_state_wide, state), df_normalised)

```

### Load the States

```{r}
#| label: states
#| message: false
#| warning: false
#| cache: false
#| results: asis

vect_state_geo <- vect("D:/DatosProyecto/SpatialData/Africa/Pais/Nigeria/NGA Admin 2019-2025 OSGOF-Ehealth-UNCS/nga_admin1_topo.topojson")

crs(vect_state_geo) <- "epsg:4326"

vect_state_geo$adm1_pcode <- as.factor(vect_state_geo$adm1_pcode)

# levels(vect_state_geo$adm1_pcode) # Check levels

ggplot(vect_state_geo)+
  geom_spatvector(aes(fill = adm1_pcode)) +
  geom_spatvector_text(aes(label = adm1_pcode), size = 3) +
  labs(
    title = "Nigeria States",
    fill = "State<br>Code") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

df_state <- as.data.frame(vect_state_geo)


```

### Load the geopolitical zones

The geopolitical zones are a combination of states with similar ethnic groups, and/or common political history.

There are six zones:

+---------------+----------------------------------------------------------------------------------------+
| Zone name     | States                                                                                 |
+===============+========================================================================================+
| North Central | Benue, Kogi, Kwara, Nasarawa, Niger, and Plateau States, and Federal Capital Territory |
+---------------+----------------------------------------------------------------------------------------+
| North East    | Adamawa, Bauchi, Borno, Gombe, Taraba, and Yobe                                        |
+---------------+----------------------------------------------------------------------------------------+
| North West    | Jigawa, Kaduna, Kano, Katsina, Kebbi, Sokoto, and Zamfara                              |
+---------------+----------------------------------------------------------------------------------------+
| South East    | Abia, Anambra, Ebonyi, Enugu, and Imo                                                  |
+---------------+----------------------------------------------------------------------------------------+
| South South   | Akwa Ibom, Bayelsa, Cross River, Delta, Edo, and Rivers                                |
+---------------+----------------------------------------------------------------------------------------+
| South West    | Ekiti, Lagos, Ogun, Ondo, Osun, and Oyo                                                |
+---------------+----------------------------------------------------------------------------------------+

The zones are created by joining the states.

```{r}
#| label: create_zones
#| message: false
#| warning: false

adm1_pcode <- c("NG001", "NG002", "NG003", "NG004", "NG005", "NG006", "NG007", "NG008", "NG009", "NG010", "NG011", "NG012", "NG013", "NG014", "NG015", "NG016", "NG017", "NG018", "NG019", "NG020", "NG021", "NG022", "NG023", "NG024", "NG025", "NG026", "NG027", "NG028", "NG029", "NG030", "NG031", "NG032", "NG033", "NG034", "NG035", "NG036", "NG037")

geopolitical_zone<- c("South East", "North East", "South South", "South East", "North East", "South South", "North Central", "North East", "South South", "South South", "South East", "South South", "South West", "South East", "North Central", "North East", "South East", "North West", "North West", "North West", "North West", "North West", "North Central", "North Central", "South West", "North Central", "North Central", "South West", "South West", "South West", "South West", "North Central", "South South", "North West", "North East", "North East", "North West")

df_gpz <- tibble(adm1_pcode, geopolitical_zone)

vect_gpz_geo <- dplyr::left_join(vect_state_geo, df_gpz, by = c("adm1_pcode")) |> aggregate(by = "geopolitical_zone")

ggplot(vect_gpz_geo)+
  geom_spatvector(aes(fill = geopolitical_zone)) +
  geom_spatvector_text(aes(label = geopolitical_zone), size = 3) +
  theme_minimal()

```

## Join back to States dataset

```{r}
#| label: GHS_crop_rescaled_consumption_state_join
#| message: false
#| warning: false
#| cache: false
#| results: asis

vect_state_geo_GHS_crop_rescale <- dplyr::left_join(vect_state_geo, df_GHS_pop_crop_state_rescale, by = c("adm1_ref_name" = "state"))

```

## Cassava consumption

```{r}
#| label: GHS_2023_24_state_cassava_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Cassava)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cassava consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Cocoyam consumption

```{r}
#| label: GHS_2023_24_state_cocoyam_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Cocoyam)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cocoyam consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Cowpea consumption

```{r}
#| label: GHS_2023_24_state_cowpea_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Cowpea)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Cowpea consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Maize consumption

```{r}
#| label: GHS_2023_24_state_maize_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Maize)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Maize consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Millet consumption

```{r}
#| label: GHS_2023_24_state_millet_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Millet)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Millet consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Potato consumption

```{r}
#| label: GHS_2023_24_state_potato_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Potato)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Potato consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Plantain consumption

```{r}
#| label: GHS_2023_24_state_plantain_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Plantain)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Plantain consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Rice consumption

```{r}
#| label: GHS_2023_24_state_rice_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Rice)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Rice consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Sorghum consumption

```{r}
#| label: GHS_2023_24_state_sorghum_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Sorghum)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sorghum consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Sweet potato consumption

```{r}
#| label: GHS_2023_24_state_sweet_potato_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Sweet.potato)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Sweet potato consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Wheat consumption

```{r}
#| label: GHS_2023_24_state_wheat_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Wheat)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Wheat consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

## Yam consumption

```{r}
#| label: GHS_2023_24_state_yam_rescaled_consumption_map
#| message: false
#| warning: false
#| cache: false
#| results: asis

ggplot(vect_state_geo_GHS_crop_rescale) +
  geom_spatvector(aes(fill = Yam)) +
  scale_fill_stepsn(
    n.breaks = 10,
    limits = c(0, 1),
    colours = hcl.colors(palette = "Grays", 10, rev = T),
    guide = "legend",
    na.value = "pink",
    labels = scales::comma
  ) +
  labs(title = "Yam consumption per state (smoothed, benchmarked & rescaled)", fill = "") +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

```

# Micronutrient Deficiency Index

## Serum retinol (vitamin A) children between 6 and 59 months




## Age standardised DALYs Vitamin A Deficiency (VAD) available at subnational level for 2023






- Percentage of population at risk of inadequate intake of zinc. This is now available at the subnational level from the 2021 NFCMS as *womens's prevalence of inadequacy of zinc intakes*. A further indicator is also available:  *prevalence of women of reproductive age with zinc deficiency* and *prevalence of children with zinc deficiency*
These are only available at the geopolitical zone level in Nigeria, but are potentially an improvement on both the global and 2018 subnational indicators

- Prevalence of stunting among children age 6-59 months available at subnational level for 2023-24

- Age-standardized DALYs lost per 100,000 inhabitants to IDA. The IMHE GBD database includes Age standardised DALYs Dietary iron Deficiency - a slightly different indicator, which focuses on the dietary component of anaemia, but is available at the state level

- Proportion of preschool-age children with Hb \< 110 g/l - the NFCMS includes Iron deficiency in children defined as serum ferritin concentration \<15 µg/L corrected for inflammation. Both the NFCMS and NDHS include anaemia for pre-school children - for the NDHS at the state level, and for NFCMS at the zone level. The definition of any anaemia in the NDHS is is when haemoglobin level is <12.0 g/dl. A closer level of Haemoglobin to the global BPI is for moderate and severe anaemia, which combined have an upper limit of 11 g/dl.
