---
title: "Biofortified Varieties in target agro-ecological zones"
format:
  html:
    toc: true
    toc-depth: 1
    toc-location: left
    page-layout: full
    embed-resources: true
    self-contained-math: true
    toc-expand: true
    theme: united
    code-fold: true
    code-summary: "Show the code"
    css: /css/style.css
bibliography: biofort_quarto_better.bib
editor: 
  markdown: 
    wrap: 72
---

```{r results='hide', warning=FALSE, echo=FALSE, message=FALSE}

library(tidyverse)
library(terra)
library(tidyterra)
library(readODS)
library(ggplot2)
library(ggtext)
library(tidyr)

```

# Introduction

The interventions associated with crop biofortification vary depending
on the stage of the genetic innovation pipeline [@lungaho2025], from
design to delivery.

The spatial components of those interventions are most visible in the
product development phase, during the testing of materials, and the
delivery phase, where new Varieties, are released and are multiplied
within the seed system.

Information on testing locations is generally sparse, but the testing
regime provides information (e.g. via National Variety Trials) on the
geographies for which a new biofortified crop variety is adapted. When
the variety is released it can be linked to the zones for which it is
adapted and mapped.

# Methods 

## Extracting released variety information

Variety adaptation is one of the pieces of information contained in the
Nigerian Seed Portal Initiative [@nigerianseedportalinitiative2026].

The seed portal contains information about released seed Varieties, or
other improved planting material. The portal has information on improved
varieties of over 30 crops ( @tbl-cropvar).

+------------------------------------------------+---------------------+
| Crop                                           | Improved Varieties  |
+================================================+=====================+
| Maize                                          | 210                 |
+------------------------------------------------+---------------------+
| Rice                                           | 85                  |
+------------------------------------------------+---------------------+
| Sorghum                                        | 60                  |
+------------------------------------------------+---------------------+
| Cassava                                        | 56                  |
+------------------------------------------------+---------------------+
| Cowpea                                         | 44                  |
+------------------------------------------------+---------------------+
| Yam                                            | 31                  |
+------------------------------------------------+---------------------+
| Groundnut                                      | 29                  |
+------------------------------------------------+---------------------+
| Soybean                                        | 28                  |
+------------------------------------------------+---------------------+
| Wheat                                          | 26                  |
+------------------------------------------------+---------------------+
| Tomato                                         | 24                  |
+------------------------------------------------+---------------------+
| Sugarcane                                      | 21                  |
+------------------------------------------------+---------------------+
| Pearl Millet                                   | 19                  |
+------------------------------------------------+---------------------+
| Cotton                                         | 17                  |
+------------------------------------------------+---------------------+
| Rubber, Cocoa                                  | 14                  |
+------------------------------------------------+---------------------+
| Forage Legume, Pepper , Sweet Orange           | 11                  |
+------------------------------------------------+---------------------+
| Potato (& Irish Potato)                        | 10                  |
+------------------------------------------------+---------------------+
| Sweet Potato                                   | 9                   |
+------------------------------------------------+---------------------+
| Amaranthus                                     | 8                   |
+------------------------------------------------+---------------------+
| Sesame                                         | 7                   |
+------------------------------------------------+---------------------+
| Shika Grass, Barley                            | 6                   |
+------------------------------------------------+---------------------+
| Oat, Castor, Corchorus                         | 5                   |
+------------------------------------------------+---------------------+
| Sunflower, Solanum                             | 4                   |
+------------------------------------------------+---------------------+
| Sokoyokoto, Coconut, Kola, Okra, Kenaf         | 3                   |
+------------------------------------------------+---------------------+
| Ginger, Melon, Coffee, Cabbage, Acha           | 2                   |
+------------------------------------------------+---------------------+

: \# of varieties per crop {#tbl-cropvar}

It would be time consuming to extract information about these varieties
manually. Instead a script was written in R to ‘scrape’ the website and
automatically extract the pertinent information.

Each ‘page’ of the seed portal contains summaries of 20 varieties and
there are 41 pages.

For each variety summary the scraper extracts:

- Entry number

- Crop type

- Variety name + variety code (split from the combined field)

- Release year & registration year

- Outstanding characteristics (full text)

- Yield (extracted separately from characteristics, e.g. `10.8t/ha`)

The script loops through all 41 pages:

- Loops pages 1–41 using the URL format provided

- Polite delay of 1.5 seconds between requests `(DELAY_SECS`) — keeps
  you from hammering the server; adjust as needed

- Auto-retry — if a page fetch fails, it waits 3 seconds and tries once
  more before skipping

- Page number column added to the output so each entry can be traced to
  a page

- Combined output saved to a single `seedportal_varieties.csv` and
  `seedportal_varieties.xlsx`

Each seed variety has a separate page with more information, the detail
pages contain these extra fields:

- National Code

- Original Name

- Agroecological Zones

- Origin/Source

- Developing Institute

- Breeder(s)/Collaborators

- Potential Yield (t/ha)

The links on the list pages contain both `varid` and `cropid`, so the
strategy is: scrape the links from each list page, then visit each
detail page.

The script is a two-stage scraper.

Stage 1 — Collect links loops all 41 list pages and harvests every
`varid`= link, then deduplicates them. The detail page URL format is
`variety.php?varid=XXX&cropid=YYY&task=view`.

Stage 2 — Scrape detail pages visits each link and extracts the full set
of fields that only appear on the detail page:

| Detail                       | Example                              |
|------------------------------|--------------------------------------|
| Crop Name                    | Cowpea                               |
| National Code                | NGVU-16-34                           |
| Variety Name                 | FUAMPEA 2                            |
| Original Name                | UAM09 1051-1                         |
| Outstanding Characteristics  | Resistance to Striga…                |
| Agroecological Zones         | Northern Guinea and Sudan Savanna    |
| Origin/Source                | IAR-ABU, Zaria and IITA              |
| Developing Institute         | FUAM and IAR-ABU, Zaria              |
| Breeder(s)/Collaborators     | L.O. Omoigui, L.L. Bello…            |
| Potential Yield (t/ha)       | 2 (numeric)                          |
| Year of Release              | 2016                                 |
| Year of Registration         | 2016                                 |

: Details of variety

The scraper further parses the results after scraping to assign flags
depending on whether the variety characteristics include high or notable
levels of micronutrients:

(1) Add a new column called ‘zinc’. If the ‘outstanding characteristic’
    column includes the words ‘Zinc’ or ‘Zn’ then code the zinc column
    as 1, otherwise 0

(2) Add a new column called ‘vit_A’. If the ‘outstanding characteristic’
    column includes the words ‘vitamin A’ or ‘vit A’ or ‘beta carotene’
    , then code the vit_A column as 1, otherwise 0.

(3) Add a new column called ‘iron’. If the ‘outstanding characteristic’
    column includes the words ‘Iron’ or ‘Fe’ then code the iron column
    as 1, otherwise 0 .

(4) Add a new column called ‘n_guinea’ . If the ‘Agroecological Zones’
    column includes the word ‘northern guinea’ or ‘N guinea’ or just
    ‘northern’ then code the ‘n_guinea’ column as 1, otherwise 0 .

(5) Add a new column called ‘s_guinea’ . If the ‘Agroecological Zones’
    column includes the word ‘southern guinea’ or ‘S guinea’ or just
    ‘southern’ then code the ‘n_guinea’ column as 1, otherwise 0 .

(6) . If the ‘Agroecological Zones’ column includes the word ‘guinea’
    without northern or southern then code both the ‘n_guinea’ and
    ‘s_guinea’ column as 1 .

(7) Add a new column called ‘sudan’ . If the ‘Agroecological Zones’
    column includes the word ‘sudan’ or ‘sudano’ then code the ‘sudan’
    column as 1, otherwise 0 .

(8) Add a new column called ‘derived’ . If the ‘Agroecological Zones’
    column includes the word ‘derived’ or ‘transition’ then code the
    ‘derived’ column as 1, otherwise 0 .

(9) Add a new column called ‘forest’ . If the ‘Agroecological Zones’
    column includes the word ‘forest’ then code the ‘forest’ column as
    1, otherwise 0 .

(10) . If the ‘Agroecological Zones’ column includes the word ‘savanna’
     or ‘savanna’ without ‘guinea’ or ‘sudan’ then code the ‘n_guinea’,
     ‘sudan’ and ‘s_guinea’ column as 1 .

(11) Add a new column called ‘mid_alt’ . If the ‘Agroecological Zones’
     column includes the word ‘mid altitude’ or ‘mid alt’ then code the
     ‘mid_alt’ column as 1, otherwise 0 .

(12) Add a new column called ‘lowland’ . If the ‘Agroecological Zones’
     column includes the word ‘lowland’ or ‘low’ then code the ‘lowland’
     column as 1, otherwise 0 .

(13) Add a new column called ‘sahel’ . If the ‘Agroecological Zones’
     column includes the word ‘sahel’ or ‘sahelian’ then code the
     ‘sahel’ column as 1, otherwise 0 . In all cases ignore the case of
     the word and also include parts of words.

(14) Add a new column called ‘all’ . If the ‘Agroecological Zones’
     column includes the word ‘all’ then code the ‘all’ column as 1,
     otherwise 0 . In all cases ignore the case of the word and also
     include parts of words.

+----------------------------------+----------------------------------+
| Column                           | Logic                            |
+==================================+==================================+
| zinc                             | “zinc” or “Zn” in                |
|                                  | characteristics                  |
+----------------------------------+----------------------------------+
| vit_A                            | “vitamin A”, “vit A”, or “beta   |
|                                  | carotene”                        |
+----------------------------------+----------------------------------+
| iron                             | “iron” or “Fe”                   |
+----------------------------------+----------------------------------+
| n_guinea                         | northern guinea / N guinea /     |
|                                  | northern / plain “guinea” /      |
|                                  | plain “savanna”                  |
+----------------------------------+----------------------------------+
| s_guinea                         | southern guinea / S guinea /     |
|                                  | southern / plain “guinea” /      |
|                                  | plain “savanna”                  |
+----------------------------------+----------------------------------+
| sudan                            | sudan / sudano / plain “savanna” |
+----------------------------------+----------------------------------+
| derived                          | “derived” or “transition”        |
+----------------------------------+----------------------------------+
| forest                           | “forest”                         |
+----------------------------------+----------------------------------+
| mid_alt                          | “mid altitude” or “mid alt”      |
+----------------------------------+----------------------------------+
| lowland                          | “lowland” or “low”               |
+----------------------------------+----------------------------------+
| sahel                            | “sahel” or “sahelian”            |
+----------------------------------+----------------------------------+
| all_zones                        | “all”                            |
|                                  |                                  |
| (named all_zones to avoid        |                                  |
| conflict with R’s built-in       |                                  |
| all())                           |                                  |
+----------------------------------+----------------------------------+

: extraction logic

The savanna cascade (rule 10) and the bare “guinea” rule (rule 6) are
handled cleanly through intermediate boolean flags before the final
as.integer() conversion.

## Constructing Agro-Ecological Zones for Nigeria

There is no source of agro-ecological zones that incorporate the zones
most commonly referenced and used in Nigeria i.e. the “Sahel Savanna”,
“Sudan Savanna”, “Northern Guinea Savanna”, “Southern Guinea Savanna”,
“Derived/Transition Savanna”, “Forest”, and “Mid-Altitude zones.

In 2015 HarvestChoice developed Agro-Ecological Zones (AEZ) for Africa
South of the Sahara (SSA), based on the methodology developed by FAO and
IIASA. The dataset includes three classification schemes: 5, 8, and 16
classes, referred to as the AEZ5, AEZ8, and AEZ16, respectively.

The AEZ16 scheme specifically groups regions into classes based only on
climatic criteria and thresholds. There are no soil or landform
criteria, in contrast to the FAO/IIASA Global Agro-Ecological Zones
(GAEZ) methodology:

Climatic Characteristics

Length of Growing Period (LGP): The number of days in a year when
temperature and moisture conditions are suitable for crop growth.

Thermal Regimes: Defined by temperature regimes (e.g., tropical,
subtropical, temperate) that influence crop suitability.

Aridity Index: A measure of water availability, calculated as the ratio
of precipitation to potential evapotranspiration. This helps distinguish
between arid, semi-arid, and humid zones

The HarvestChoice dataset only has five zones for Nigeria:

- Tropic - warm / humid

- Tropic - warm / subhumid

- Tropic - cool / subhumid

- Tropic - warm / semiarid

- Tropic - warm / arid

The agro-ecological zones used in Nigeria do not map exactly onto the
harvest choice zones, but they do use similar criteria
[@akinwumiju2020a].

For this analysis I use the following criteria and thresholds

+------------------------+-------------------+------------------------+
| AEZ Name               | LGP (2020) (days) | Elevation              |
|                        |                   | (masl)[@reuter2007a]   |
+========================+===================+========================+
| Sahel Savanna          | \<= 90            | \< 800                 |
+------------------------+-------------------+------------------------+
| Sudan Savanna          | 90 - 150          | \< 800                 |
+------------------------+-------------------+------------------------+
| Northern Guinea        | 150 - 181         | \< 800                 |
| Savanna                |                   |                        |
+------------------------+-------------------+------------------------+
| Southern Guinea        | 181 - 211         | \< 800                 |
| Savanna                |                   |                        |
+------------------------+-------------------+------------------------+
| Derived/Transition     | 211 - 270         | \< 800                 |
| Savanna                |                   |                        |
+------------------------+-------------------+------------------------+
| Forest                 | \> 270            | \< 800                 |
+------------------------+-------------------+------------------------+
| Mid-Altitude zones     |                   | \> 800                 |
+------------------------+-------------------+------------------------+

# Load the Agro-Ecological Zones

```{r}
#| label: fig-aez
#| message: false
#| warning: false
#| fig-cap: " Nigeria Agro-ecological zones based on 2020 LGP and elevation"
#| 
vect_aez_geo <- vect("spatial_data/IITA_AEZ_LGP_2020_SRTM_RCL_NGA_topo.topojson")
vect_aez_geo$AEZ <- as.factor(vect_aez_geo$AEZ)
#levels(vect_aez_geo$AEZ) # Check levels
#plot(vect_aez_geo, main = "Nigeria Agro-ecological zones based on 2020 LGP and elevation", levels=TRUE)

AEZ <- 1:7
AEZ_name <- c("Sahel Savanna", "Sudan Savanna", "Northern Guinea Savanna", "Southern Guinea Savanna", "Derived/Transition Savanna", "Forest", "Mid-Altitude")

df_aez_names <- tibble(AEZ, AEZ_name)
df_aez_names$AEZ <- as.factor(df_aez_names$AEZ)
#levels(df_aez_names$AEZ) # Check levels

vect_aez_geo <- dplyr::left_join(vect_aez_geo, df_aez_names, by = c("AEZ") )

ggplot(vect_aez_geo)+
  geom_spatvector(aes(fill = AEZ_name)) +
  theme(
    plot.title = element_markdown(),
    legend.title = element_markdown(),
    axis.title = element_blank()
  )

df_aez <- as.data.frame(vect_aez_geo)

```

# Load the Variety data

## Load the Nigeria Seed Portal variety database

```{r}
#| label: nsp_variety_data
#| message: false
#| warning: false

df_nsp <- read.csv("tab_data/seedportal_varieties.csv")

#str(df_nsp)
#summary(df_nsp)

```

## Filter Bio-fortified Varieties

```{r}
#| label: filter_biofort
#| message: false
#| warning: false

df_nsp_biofort <- dplyr::filter(df_nsp, zinc == 1 |
                                  vit_A == 1 | iron == 1)

paste("Number of Bio-fortified varieties =", nrow(df_nsp_biofort))
biofort_crop_list <- as.list(unique(df_nsp_biofort$Crop.Name))


df_nsp_biofort_zinc <- dplyr::filter(df_nsp, zinc == 1)
paste("Number of varieties Bio-fortified with zinc =", nrow(df_nsp_biofort_zinc))

df_nsp_biofort_crop_zinc_sum <- df_nsp_biofort_zinc %>%
  group_by(Crop.Name) %>%
  summarise(across(
    c(
      n_guinea,
      s_guinea,
      sudan,
      derived,
      forest,
      mid_alt,
      lowland,
      sahel,
      all_zones
    ),
    sum,
    na.rm = TRUE
  ))

df_nsp_biofort_crop_zinc_sum |> 
  DT::datatable(
    rownames = F,
    caption = "Number of varieties per Crop and AEZ with notable levels of zinc",
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

df_nsp_biofort_iron <- dplyr::filter(df_nsp, iron == 1)
paste("Number of varieties Bio-fortified with iron =", nrow(df_nsp_biofort_iron))

df_nsp_biofort_crop_iron_sum  <- df_nsp_biofort_iron %>%
  group_by(Crop.Name) %>%
  summarise(across(
    c(
      n_guinea,
      s_guinea,
      sudan,
      derived,
      forest,
      mid_alt,
      lowland,
      sahel,
      all_zones
    ),
    sum,
    na.rm = TRUE
  ))

df_nsp_biofort_crop_iron_sum |> 
  DT::datatable(
    rownames = F,
    caption = "Number of varieties per Crop and AEZ with notable levels of iron",
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

df_nsp_biofort_vit_A <- dplyr::filter(df_nsp, vit_A == 1)
paste("Number of varieties Bio-fortified with vitamin A =", nrow(df_nsp_biofort_vit_A))

df_nsp_biofort_crop_vit_A_sum <- df_nsp_biofort_vit_A %>%
  group_by(Crop.Name) %>%
  summarise(across(
    c(
      n_guinea,
      s_guinea,
      sudan,
      derived,
      forest,
      mid_alt,
      lowland,
      sahel,
      all_zones
    ),
    sum,
    na.rm = TRUE
  ))

df_nsp_biofort_crop_vit_A_sum |> 
  DT::datatable(
    rownames = F,
    caption = "Number of varieties per Crop and AEZ with notable levels of pro Vitamin A",
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))

```

# Map the varieties in each AEZ

## Step 1: Reshape varieties data to long format by AEZ

```{r}
#| label: aez_var_reshape
#| message: false
#| warning: false


# Map AEZ classes to your column names
aez_map <- c(
  "1" = "sahel",
  "2" = "sudan",
  "3" = "n_guinea",
  "4" = "s_guinea",
  "5" = "derived",
  "6" = "forest",
  "7" = "mid_alt"
)

# For each AEZ class, count varieties per crop & micronutrient
# A variety "belongs" to an AEZ if its AEZ column value == 1

df_nsp_biofort_long <- df_nsp_biofort %>%
  pivot_longer(cols = c(sahel, sudan, n_guinea, s_guinea, derived, forest, mid_alt),
               names_to = "aez_name",
               values_to = "in_aez") %>%
  filter(in_aez == 1) %>%                         # keep only varieties present in that AEZ
  mutate(AEZ = as.character(match(aez_name,        # convert name back to AEZ class number
                                   c("sahel","sudan","n_guinea",
                                     "s_guinea","derived","forest","mid_alt"))))
```

## Step 2: Count varieties per Crop, Micronutrient, and AEZ

```{r}
#| label: aez_var_num
#| message: false
#| warning: false


micronutrients <- c("zinc", "iron", "vit_A")

df_nsp_biofort_counts <- df_nsp_biofort_long %>%
  pivot_longer(cols = all_of(micronutrients),
               names_to  = "micronutrient",
               values_to = "has_trait") %>%
  filter(has_trait == 1) %>%                       # keep only varieties with that micronutrient
  group_by(Crop.Name, micronutrient, AEZ) %>%
  summarise(n_varieties = n(), .groups = "drop")

df_nsp_biofort_counts |> 
  DT::datatable(
    rownames = F,
    caption = "Number of varieties per Crop, Micronutrient, and AEZ",
    filter = "bottom",
    selection = list(mode = "none"),
    editable = FALSE,
    options = list(
      digits = 0,
      scrollX = TRUE,
      scrollCollapse = TRUE,
      autoWidth = TRUE,
      fixedColumns = list(leftColumns = 2),
      lengthMenu = c(15, 40),
      pageLength = 15
    ))


```

## Step 3: Join to SpatVector and plot

```{r}
#| label: fig-aez_var_num
#| fig-width: 12
#| fig-height: 18
#| out-width: "100%"
#| fig-align: "center"
#| fig-cap: "Number of Biofortified Varieties per AEZ"
#| message: false
#| warning: false


# Join counts to spatial data
vect_aez_geo_joined <- vect_aez_geo %>%
  mutate(AEZ = as.character(AEZ)) %>%
  left_join(df_nsp_biofort_counts, by = "AEZ", relationship = "many-to-many")

vect_aez_geo_joined <- vect_aez_geo_joined |> 
  filter(!is.na(micronutrient))|> 
  filter(!is.na(Crop.Name))


# Plot: one facet per Crop x Micronutrient combination
ggplot(vect_aez_geo_joined) +
  geom_spatvector(data = vect_aez_geo, fill = "lightgrey", color = "lightgrey", linewidth = 0.2) +
  geom_spatvector(aes(fill = n_varieties), color = "lightgrey", linewidth = 0.2) +
  scale_fill_viridis_c(option = "blues", na.value = "lightgrey",
                       name = "# Varieties", direction = -1) +
  facet_grid(Crop.Name ~ micronutrient, drop = TRUE) +
  labs(title = "Number of biofortified varieties per AEZ",
       caption = "Grey = no varieties for that crop/nutrient/AEZ combination") +
  theme_minimal() +
  theme(
    axis.text    = element_blank(),
    axis.ticks   = element_blank(),
    panel.grid   = element_blank(),
    strip.text   = element_text(face = "bold", size = 16),
    legend.position = "bottom",
    panel.spacing = unit(0.1, "lines"),
    plot.margin = margin(0, 0, 0, 0, "cm")
  )

```
